---
title: "Metacognitive uncertainty across the post-decisional response window"
author: Jesper Fischer Ehmsen^1^,
        Siebe Everaets^2,
        Arthur S. Courtin^1,3,
        Kelly Hoogervorst^1^,
        Francesca Fardo^1\*^,
        Micah G. Allen^1,4,\*^
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
  word_document:
    reference_docx: Knitting files/docx_template.docx
    toc: true
    toc_depth: 3
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
    number_sections: false
bibliography:
- "Knitting files/references.bib"
link-citations: yes
linkcolor: blue
number_sections: true
csl: Knitting files/chicago-author-date.csl
always_allow_html: yes
---

::: {custom-style="Title text"}
^1^Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University

^2^KU Leuven; Faculty of Psychology and Educational Sciences; Leuven; Belgium

^3^Institute of NeuroScience, Université catholique de Louvain, Brussels (Belgium)

^4^Cambridge Psychiatry, University of Cambridge, United Kingdom

^\*^Equal contributions

Corresponding author: Jesper Fischer Ehmsen (jesperfischer@outlook.dk)

:::

_______________________________________________________

```{r load packages, warning = F, message = F, echo = F, include=FALSE}
# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,        # Hide code by default
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  dpi = 300,          # High quality figures
  fig.align = 'center' # Center figures
)

# seed
set.seed(123)

# packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load("here", "knitr","yaml","cowplot", "magick",
               "tidyverse", "flextable","ordbetareg","patchwork","truncnorm")
```


This preregistration consists of three parts. The first two parts provide justification for the experimental design and the computational model, respectively. The third part is the main preregistration.

1. **Considerations of models and empirical data from metacognitive perceptual decision-making tasks**
2. **Outline of our computational model**
3. **Design, methods, hypotheses, and sample size considerations of the current study**

Each part is maintained in a separate document; this document includes all three.

_______________________________________________________


```{r child1, child="part1-models-data.Rmd"}
```

_______________________________________________________


```{r child2, child="part2-current-model.Rmd"}
```

_______________________________________________________

```{r child3, child="part3-design-methods.Rmd"}
```

_______________________________________________________



# References

::: {#refs}
:::