---
title: "Current model"
author: Jesper Fischer Ehmsen^1^,
        Siebe Everaets^xx^,
        Micah G. Allen^1,2,\*^,
        Francesca Fardo^1,3,\*^
output:
  html_document: default
  word_document:
    # reference_docx: Knitting files/docx_template.docx
  pdf_document:
    latex_engine: xelatex
# bibliography: 
  # - "Knitting files/References.bib"
link-citations: yes
linkcolor: blue
# csl: Knitting files/science-advances.csl
always_allow_html: yes
---

::: {custom-style="Title text"}
^1^Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University

^2^Cambridge Psychiatry, University of Cambridge, United Kingdom

^3^Danish Pain Research Center, Department of Clinical Medicine, Aarhus University

^\*^Equal contributions

Corresponding author: Jesper Fischer Ehmsen (jesperfischer@cfin.au.dk)
:::

```{r load packages, warning = F, message = F, echo = F, include=FALSE}
knitr::opts_chunk$set(dpi=300)

# seed
set.seed(123)

# packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load("here", "knitr","yaml")

knitr::opts_chunk$set(echo = TRUE)

#renv::restore(project=here::here())

required_packages = c("cowplot", "magick", "tidyverse", "flextable")
lapply(required_packages, library, character.only = TRUE)
```




### Static models of metacognition

Here we outline the general idea between the static models of perceptual deicison making and confidence judgements.

In SDT, an actor makes a binary choice $a \in \{-1,+1\}$ based on evidence $e$ about the true state $D \in \{-1,+1\}$. Evidence is drawn from a normal distribution:

$$
e_i \sim \mathcal{N}(X_iD_i,\sigma)
$$

Here inorder to incompass **3** from the Patterns of metacognition section we introduce a parameter that can shift this evidence distribution, essentially introducing a bias in the encoding of evidence.

$$
e_i \sim \mathcal{N}(X_iD_i - \beta,\sigma)
$$

where $X_i$ is stimulus strength. The actor compares evidence to criterion $c$:

$$
a_i = \begin{cases}+1 & \text{if } e_i > c \\ -1 & \text{if } e_i \le c\end{cases}
$$

For metacognition, the agent accesses this evidence for information about the correctness of the action.

Models differ in whether the rater has independent information above the evidence used to make the action.
Fleming and Daw formalized a second order rater where evidence for the action and evidence for the confidence judgement is draw from a multivariate normal with correlation $\rho$ governing information sharing. I.e. if $\rho$ = 1, the confidence judgement is based on the same information as the action, however if $\rho$ = 0 the confidence judgement is based on completely independent information.

Note we here write this model as orginally proposed by Fleming and Daw without the evidence bias

$$
\begin{bmatrix}e_i \\ r_i\end{bmatrix}\sim \mathcal{N}\left(\begin{bmatrix}X_iD_i \\ X_iD_i\end{bmatrix},\begin{bmatrix}\sigma_e^2 & \rho \cdot \sigma_e \cdot \sigma_m \\ \rho \cdot \sigma_e \cdot \sigma_m & \sigma_m^2\end{bmatrix}\right)
$$

The rater uses this information to compute $P(D = a | r_i)$, the subjective probability of being correct.

One of the main arguments for the development of this and other models with independent information for the confidence judgement was to explain error-monitoring.

Below we formalize the idea that such independent information is not necessary if instead it is assumeed that agents have varability in their choice given the evidence. I.e. instead of a deterministic decision policy agents have a noisy decision process that can produce errors even though the evidence is in favor of a different choice.

## Choice Variability

We extend the standard SDT framework by introducing stochastic choice policy. Evidence is drawn as:

$$
e_i \sim \mathcal{N}(X_iD_i- \beta,\sigma_e)
$$

But now the criterion varies trial-by-trial: $c_i \sim \mathcal{N}(0,\sigma_c)$. This yields:


$$
a_i = \begin{cases}+1 & \text{if } e_i > c_i \\ -1 & \text{if } e_i \le c_i\end{cases}
$$


$$
P(a = 1 \mid D_i, X_i) = \Phi\left(\frac{D_i \cdot X_i- \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$
Or equivalently:

$$
P(a = 1 \mid e_i) = \Phi\left(\frac{e_i}{\sigma_c}\right)
$$

From choices alone, only $\sigma_e^2 + \sigma_c^2$ is identifiable, both noise sources equivalently flatten the psychometric curve. However, by including confidence ratings these can be disentangled .

For confidence, we define:

$$
\text{confidence} = P(\text{correct} \mid r, a) = P(D = a \mid r)
$$

The rater accesses noisy evidence: $r_i \sim \mathcal{N}(e_i, \sigma_m^2)$, where $\sigma_m^2$ is metacognitive noise. This can be shown to yield:

$$
P(D=1 \mid r_i, \mu_i) = g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right)
$$

giving:

$$
P(\text{correct} \mid r_i, a, \mu_i) = \begin{cases}g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases}
$$

From confidence alone, only $\sigma_e^2 + \sigma_m^2$ is identifiable. However, combining choices and confidence allows estimating all three parameters: $\sigma_e$, $\sigma_c$, and $\sigma_m$. This requires Bayesian data augmentation to estimate latent trial-by-trial evidence. As the normative probability of being correct i.e. $P(correct | r_i,a)$ is a logit-normally distributed variable it cannot be marginalized as the choice probability can.


## Simulations

We demonstrate that this model reproduces key behavioral patterns (folded-X, double-increase) under different parameter combinations. Simulations show that the ratio $\frac{\sigma_e}{\sigma_c}$ determines these patterns: 

A low ratio (high $\sigma_c$ low $\sigma_e$) produces folded-X (strong error-monitoring), while a high ratio (low $\sigma_c$ high $\sigma_e$) produces double-increase (weak error-monitoring).

Metacognitive noise $\sigma_m$ flattens confidence curves regardless of accuracy.

Note that for these simulations we assume that the bias in evidence encoding $\beta$ is 0 as this just shifts the psychometric function as well as the confidence curves horizontally

```{r}
set.seed(199742)

n_trials <- 500

simulate_interaction <- function(sigma_m, sigma_k,sigma_e) {
  X <- abs(rnorm(n_trials,2,1))
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_k)
  p_gauss <- pnorm(e / sigma_k)
  a <- rbinom(n_trials, 1, p_gauss)

  data.frame(e = e,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}


df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_k = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_k,sigma_e))) %>% 
  unnest(data)

```

Now lets turn to how evidence, the physical stimulus and accuracy interacts across parameters:

```{r, fig.height=6, fig.width=10,warning=F,message=F}

ACC = df_interaction %>%
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = X*D, y = e, col = Accuracy)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point()+
  facet_grid(sigma_e~sigma_k, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))

ACC
```

From the above plot its obvious that $\sigma_e$ scales the the range of evidence observed (see y-axis range for the different rows). When $\sigma_k$ is low (left column) choices are consistent (when the the evidence is positive and the signed stimulus is positive i.e. D = 1 then a correct response is given). When $\sigma_k$ increases this relationship essentially becomes more noisy with correct and incorrect choices being less distinguishable across the evidence = 0 line. Another way to display this is by plotting the actions instead of the accuracy: 

here the space is only split in two quadrants instead of 4.

```{r, fig.height=6, fig.width=10,warning=F,message=F}
Act = df_interaction %>%
  mutate(Action = as.factor(a)) %>% 
  ggplot(aes(x = X*D, y = e, col = Action)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point()+
  facet_grid(sigma_e~sigma_k, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))
Act
```


Next we look at the realized confidence ratings as a function of the three different sigmas:

```{r}
set.seed(199742)

n_trials <- 5000

conf_prec <- 100
c0 <- -3
c1 <- 3

simulate_interaction <- function(sigma_m, sigma_k,sigma_e) {
  X <- sample(c(0.01,1,2,5), n_trials, replace = TRUE)
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_k)
  p_gauss <- pnorm(e / sigma_k)
  a <- rbinom(n_trials, 1, p_gauss)
  C_mu <- ifelse(a == 1, 
                 plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)),
                 1 - plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)))
  

  c_mu <- pmax(0.001, pmin(0.999, C_mu))
  C <- ordbetareg::rordbeta(n_trials, c_mu, conf_prec, c(c0, c1))
  
  data.frame(e = e,
             C = C,
             C_mu = C_mu,
             # C_mu2 = C_mu2,
             ehat = ehat,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}

df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_k = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_k,sigma_e))) %>% 
  unnest(data)
```



First we can look to fix $\sigma_k = 1$ and then we look at how $\sigma_e$ & $\sigma_m$ change confidence ratings as a function of the Stimulus:

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_k == 1) %>% 
  mutate(SS = D*X) %>%
  group_by(SS,ACC, sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

From the above plot we see that when we plot confidence on the Signed Stimulus scale we end up with a somewhat double increase pattern, when the stimulus encoding uncertainty is high (i.e. 3 or 5). Furthermore we see a (")half) folded X when stimulus encoding uncertainty  is low. Generally we see that meta uncertainty makes the confidence ratings more flat lined at 0.5 regardless of their correctness.  

Now we fix $\sigma_e = 1$ and then we look at $\sigma_k$ & $\sigma_m$:

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_e == 1) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_k,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_k~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

This plot generally shows the same as the one with $\sigma_k = 1$ but opposite in pattern. I.e. high $\sigma_k$ here is more of a folded X with lower values (espeically below 1) make a double increase.

lastly we might plot $sigma_m = 1$

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_m == 1) %>%
  mutate(sigma = sigma_e/sigma_k) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_k,sigma,sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>%
  mutate(sigma_m = as.factor(sigma_m)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~sigma_k, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We see that the diagonal where $\sigma_e = \sigma_k$ that there generally is a half-folded X, with incorrect trials being flat with stimulus strength with correct trials increasing in confidence with stimulus strength. Going to the upper triangle we move into the space where $\sigma_k > \sigma_e$ where we see folded-X patterns of varying "strengths". Moving to the lower triangle we see the opposite pattern with double increase patterns emerging again with varying "strengths". 

This means, that from the models perspective, double increase and Folded-X patterns arise from the ratio of uncertainty in the encoding of the stimulus and the choice variability. When choice variability is high compared to stimulus encoding a folded X pattern arises whereas choice variability is low compared to stimulus encoding a double increase pattern arises. $\sigma_m$, meta uncertainty is agnostic to the accuracy of the rater and when it increases it makes the curves of the confidence ratings more flat regardless of correctness.

## Fitting

We fit this model using full Bayesian inference with Stan's NUTS sampler, treating latent trial-by-trial evidence as parameters estimated jointly with the noise parameters. This model however given our pilot data is not particularly informed by the data, meaning that the sampler diagnostics aren't particularly great for all pilot data-sets and especially not when during hierarchical bayesian estimation, i.e. estimating group level parameters which subjects level parameters are drawn from. We therefore build a heurestic model that approximates the above model, without the need for the trial-by-trial evidence. This is done such that hierarchical bayesian inference can be done reliably for the Sequential sampling approach used for sample size determined in the current study (see Sequential sampling section).

## A heuristic model

While the mechanistic model is powerful, its computational complexicirty (especially in hierarchical settings) motivate a simpler heuristic approximation. The key challenge is that confidence requires trial-specific evidence $r_i$, which cannot be marginalized analytically.

**Heuristic approach**: Instead of computing confidence from trial-specific evidence, we compute it from the psychometric function:

$$
P(a = D) = \Phi\left(\frac{|D_i \cdot X_i - \beta|}{\sigma_e + \sigma_\gamma}\right)
$$

$$
P(a \neq D) = \Phi\left(\frac{|D_i \cdot X_i - \beta|}{\sigma_\eta}\right)
$$

with $\sigma_\eta \in c(-\inf,\inf)$ and $\sigma_\gamma \in c(0,\inf)$. These simplifying assumptions make it such that the confidence curve can only be as steep as the psychometric curve i.e. when $\sigma_\gamma = 0$ and never decreasing with stimulus intensity (i.e. being flat with increasing stimulus intensity when $\sigma_\gamma = \inf$). For incorrect trials we assume that there can be an increase or decrease with stimulus intensity depending on the sign of  $\sigma_\eta \in c(-\inf,\inf)$ with negative values being a folded X pattern and positive values being a double increase pattern.
Below we will show that this Heurestic model is computationally much less complex and can be fit without issues. Furthermore, we will show that it provides a useful heurestic to the parameter of interest $\sigma_m$ which is the above $\sigma_\gamma$. Lastly, as this model does not estimate the choice varability and thus will assume a steeper psychometric slope we allow for the asympotes of the psychometric function to not converge to 0 and 1 but to a particular lapse rate. This is a common assumption when analyzing purely binary data, as not accounting for flipped responses at high stimulus intensities can greatly comprimise the estimation of the psychometric slope (i.e. $\sigma_e$)

**Critical distinction**: The mechanistic model computes $\mathbb{E}_{r_i}[g(r_i)]$ (average confidence over evidence), while the heuristic computes $g(\mathbb{E}_{r_i}[r_i])$ (confidence from average evidence). 

The mechanistic model marginalizes after computing confidence (via MCMC), while the heuristic marginalizes before. By Jensen's inequality, these differ. The heuristic requires separate slopes for correct/error trials due to different conditional evidence distributions.



**Asymmetry**: The mean difference between these two models arise in the incorrect trials. If the above formulations did not have a separate parameter for the denominator for incorrect trials the argued for confidence patterns couldn't emerge.


## simulations Heuristic model

```{r}
set.seed(199742)

n_trials <- 5000

simulate_interaction <- function(sigma_mc, sigma_mi,sigma_e) {
  X <- sample(c(0.01,1,2,5), n_trials, replace = TRUE)
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  
  
  p_gauss <- pnorm(D*X / sigma_e)
  a <- rbinom(n_trials, 1, p_gauss)
  ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0))

  C_mu <- ifelse(ACC == 1, 
                 pnorm(abs(D*X)/ (sigma_e + sigma_mc)),
                 pnorm(sigma_mi * abs(D*X)))
  
  
  

  data.frame(C_mu = C_mu,
             a = a,
             X = X,
             D=D,
             p_gauss = p_gauss,
             # sigma_e = sigma_e,
             # sigma_mi = sigma_mi,
             # sigma_mc = sigma_mc,
             ACC = ACC)
             

   }

df_interaction <- expand.grid(
  sigma_mc = c(0.1, 1 ,3),
  sigma_mi = c(-0.2, 0 ,0.2),
  sigma_e = c(0.1, 0.5 ,1)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_mc, sigma_mi,sigma_e))) %>% 
  unnest(data)



```



First we can look to fix $\sigma_k = 1$ and then we look at how $\sigma_e$ & $\sigma_m$ change confidence ratings as a function of the Stimulus:

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_e == 1) %>% 
  mutate(SS = D*X) %>%
  group_by(SS,ACC, sigma_mi,sigma_mc) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_mi~ sigma_mc, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction <- expand.grid(
  sigma_mc = c(0, 1 ,3),
  sigma_mi = c(-0.2, 0 ,0.2),
  sigma_e = c(0.5 ,1, 2)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_mc, sigma_mi,sigma_e))) %>% 
  unnest(data)


df_interaction %>%
  mutate(SS = D*X) %>%
  group_by(SS,ACC, sigma_mi,sigma_mc,sigma_e) %>% 
  summarize(mean_conf = mean(C_mu),
            mean_a = mean(p_gauss)) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean_a))+
  geom_point(aes(x = SS, y = mean_conf, col = Accuracy))+
  geom_point()+
  theme_classic()+
  facet_wrap(sigma_mc~sigma_e, labeller = label_both)
```


```{r, fig.height=6, fig.width=10, message=F, warning=F}
# Step 1: Create basic psychometric curves for two sigma_e values
df_psychometric <- expand.grid(
  sigma_mc = 0,  # No metacognitive noise for pure psychometric
  sigma_mi = c(-0.1, 0, 0.1),  # Different bias levels (not 0 anymore)
  sigma_e = c(1, 2)  # Two evidence noise levels
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_mc, sigma_mi, sigma_e))) %>% 
  unnest(data) %>%
  mutate(SS = D*X) %>%
  group_by(SS, sigma_e, sigma_mi, ACC) %>% 
  summarize(mean_a = mean(p_gauss), 
            mean_conf = mean(C_mu), .groups = 'drop')

 df_psychometric %>%
  ggplot(aes(x = SS, y = mean_a)) +
  geom_line(aes(linetype = factor(sigma_e)), 
            color = "black", 
            size = 1, 
            alpha = 0.7) +
  # Confidence curves for incorrect trials (colored)
  geom_line(data = filter(df_combined, ACC == 0),
            aes(y = mean_conf, color = sigma_mi_label),
            size = 1.2) +
  scale_linetype_manual(
    values = c("1" = "dotted", "3" = "dashed"),
    labels = c("1" = "σ_e = 1", "3" = "σ_e = 3"),
    name = "Evidence Noise"
  ) +
   scale_color_manual(values = c("red4","red","orange"))
  labs(x = "Stimulus Strength", 
       y = "P(a)") +
  theme_classic() +
    theme(legend.position = "none")
  # theme(legend.position = c(0.4,0.25),
        # legend.title = element_blank())


```



```{r}
# Step 1: Create dataframe with both psychometric curves AND confidence curves
df_combined <- expand.grid(
  sigma_mc = 0,  # No metacognitive noise for now
  sigma_mi = c(-0.2, 0, 0.2),  # Different bias levels (not 0 anymore)
  sigma_e = c(1, 3)  # Two evidence noise levels
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_mc, sigma_mi, sigma_e))) %>% 
  unnest(data) %>%
  mutate(SS = D*X) %>%
  group_by(SS, sigma_e, sigma_mi, ACC) %>% 
  summarize(mean_a = mean(p_gauss), 
            mean_conf = mean(C_mu), .groups = 'drop') %>%
  mutate(sigma_mi_label = case_when(
    sigma_mi == -0.2 ~ "σ_mi < 0",
    sigma_mi == 0 ~ "σ_mi = 0", 
    sigma_mi == 0.2 ~ "σ_mi > 0"
  ))

# Plot both psychometric (for sigma_mi = 0 only) and confidence curves
df_combined %>%
  ggplot(aes(x = SS)) +
  # Psychometric curves (black background) - only for unbiased case
  geom_line(data = filter(df_combined, sigma_mi == 0),
            aes(y = mean_a, linetype = factor(sigma_e)), 
            color = "black", 
            size = 1, 
            alpha = 0.7) +
  # Confidence curves for incorrect trials (colored)
  geom_line(data = filter(df_combined, ACC == 0),
            aes(y = mean_conf, color = sigma_mi_label),
            size = 1.2) +
  geom_point(data = filter(df_combined, ACC == 0),
             aes(y = mean_conf, color = sigma_mi_label),
             size = 2) +
  scale_linetype_manual(
    values = c("1" = "dotted", "3" = "dashed"),
    labels = c("1" = "σ_e = 1", "3" = "σ_e = 3"),
    name = "Evidence Noise"
  ) +
  scale_color_viridis_d(name = "Metacognitive Bias\n(Incorrect Trials)", 
                        option = "plasma") +
  labs(x = "Stimulus Strength", 
       y = "P(a) / Confidence",
       title = "Psychometric Functions + Confidence for Incorrect Trials") +
  theme_classic() +
  theme(legend.position = "right")
```






```{r, fig.height=6, fig.width=10,warning=F,message=F}
# Compute approximation errors
error_summary <- df_interaction %>%
  group_by(sigma_e, sigma_m, sigma_k, ACC, X) %>%
  summarize(
    mean_mechanistic = mean(C_mu),
    mean_heuristic = mean(C_mu2),
    abs_error = abs(mean(C_mu) - mean(C_mu2)),
    n_trials = n(),
    .groups = "drop"
  )

error_summary %>%
  filter(ACC == 1) %>% 
  group_by(sigma_e, sigma_m, sigma_k, ACC) %>%
  summarize(
    mean_abs_error = mean(abs_error),
    q5_abs_error = quantile(abs_error,0.05),
    q95_abs_error = quantile(abs_error,0.95),
    .groups = "drop"
  ) %>% ggplot(aes(x = sigma_e, y = mean_abs_error, col = as.factor(ACC)))+
  geom_pointrange(aes(ymin = q5_abs_error,ymax = q95_abs_error))+
  facet_grid(sigma_k~sigma_m,scales = "free", labeller = label_both)


error_summary %>%
  filter(ACC == 0) %>% 
  group_by(sigma_e, sigma_m, sigma_k, ACC) %>%
  summarize(
    mean_abs_error = mean(abs_error),
    q5_abs_error = quantile(abs_error,0.05),
    q95_abs_error = quantile(abs_error,0.95),
    .groups = "drop"
  ) %>% ggplot(aes(x = sigma_e, y = mean_abs_error, col = as.factor(ACC)))+
  geom_pointrange(aes(ymin = q5_abs_error,ymax = q95_abs_error))+
  facet_grid(sigma_k~sigma_m,scales = "free", labeller = label_both)
```


```{r, fig.height=6, fig.width=10,warning=F,message=F}
df_interaction %>%
  filter(ACC == 1) %>% 
  mutate(dif = C_mu-C_mu2)  %>% 
  ggplot(aes(x = as.factor(sigma_e), y = dif, col = as.factor(ACC)))+
  geom_boxplot()+
  facet_grid(sigma_k~sigma_m,scales = "free", labeller = label_both)




df_interaction %>%
  filter(ACC == 0) %>% 
  mutate(dif = C_mu-C_mu2)  %>% 
  ggplot(aes(x = as.factor(sigma_e), y = dif, col = as.factor(ACC)))+
  geom_boxplot()+
  facet_grid(sigma_k~sigma_m,scales = "free", labeller = label_both)


```




