
<!-- Part 2: Current Model -->

```{r packages3, warning = F, message = F, echo = F, include=FALSE}
# Load additional packages needed for this section
if (!require("pacman")) install.packages("pacman")
pacman::p_load("here", "knitr","yaml")

#renv::restore(project=here::here())

required_packages = c("cowplot", "magick", "tidyverse", "flextable","truncnorm")
lapply(required_packages, library, character.only = TRUE)

```

# Part 2: Current Model

In this section we first describe the general structure of ideas of static models of perceptual decision making and confidence ratings. Next we describe, derive and simulate from our proposed model, demonstrating that it can produce some important patterns highlighted in the [Design considerations](Models and Data in Perceptual Decision-Making with Confidence.Rmd).

## Static models of metacognition

In SDT, an actor makes a binary choice on each trial i, $a_i \in \{-1,+1\}$ based on some internal evidence $e_i$. The action reflects the actor's inference on the true state of the stimulus $D_i \in \{-1,+1\}$. 

In SDT, evidence is drawn from a normal distribution, as perception is a noisy process. The mean of this distribution reflects the true state and scales with stimulus intensity $X_i$. 
$$
e_i \sim \mathcal{N}(X_iD_i,\sigma_e)
$$

The actor then produces an action by comparing the internal evidence to a criterion $c$:

$$
a_i = \begin{cases}+1 & \text{if } e_i > c \\ -1 & \text{if } e_i \le c\end{cases}
$$

Within most frameworks, this criterion is thought to be static within an experimental setting for each agent.

#### Confidence Ratings

Having made a binary choice, the agent is asked to give a confidence judgement in their choice. A particularly prominent way to describe this is assume that the agent computes the probability that the action was correct.  @FLEMINGSelfEvaluationDecisionMakingGeneralBayesianFrameworkMetacognitive2017 proposed an approach within the framework discussed above that could also account for error-monitoring.<!-- this goes very fast, i feel like it needs some transition from SDT confidence (P(correct)) to fleming and daw. Feels kinda sudden now-->

<!-- The way i view this section is just briefly mentioning what others have done and why. The most relevant here is to get SDT across with static citerion and some model that accounts for error monitoring, but also has this independent information aspect.-->


In their formulation confidence judgement are based on <!-- I would mention it is based on ehat, now it feels like confidence is being drawn from a MVN -->draws from a multivariate normal, which contains the evidence used for the action $e_i$ and  pseudo-independent information only accessible to the rater $r_i$. These two quantities are assumed to be correlated with $\rho$ governing information sharing. I.e. if $\rho$ = 1, the confidence judgement is based on the same information as the action, however if $\rho$ = 0 the confidence judgement is based on completely independent information.

$$
\begin{bmatrix}e_i \\ r_i\end{bmatrix}\sim \mathcal{N}\left(\begin{bmatrix}X_iD_i \\ X_iD_i\end{bmatrix},\begin{bmatrix}\sigma_e^2 & \rho \cdot \sigma_e \cdot \sigma_m \\ \rho \cdot \sigma_e \cdot \sigma_m & \sigma_m^2\end{bmatrix}\right)
$$

The rater is then thought to use this information to compute the subjective probability of being correct which is used to generate a confidence rating. One of the main arguments for such model and other models with independent information was to explain error-monitoring i.e. agents' ability to know that they produced an error and give low (even below guessing) confidence ratings. The necessity of independent information for error monitoring follows from the fact that a rational agent who bases both choice and confidence on the same information (i.e. $\rho = 1$ in the above model), and whose choices are deterministic, cannot make an error while being aware of it. Any information indicating that the choice is incorrect would also determine the choice itself, thereby preventing the error


## Efference copy (Choice Variability) model of metacognition (EA/CV - Meta)

### Choices

Below we will argue that independent information is not required to explain error monitoring if we instead assume that agents show variability in their choices given the evidence. The difference from the above account is that instead of a deterministic decision policy (which is usually assumed in SDT), agents have a noisy decision process that can produce errors even though the evidence is in favor of the correct choice. Such errors could be related to speed stress conditions or more generally motor errors. By extending the standard SDT framework with the introduction of a stochastic choice policy it is possible for agents to display error monitoring. One way to implement this is to assume that the criterion varies from trial-by-trial: $c_i \sim \mathcal{N}(0,\sigma_c)$. This gives us:

$$
a_i = \begin{cases}+1 & \text{if } e_i > c_i \\ -1 & \text{if } e_i \le c_i\end{cases}
$$

Due to the fact that both $e_i$ and $c_i$ are normally distributed, it is possible to marginalize their difference to obtain the probability of the action:

$$
P(a = 1 \mid D_i, X_i) = \Phi\left(\frac{D_i \cdot X_i}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$

Before moving on to the confidence ratings we also assume that the evidence distribution might be biased for each subject. This assumption will help explain the fact that the psychometric function of the choices generally shift the other behavioral variables such as confidence ratings [@CAZIOTPerceptualConfidenceJudgmentsReflectSelfconsistency2021; @SANCHEZ-FUENZALIDAConfidenceReportsPerceptualDecisionMakingDissociate2025; @MIHALIIntrospectiveInferenceCounteractsPerceptualDistortion2023a; @NAVAJASIdiosyncraticNatureConfidence2017].

This leads to:

$$
e_i \sim \mathcal{N}(X_iD_i- \beta,\sigma_e)
$$
And when marginalized

$$
P(a = 1 \mid D_i, X_i) = \Phi\left(\frac{D_i \cdot X_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right) \tag{1}
$$

From choices alone, only the sum $\sigma_e^2 + \sigma_c^2$ can be identified, as both noise sources equivalently flatten the psychometric function. However, by including confidence ratings these can be disentangled.

### Confidence ratings

To compute confidence ratings, we assume that the agent internally computes the subjective probability that the choice was correct.

$$
\text{confidence} = P(\text{correct} \mid r, a) = P(D = a \mid r)
$$
Here we assume that the rater has access to a noisy readout of the evidence: $r_i \sim \mathcal{N}(e_i, \sigma_m)$, where $\sigma_m$ is metacognitive noise. 

This can be shown to yield: (See the full derivation in [Additional Derivations](Additional Derivations/inverselogit.Rmd)).

```{r, child="Additional Derivations/inverselogit.Rmd", eval = knitr::is_html_output()}
```

$$
P(\text{correct} \mid r_i, a, X_i) = \begin{cases}g\left(\frac{2 \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - g\left(\frac{2 \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases} \tag{2}
$$

Given that we assume a uniform prior on D, i.e. $P(D_i = 1) = P(D_i = -1) = 0.5$, which represents the true state of the world (i.e. in the Random Dot Motion task whether the dots are moving up or down).

Where $g(.)$ is the inverse logit transformation i.e. $g(x)= \frac{1}{1+exp(-x)}$. 

From confidence alone, only the sum $\sigma_e^2 + \sigma_m^2$  can be identified. However, combining choices and confidence allows estimating all three parameters: $\sigma_e$, $\sigma_c$, and $\sigma_m$. 

As can be seen from equation (2), to compute the probability of being correct the rater needs access to the stimulus intensity, $X_i$. Even though that it is assumed in some models [@FLEMINGNeuralMediatorsChangesMindPerceptualDecisions2018a], we believe that the rater cannot have access to the objective stimulus intensity. We will get back to this and a final formulation without this term later. Interested readers are referred to [@RAUSCHFoldedXpatternNotNecessarilyStatisticalSignature2019], were simulations are conducted with different generative processes for the stimulus intensity.

Below we showcase a plate notation schematic of the above proposed model, demonstrating a visualization of the noise parameters. 


```{r platenotation, out.width="100%", fig.align='center', message=F, error=F, warning=F}

invisible(tinytex::latexmk(here::here("Preregistration","Figure scripts","Platenotations", "platenotation.tex")))

# For PDF output, include the PDF directly; for HTML/Word, convert to image
if (knitr::is_latex_output()) {
  knitr::include_graphics(here::here("Preregistration", "Figure scripts","Platenotations", "platenotation.pdf"))
} else {
  img1 <- image_read_pdf(here::here("Preregistration", "Figure scripts","Platenotations", "platenotation.pdf"),
                         density = 600)
  img1 <- image_trim(img1)
  img1 <- image_resize(img1, "1200x")
  print(img1)
}
```

The agent is presented with a stimulus that has a direction $D_i$ (the true state of the world) and a strength $X_i$. An internal representation of the physical stimulus (evidence) is accessible to the agent $e_i$ (red point), which is a noisy readout of the physical stimulus. This internal evidence representation is then sent to the pre-motor and motor cortex where an efferent signal is produced together with an efference copy (red split path). Due to additional noise ($\sigma_c$)  in the processing and transmission of the signal to the limb generating the action, it is possible that the action generated, is in the opposite side of what the evidence entailed (green point). Simultaneously the efference copy might also be subjected to additional noise ($\sigma_m$), which is used together with the action to produce a prediction error in the comparator $C_t$. This prediction error is then used to compute confidence ratings $C_t$. <!-- We should definitely elaborate on the underlying neural representation more and earlier, reading it in the annotation feels weird--> 

Next, we present simulations from the model to build intuition about the parameters and how they interact.

### Simulations

Below we demonstrate that the model reproduces key behavioral patterns of choice and confidence ratings under different parameter combinations. Foreshadowing the results, our simulations show that the ratio $\frac{\sigma_e}{\sigma_c}$ determines the degree of error-monitoring and the main confidence patterns of folded-X or double increase.

<!-- Same here: not sure about weak error-monitoring, also we should mention here that the second one is why metanoise can be seen as a measure of metaC efficiency--> 

<!-- Not sure i agree with you in how these terms (efficiency and sensitivity) are used in the meta-cognitive literature. I would argue that both relate correct and incorrect choices to confidence ratings with efficiency being dependent on stimulus dificulty. So metanoise is related but not the main player, the ratio is the main contributor to see a difference between correct and incorrect trials in the confidence ratings--> 



* A low ratio (high $\sigma_c$ low $\sigma_e$) produces folded-X (strong error-monitoring), while a high ratio (low $\sigma_c$ high $\sigma_e$) produces double-increase (no error-monitoring). 


* The metacognitive noise $\sigma_m$ flattens the confidence curves regardless of accuracy.

Note: For all simulations below we assume that the bias in evidence encoding $\beta$ is 0. We will get back to this, which relates to the assumption of subjects have access to the objective stimulus intensity.

```{r}
set.seed(199742)

n_trials <- 500

simulate_interaction <- function(sigma_m, sigma_c,sigma_e) {
  X <- abs(rnorm(n_trials,2,1))
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_c)
  p_gauss <- pnorm(e / sigma_c)
  a <- rbinom(n_trials, 1, p_gauss)

  data.frame(e = e,
             ehat = ehat,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}


df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_c = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  as_tibble() %>% 
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_c,sigma_e))) %>% 
  unnest(data)


```

First, we might look at the distribution of internal evidence over the objective signed stimulus (i.e. the product of the stimulus intensity and the true state) across the two parameters governing the choices i.e. $\sigma_e$ and $\sigma_c$ being the evidence encoding and choice variability respectively.

```{r, fig.height=6, fig.width=10,warning=F,message=F}

ACC = df_interaction %>%
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = X*D, y = e, col = Accuracy)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point(alpha = 0.2)+
  facet_grid(sigma_e~sigma_c, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))

ACC
```

From the above plot we see that $\sigma_e$ broadens the range of evidence observed (essentially making it more noisy based on the same external stimulus), moving down the rows the y-axis greatly changes its range. When $\sigma_c$ is low (left column) choices are consistent (when the the evidence is positive and the signed stimulus is positive (i.e. D = 1), a correct response is given). When $\sigma_c$ increases this relationship essentially becomes more noisy, such that correct and incorrect choices are less distinguishable across the zero-evidence line and zero-signed-stimulus line. 

Another way to display this is by plotting the actions instead of the accuracy: 

```{r, fig.height=6, fig.width=10,warning=F,message=F}
Act = df_interaction %>%
  mutate(Action = as.factor(a)) %>% 
  ggplot(aes(x = X*D, y = e, col = Action)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point(alpha = 0.2)+
  facet_grid(sigma_e~sigma_c, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))
Act
```

Next we look at the confidence ratings as a function of the three different sigmas.

```{r}
set.seed(199742)

n_trials <- 5000

conf_prec <- 100
c0 <- -3
c1 <- 3

simulate_interaction <- function(sigma_m, sigma_c,sigma_e) {
  X <- sample(c(0.01,1,2,5), n_trials, replace = TRUE)
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_c)
  p_gauss <- pnorm(e / sigma_c)
  a <- rbinom(n_trials, 1, p_gauss)
  C_mu <- ifelse(a == 1, 
                 plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)),
                 1 - plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)))
  

  c_mu <- pmax(0.001, pmin(0.999, C_mu))
  C <- ordbetareg::rordbeta(n_trials, c_mu, conf_prec, c(c0, c1))
  
  data.frame(e = e,
             C = C,
             C_mu = C_mu,
             ehat = ehat,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}

df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_c = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_c,sigma_e))) %>% 
  unnest(data)
```

First we fix $\sigma_c = 1$ and then we look at how $\sigma_e$ & $\sigma_m$ change confidence ratings as a function of the signed stimulus intensity: 

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_c == 1) %>% 
  mutate(SS = D*X) %>%
  group_by(SS,ACC, sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  # geom_vline(xintercept = 0, linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  geom_line()+
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

From the above plot we see that when we plot confidence on the Signed Stimulus scale we end up with a double increase pattern when $\sigma_e$ is high (i.e. moving down the facets). Furthermore, when $\sigma_e$ is low (moving up facets) we see a folded X pattern. Lastly, when meta uncertainty ($\sigma_m$) increase (moving left to right in the facets) both confidence curves becomes increasingly flat (i.e. closer to a flat line at 0.5) regardless of their correctness. 

Next we fix $\sigma_e = 1$ and then we look at $\sigma_c$ & $\sigma_m$:


```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_e == 1) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_c,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_c~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
    geom_line()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Here we observe a similar patterns as before with $\sigma_c = 1$. Here a high $\sigma_c$ reflects a folded X, whereas lower values (especially below 1) make a double increase. As before when meta uncertainty ($\sigma_m$) increase both confidence curves becomes increasingly flat (i.e. closer to a flat line at 0.5) regardless of their correctness. 

Lastly we plot the simulated data for varying $\sigma_c$ and $\sigma_e$ when $\sigma_m = 1$

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_m == 1) %>%
  mutate(sigma = sigma_e/sigma_c) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_c,sigma,sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>%
  mutate(sigma_m = as.factor(sigma_m)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~sigma_c, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
      geom_line()+

  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Here we see that when $\sigma_e = \sigma_c$ (the diagonal), it becomes difficult to distinguish a folded‑X pattern from a double‑increase pattern, because incorrect trials cluster around 0.5 and show little dependence on stimulus intensity. Correct trials increase in confidence across all parameter combinations with stimulus intensity. Looking at the upper triangle (where $\sigma_c > \sigma_e$), we see folded-X patterns, whereas in the lower triangle ($\sigma_c < \sigma_e$) we see the opposite pattern, double increase patterns. 
<!-- Can we actually give meaning to this ratio, it would be cool if we can show "this is why it produces these patterns"--> 
<!-- Yea lets think about this for the paper!"--> 


These simulations show that from the current model's perspective, the double increase and Folded-X pattern arise due to different ratios of uncertainty in the encoding and choice. When choice variability is high compared to stimulus encoding, a folded X pattern arises whereas choice variability is low compared to stimulus encoding a double increase pattern arises. Meta uncertainty ($\sigma_m$) is agnostic to the accuracy of the rater and when it increases it makes the confidence curves more flat regardless of correctness.

### Fitting

We want to fit this model using full Bayesian inference with Stan's NUTS sampler, treating the latent trial-by-trial evidence as parameters estimated jointly with the subject-by-subject noise parameters. Investigating our pilot data we saw that on a single-subject-level this model was possible to fit with sufficiently good sampler diagnostics. However, trying to fit this model in a hierarchical framework (i.e. drawing subject level parameters from group level parameters), we ran into sampling issues. As the current project is going to use a sequential sampling paradigm, a robust and efficient model is necessary. We therefore build a heuristic model that approximates the model shown above, without the need for estimation of trial-by-trial evidence.

### The Heuristic model

While the above model is powerful, its computational complexity (especially in hierarchical settings) motivate a simpler heuristic approximation. The key challenge is that confidence requires trial-specific evidence $r_i$, which cannot be marginalized analytically, because the random variable is inside the inverse logit function. 

<!-- is it actually impossible to marginalize or are we just too stupid? lol --> 
<!-- Hahaha, well i think with current methods it is not possible to do analytically --> 

The inverse logit function and the cumulative normal distribution are in practice close to being indistinguishable up to a constant i.e. $g(.) \sim \Phi(\frac{1}{1.7})$, see plot below:

```{r, fig.height=6, fig.width=10,warning=F,message=F}
# Compare inverse logit with cumulative normal approximation
data.frame(
  x = seq(-6, 6, by = 0.01)
) %>% 
  mutate(
  `Inverse Logit g(x)` = plogis(x),
  `Scaled CDF Φ(x/1.7)` = pnorm(x / 1.7)) %>% 
  pivot_longer(cols = -x, names_to = "Function", values_to = "Value") %>% 
  ggplot(aes(x = x, y = Value, color = Function, linetype = Function)) +
  geom_line(size = 1) +
  labs(x = "x", y = "Probability",
       title = "Comparison of Inverse Logit and Scaled Cumulative Normal") +
  theme_minimal() +
  theme(text = element_text(size = 12), legend.position = "top")
```

Using this approximation it is possible to marginalize out the trial-by-trial evidence and arrive at a computationally less expensive model (and an equation for the expectation of confidence) that uses the same parameters. 

With this approximation we go from equation (2) to the following:

$$
P(\text{correct} \mid r_i, a, X_i) = \begin{cases}\Phi\left(\frac{C \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - \Phi\left(\frac{C \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases} \tag{3}
$$

Where $C = \frac{2}{1.7}$. 

Focusing on the instance where the choice was one.

$$
\Phi\left(\frac{C \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right)
$$

We want to marginalize over $r_i$, i.e. calculate the expectation of above equation, across $r_i$. 

This can be done using the general formula for the expectation of a random continuous variable.

$$
\mathbb{E(X)} = \int_{-\infty}^\infty x \ p(x) dx
$$

Substituting the variables, with the following $b = \frac{CX_i}{\sigma_e^2 + \sigma_m^2}$

$$
\int_{-\infty}^\infty \Phi(b \cdot r_i) \ p(r_i) dr_i
$$

with $r_i \sim \mathcal{N}(\mu,\sigma)$

This integral can be shown to have the [solution](https://www.tandfonline.com/doi/abs/10.1080/03610918008812164):

$$
\Phi \left(\frac{\mu \cdot b}{\sqrt{1+\sigma^2 \cdot b^2}}\right)
$$


Given that we have $r_i \sim \mathcal{N}(X_iD_i - \beta, \sqrt{\sigma_m^2+\sigma_e^2})$, we get: 

$$
\Phi \left(\frac{\mathbb{E}(r_i) \cdot b}{\sqrt{1+(\sigma_e^2+\sigma_m^2) \cdot b^2}}\right)
$$

Here we write the expectation of $r_i$, instead of the mean of the $r_i$ distribution ($X_iD_i - \beta$), as this will be conditional on the choice (remember we are looking at $a_i = 1$). 

Writing everything out and simplifying, we obtain:

$$
\Phi\left(\frac{\mathbb{E}(r_i) \cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)
$$

This equation gives the probability of being correct when the choice is 1. However, because the rater knows the actor’s choice, and the rater’s evidence is a noisy readout of the evidence used to generate that choice, observing the choice provides the rater with additional information about the underlying evidence. Mathematically this means that if the rater uses all the available evidence then the expectation of the available evidence needs to be conditional on the actors choice.


Equation (3), therefore now becomes:


$$
P(\text{correct} \mid a, X_i) = \begin{cases}
\Phi\left(\frac{\mathbb{E}(r_i \mid a = 1) \cdot CX_i}{\sigma_e^2 + \sigma_m^2} \cdot \frac{1}{\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right) & \text{if } a = 1 \\ 
1 - \Phi\left(\frac{\mathbb{E}(r_i \mid a = -1) \cdot CX_i}{\sigma_e^2 + \sigma_m^2} \cdot \frac{1}{\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right) & \text{if } a = -1
\end{cases} \tag{4}
$$


It can be shown that this expectation given the choice is equal to the following equation: (For a full walk through of the derivation see below)


```{r, child="Additional Derivations/Conditional Expectations.Rmd", eval = knitr::is_html_output()}
```



$$
\mathbb{E}(r_i \mid a = 1) = (X_iD_i - \beta) + \rho \cdot \sqrt{\sigma_e^2 + \sigma_m^2} \cdot \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$

For $a = -1$.

$$
\mathbb{E}(r_i \mid a = -1) = (X_iD_i - \beta) - \rho \cdot \sqrt{\sigma_e^2 + \sigma_m^2} \cdot \lambda\left(-\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$

Here $\lambda(x) = \frac{\phi(x)}{\Phi(x)}$, and $\rho = \frac{\sigma_e^2}{\sqrt{(\sigma_e^2 + \sigma_c^2)} \cdot \sqrt{(\sigma_e^2 + \sigma_m^2)}}$


### Final confidence model

Substituting these conditional expectations back into our marginalized confidence equation (4). This gives us after simplifying:

$$
P(correct \mid a = 1, X_i, D_i) = \Phi\left(\left[\frac{(CX_i^2D_i - C\beta X_i)}{\sigma_e^2 + \sigma_m^2} + \rho \cdot \frac{ \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)CX_i}{\sqrt{\sigma_e^2 + \sigma_m^2} }\right] \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)
$$


The above equation gives us a fully marginalized model for confidence ratings (for choices of 1) that depends only on the parameters $\Theta_1 = (\beta, \sigma_e, \sigma_c, \sigma_m)$ and the observed data (choices $a_i$, stimuli $X_i$ and true states $D_i$), without requiring trial-by-trial estimation of latent evidence variables $e_i$ and $r_i$.

Before interpreting the equation we need to revert back to one of the key assumptions of this model as was outline just after introducing equation (2). When the subject computes the subjective probability of being correct it is assumed that the subject has access to the stimulus intensity. Below we show which of the stimulus intensities this is related to (the two $X_i$ terms that we highlight):

$$
P(\text{correct} \mid a = 1, X_i, D_i) = \Phi\left(\frac{\left[(X_iD_i - \beta) + \rho \cdot \sqrt{\sigma_e^2 + \sigma_m^2} \cdot \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)\right] \cdot C \mathbf{X_i}}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2\mathbf{X^2_i}}{\sigma_e^2+\sigma_m^2}}}\right) 
$$

Interestingly these two stimulus intensities (shown in bold), especially the first one (which is not from the marginalization) serves to make sure that at intensity 0 the predicted confidence will always be 0.5. So not only is it a strong assumption that the agent has access to the objective stimulus strength, this assumption does not even hold given the empirical observation that the lowest confidence rating shifts with the bias which we highlighted in part 1. If confidence ratings are treated as noisy, biased subjective probabilities of being correct, the formulation above becomes problematic due to its reliance on an objective anchor at $X_i = 0$. Due to these consideration, supported by some empirical validation on our pilot data below, we believe that this particular stimulus intensity, coming from an unjustified assumption, needs to be accessed. 


One interesting idea of how to deal with this, is to assume that the agent has a noisy readout of the stimulus strength, $\hat{X} \sim p(\hat{x}|\theta,c)$. The problem with such formulation is defining the probability distribution of the internal stimulus intensity. This probability distribution needs to be positively bound as $\hat{X} > 0$, which introduces an arbitrary choice, but more importantly makes it non trivial to marginalize the trial-by-trial readout. Another approach would be to replace this stimulus intensity value with a constant, which could vary across subjects. Such parameterization would serve as a placeholder for idiosyncratic differences in how subjects might use the stimulus intensity above and beyond what they should have access to. 

Instead of opting for these options, we chose to assume that the agents do not have access to this stimulus intensity and therefore fix it to 1. This approach in its simplicity circumvents some of the troublesome computational aspects, as well as obeying to what the agents is assumed to have access to. This also importantly this makes it possible for the model to accommodate the empirical observation that the confidence curve follows the psychometric function of the binary choices. Below we will showcase that from our pilot data the two models (fixing this particular stimulus intensity to 1, or letting it vary based on the true stimulus intensity) have equivalent approximate leave one out cross-validation performance.

This means that the final equations become

$$
P(\text{correct} \mid a = 1, X_i, D_i) = \Phi\left(\left[\frac{(CX_iD_i - C\beta)}{\sigma_e^2 + \sigma_m^2} + \rho \cdot \frac{ \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)C}{\sqrt{\sigma_e^2 + \sigma_m^2} }\right] \cdot\frac{1} {\sqrt{1 + \frac{C^2}{\sigma_e^2+\sigma_m^2}}}\right) \tag{7a}
$$


which for a = -1 becomes:

$$
P(\text{correct} \mid a = -1, X_i, D_i) = 1-\Phi\left(\left[\frac{(CX_iD_i - C\beta)}{\sigma_e^2 + \sigma_m^2} - \rho \cdot \frac{ \lambda\left(-\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)C}{\sqrt{\sigma_e^2 + \sigma_m^2} }\right] \cdot\frac{1} {\sqrt{1 + \frac{C^2}{\sigma_e^2+\sigma_m^2}}}\right) \tag{7b}
$$

with:

$$
\begin{aligned}
\lambda(x) &= \frac{\phi(x)}{\Phi(x)} \\
C &= \frac{2}{1.7}
\end{aligned}
$$


In the above formula it is noteworthy that the terms inside the cumulative normal distribution function have clear interpretations. 

Firstly the common factor, is a correction factor for marginalization. thus we define: 

$$
K_i = \frac{1} {\sqrt{1 + \frac{C^2}{\sigma_e^2+\sigma_m^2}}}
$$

The two other terms can be thought to have more cognitive interpretations. Looking back at equation 3 we had (now with $X_i = 1$)

$$
P(\text{correct} \mid r_i, a, X_i) = \begin{cases}\Phi\left(\frac{C \cdot r_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - \Phi\left(\frac{C \cdot r_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases} \tag{3}
$$

One can show that the first term from equation (7a / 7b) is just the expectation of the argument to the cumulative normal distribution function in equation (3). This argument, from the derivation is the log-likelihood ratio, which can be interpreted as the average strength of evidence in favor of the true state. We might call this is I;

$$
I_i = \mathbb{E}\left(\frac{C \cdot r_i}{\sigma_e^2 + \sigma_m^2}\right) = \frac{(CX_iD_i - C\beta)}{\sigma_e^2 + \sigma_m^2}
$$

The latter term in equation (7a / 7b) adjusts this average evidence, because of the information contained within the action: 

$$
S_i = -\rho \cdot \frac{ \lambda\left(-\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)C}{\sqrt{\sigma_e^2 + \sigma_m^2} }
$$

This term serves the purpose of changing the mean expectation of the confidence to incorporate the information of the action. The addition of actions into the confidence computation have been proposed in the second order rater model from Fleming and Daw, where they also re-analyzed data to show this pattern empirically [@FLEMINGSelfEvaluationDecisionMakingGeneralBayesianFrameworkMetacognitive2017]. 

<!-- I might be wrong but without it being conditioned on the action you also have no error monitorin right so maybe nice to mention that --> 

To summarize, confidence computation can in this model be thought of as a weighting of the expected normalized evidence with a correction factor for having committed to a choice. Below we introduce a couple of additional parameters in order to encompass idiosyncrasies in the task and participants, as well as probabilistic relate the above formulation to the observed confidence ratings.


#### Additional parameters

We introduce 5 additional parameters. $\Theta_2 = (\kappa_c, \gamma, \mu_{\beta}, c_0,c_1)$.
<!-- maybe explain why we introduce them-->
<!-- Feel like it is "foreshadowed" in the just above no?-->

* $\gamma$ is the lapse rate on the binary choices this is a stimulus independent factor that shifts the asympote of the psychometric function away from certainty i.e. the asymptotic becomes $\gamma$ and $1-\gamma$. 

* $\mu_{\beta}$ is a meta cognitive bias parameter, serving as an independent vertical shift of confidence ratings.

* $\kappa_c$ is the precision of the ordered beta likelihood. 

* Lastly $c_0,c_1$ reflects the two cutpoints for the ordered beta likelihood. These two parameters help generalize the beta-distribution to outcomes that contain 0 and 1 ratings. 

See [@KUBINECOrderedBetaRegressionParsimoniousWellFittingModel2023] for the ordered beta likelihood. 

The complete set of parameters for the model are: 

$$
\begin{aligned}
\Theta_s &= \begin{bmatrix}
           \beta_s \\
           \sigma_{e_s} \\
           \sigma_{c_s} \\
           \sigma_{m_s} \\
           \mu_{\beta_s} \\
           \kappa_{c_s} \\
           \gamma_s 
         \end{bmatrix} 
         \sim 
         \mathcal{N}(\mu_{\Theta},\Sigma_{\Theta}) \\
\text{where } & \beta_s  \in [-1,1], \sigma_{e_s} >0, \sigma_{c_s} >0, \sigma_{m_s} > 0, \\
& \mu_{\beta_s} \in \mathbb{R}, \kappa_{c_s} > 0, \gamma_s \in [0,0.5] \\
& (c_{1_s} , c_{0_s}) \sim \delta([1,10,1])
\end{aligned}
$$
In the above formulation that $\mathcal{N}$ is a multivariate normal distribution with mean and covariance. Further all constraints on the parameters as written were accomplished by transformation, all positively constrained parameters were exponentiated, parameters with an upper and lower bound were transformed with the inverse logit function and multiplied appropriately. The constraint of the cut-points were applied by the prior of a Dirichlet distribution ($\delta()$). Note, the cutpoints are sampled on a subject-by-subject basis. 

The priors of the model are as follows:

$$
\begin{align*}
\mu_{\beta} &\sim \mathcal{N}(0, 0.5) & \text{(group mean of threshold)} \\
\mu_{\sigma_{e_s}} &\sim \mathcal{N}(-2, 2) & \text{(group mean of encoding uncertainty)} \\
\mu_{\sigma_{c_s}} &\sim \mathcal{N}(-2, 2) & \text{(group mean of choice variability)} \\
\mu_{\sigma_{m_s}} &\sim \mathcal{N}(-2, 2) & \text{(group mean of meta uncertainty)} \\
\mu_{\beta_c} &\sim \mathcal{N}(0, 0.5) & \text{(group mean of meta bias)} \\
\mu_{\gamma_s} &\sim \mathcal{N}(2, 2) & \text{(group mean of lapse rate)} \\
\mu_{\kappa_c} &\sim \mathcal{N}(-4, 2) & \text{(group mean of confidence precision)} \\
\end{align*}
$$


$$
\begin{align*}
\tau_{\beta} &\sim \mathcal{N}(1, 1) & \text{(Between subject SD of threshold)} \\
\tau_{\sigma_{e_s}} &\sim \mathcal{N}(1, 1) & \text{(Between subject SD of encoding uncertainty)} \\
\tau_{\sigma_{c_s}} &\sim \mathcal{N}(1, 1) & \text{(Between subject SD of choice variability)} \\
\tau_{\sigma_{m_s}} &\sim \mathcal{N}(1, 1) & \text{(Between subject SD of meta uncertainty)} \\
\tau_{\beta_c} &\sim \mathcal{N}(1, 1) & \text{(Between subject SD of meta bias)} \\
\tau_{\gamma_s} &\sim \mathcal{N}(1, 1) & \text{(Between subject SD of lapse rate)} \\
\tau_{\kappa_c} &\sim \mathcal{N}(1, 1) & \text{(Between subject SD of confidence precision)} \\
\end{align*}
$$

- $\mathbf{R} \sim \mathrm{LKJ}(\eta = 2)$ (correlation matrix for the multivariate normal distribution.)
- $z_{expo} \sim \mathcal{N}(0, 1)$ (participant deviations)
- $c_0, c_{1}$: see Dirichlet prior in Stan code

The full model becomes.

$$
\begin{aligned}
P(a = 1) &= \Phi\left(\frac{D_i \cdot X_i- \beta_s}{\sqrt{\sigma_{e_s}^2 + \sigma_{c_s}^2}}\right) \\
\mu_{c_i} &= \begin{cases}\Phi\left((I_i-S_i)\cdot C_i\right) & \text{if } a_i = 1 \\ 
1 - \Phi\left((I_i+S_i)\cdot C_i\right) & \text{if } a_i = -1\end{cases} \\
a_i &\sim \mathcal{bernoulli}(p(a_i = 1)) \\
Conf_i &\sim \mathcal{ord\beta}(\mu_{c_i},\kappa_{c_s},c_{0_s},c_{1_s})
\end{aligned}
$$

with

$$
I_i = \frac{(CX_iD_i - C\beta)}{\sigma_{e_s}^2 + \sigma_{m_s}^2}, \quad
S_i = -\rho_s \cdot \frac{ \lambda\left(-\frac{X_iD_i - \beta_s}{\sqrt{\sigma_{e_s}^2 + \sigma_{c_s}^2}}\right)C}{\sqrt{\sigma_{e_s}^2 + \sigma_{m_s}^2} }, \quad
K_i = \frac{1} {\sqrt{1 + \frac{C^2}{\sigma_{e_s}^2+\sigma_{m_s}^2}}}
$$


Next we show verification of the above formulated marginalized heuristic model, by simulating data from equation (2) (with $X_i = 1$) averaging across the signed stimulus axis and displaying for the same parameter estimates (i.e. $\Theta_1$) of the marginalized heuristic model.

In the first validation step we simulate from different combinations of noise parameters (facets).

```{r, error=F,message=F,warning=F, fig.height=10,fig.width=10}
source(here::here("Preregistration","scripts","Simulation.R"))

library(tidyverse)

trials = 500000

# Define high/low values
sigma_e_vals <- c(-2, -1.5)   # example low/high
sigma_k_vals <- c(-3, -2)
sigma_m_vals <- c(-1, 0)


param_grid <- expand.grid(
  sigma_e = sigma_e_vals,
  sigma_k = sigma_k_vals,
  sigma_m = sigma_m_vals,
  bias    = 0  # start with bias = 0 for all
)

# Add an extra row for the one case where bias != 0 (if desired)
param_grid <- param_grid%>% 
  mutate(id = row_number())

param_grid <- param_grid %>%
  mutate(facet_label = paste0("S_e=", sigma_e, 
                              ", S_k=", sigma_k, 
                              ", S_m=", sigma_m, 
                              ", B=", bias))

breaks <- seq(-1, 1, by = 0.05)
midpoints <- breaks[-length(breaks)] + 0.1

df_mech = param_grid %>% 
    rowwise() %>%
    mutate(sims = list(simulate_data_p_bias(500000, sigma_e = sigma_e, sigma_k = sigma_k, sigma_m = sigma_m, bias = bias, seed = 1))) %>% 
    ungroup() %>% 
    unnest() %>% 
    mutate(XD = X * D,
    bin = cut(XD, breaks = breaks, include.lowest = TRUE, labels = FALSE),
    XD = (breaks[bin] + breaks[bin + 1]) / 2) %>% 
  mutate(Action = as.factor(a_gauss),
         Accuracy = as.factor(ACC))  

df_heu = param_grid %>% 
    rowwise() %>%
    mutate(sims = list(simulate_data_margin_bias(500, sigma_e = sigma_e, sigma_k = sigma_k, sigma_m = sigma_m, bias = bias, seed = 1))) %>% 
    ungroup() %>% 
    unnest() %>% 
    mutate(XD = X*D) %>% 
  mutate(bin = NA)%>% 
  mutate(Action = as.factor(a_gauss),
         Accuracy = as.factor(ACC))
    
  

rbind(df_mech,df_heu) %>% ggplot()+
  geom_pointrange(data = df_mech %>% 
                    group_by(XD, Accuracy,facet_label) %>% 
                    summarize(mean = mean(c_mu), 
                              se = sd(c_mu)/sqrt(n())),
                  aes(x = XD,y = mean,ymin = mean-2*se, ymax = mean+2*se, fill = Accuracy), shape = 21, col = "black")+
  geom_pointrange(data = df_mech %>% 
                    group_by(XD, facet_label) %>% 
                    summarize(k = sum(as.numeric(as.character(Action))),
                              n = n(),
                              prop = k / n,
                              a_post = k + 1,
                              b_post = (n - k) + 1,
                              lower = qbeta(0.025, a_post, b_post),
                              upper = qbeta(0.975, a_post, b_post)),
                  aes(x = XD,y = prop,ymin = lower, ymax = upper), col = "black", shape = 21,fill = "black")+
  geom_line(data = df_heu, aes(x = XD,y = c_mu, col = Accuracy), linewidth = 2, alpha = 0.75)+
  geom_line(data = df_heu, aes(x = XD,y = p_gauss), linewidth = 2, alpha = 0.75, col = "black")+
  theme_classic()+
  facet_wrap(~facet_label, scales = "free") +
  labs(x = "Stimulus Intensity")+
  geom_vline(xintercept = 0, linetype = 2)+
  geom_hline(yintercept = 0, linetype = 2)+
  geom_hline(yintercept = 1, linetype = 2)+
  theme(legend.position = "top")
  
  
```


In the above plot we see the results from averaging `r trials` simulated trials from equation (2), for each bin of the signed stimulus intensity scale (dots are mean and approximate 95 Interval). The thick solid lines represents the marginalized heuristic model predictions. 

Next we simulate using fixed noise parameters and change the bias $\beta$. Note; that now we are displaying the action and not the accuracy of the choice.

```{r, error=F,message=F,warning=F, fig.height=10,fig.width=10}

# Define high/low values
sigma_e_vals <- c(-1.5)   # example low/high
sigma_k_vals <- c(-2)
sigma_m_vals <- c(-1.2)


param_grid <- expand.grid(
  sigma_e = sigma_e_vals,
  sigma_k = sigma_k_vals,
  sigma_m = sigma_m_vals,
  bias    = seq(-0.7,0.7,length.out = 9)
)

# Add an extra row for the one case where bias != 0 (if desired)
param_grid <- param_grid%>% 
  mutate(id = row_number())

param_grid <- param_grid %>%
  mutate(facet_label = paste0("S_e=", sigma_e, 
                              ", S_k=", sigma_k, 
                              ", S_m=", sigma_m, 
                              ", B=", bias))

breaks <- seq(-1, 1, by = 0.05)
midpoints <- breaks[-length(breaks)] + 0.1

df_mech = param_grid %>% 
    rowwise() %>%
    mutate(sims = list(simulate_data_p_bias(500000, sigma_e = sigma_e, sigma_k = sigma_k, sigma_m = sigma_m, bias = bias, seed = 1))) %>% 
    ungroup() %>% 
    unnest() %>% 
    mutate(XD = X * D,
    bin = cut(XD, breaks = breaks, include.lowest = TRUE, labels = FALSE),
    XD = (breaks[bin] + breaks[bin + 1]) / 2) %>% 
  mutate(Action = as.factor(a_gauss),
         Accuracy = as.factor(ACC))  

df_heu = param_grid %>% 
    rowwise() %>%
    mutate(sims = list(simulate_data_margin_bias(5000, sigma_e = sigma_e, sigma_k = sigma_k, sigma_m = sigma_m, bias = bias, seed = 1))) %>% 
    ungroup() %>% 
    unnest() %>% 
    mutate(XD = X*D) %>% 
  mutate(bin = NA)%>% 
  mutate(Action = as.factor(a_gauss),
         Accuracy = as.factor(ACC))
    
  

rbind(df_mech) %>% ggplot()+
  geom_pointrange(data = df_mech %>% 
                    group_by(XD, Action,facet_label) %>% 
                    summarize(mean = mean(c_mu), 
                              se = sd(c_mu)/sqrt(n())),
                  aes(x = XD,y = mean,ymin = mean-2*se, ymax = mean+2*se, fill = Action), shape = 21, col = "black")+
  geom_pointrange(data = df_mech %>% 
                    group_by(XD, facet_label) %>% 
                    summarize(k = sum(as.numeric(as.character(Action))),
                              n = n(),
                              prop = k / n,
                              a_post = k + 1,
                              b_post = (n - k) + 1,
                              lower = qbeta(0.025, a_post, b_post),
                              upper = qbeta(0.975, a_post, b_post)),
                  aes(x = XD,y = prop,ymin = lower, ymax = upper), shape = 21, col = "black",fill = "black")+
  geom_line(data = df_heu, aes(x = XD,y = c_mu, col = Action), linewidth = 2, alpha = 0.75)+
  geom_line(data = df_heu, aes(x = XD,y = p_gauss), linewidth = 2, alpha = 0.75, col = "black")+
  theme_classic()+
  facet_wrap(~facet_label, scales = "free") +
  geom_vline(xintercept = 0, linetype = 2)+
  geom_hline(yintercept = 0, linetype = 2)+
  geom_hline(yintercept = 1, linetype = 2)+
  labs(x = "Stimulus Intensity")+
  theme(legend.position = "top")
  

  
```

In the above plot we see the results from averaging `r trials` simulated trials from equation (2), for each bin of the signed stimulus intensity scale (dots are mean and approximate 95 Interval). The thick solid lines represents the marginalized heuristic model predictions. 


### Model for the current experiment.

In the current experiment we will present participants with a random dot kinematogram on each trials where subjects are asked to choose which direction the majority of dots are moving. After an experimental delay (post decisional response window) subjects are asked to give a confidence rating. We aim to test whether this post decisional response window changes the meta-cognitive noise and metacognitive bias parameter of the above model. This will be implemented in the following way:

$$
\sigma_{m_t}  = \sigma_{m_s} + \Delta{\sigma_{m_s}} \cdot PDW_t\\
\mu_{\beta_t} = \mu_{\beta_s} + \Delta{\mu_{\beta_s}} \cdot PDW_t
$$

With the subject level changes due to the post decisional window (PDW) being drawn from another group level parameter from the multivariate normal distribution highlighted above. These group-level parameters are then given the following priors:


$$
\begin{align*}
\mu_{\Delta{\sigma_{m_s}}} &\sim \mathcal{N}(0, 0.5) & \text{(group mean of effect of PDW on meta uncertainty)} \\
\mu_{\Delta{\mu_{\beta_c}}} &\sim \mathcal{N}(0, 0.5)  & \text{(group mean of effect of PDW on meta bias)} \\
\tau_{\Delta{\sigma_{m_s}}} &\sim \mathcal{N}(0, 0.5) & \text{(Between subject SD of effect of PDW on meta uncertainty)} \\
\tau_{\Delta{\mu_{\beta_c}}} &\sim \mathcal{N}(0, 0.5)  & \text{(Between subject SD of effect of PDW on meta bias)} \\

\end{align*}
$$

- with PDW being the Post Decisional response window.
- ${\Delta{\sigma_{m_s}}}$ and $\Delta{\mu_{\beta_c}}$ being the change in meta-cognitive uncertainty and meta cognitive bias based on PDW respectively.

Next, we investigate how this model with meta cognitive uncertainty and meta cognitive bias being able to vary by PDW fit to our pilot data. The first step is to showcase model comparison on approximate leave one trial out cross validation for the models assuming that the rater has access to the objective stimulus intensity and the reduced model where we assume that the this stimulus intensity is 1 i.e. $X_i = 1$


### Model comparison

```{r, echo = FALSE, message=F, error=F, warning=F}
loocomparison <- readRDS(here::here("Preregistration","Results","loocomparison.rds"))

# model_with_x <- readRDS(here::here("Experimental analysis","Heurestic models",
#                                    "saved models","saved models hier","fit_finalheu_lapse_cypertron.rds"))
# 
# 
# model_with_x1 <- readRDS(here::here("Experimental analysis","Heurestic models",
#                                     "saved models","saved models hier","fit_finalheu_lapse_x1.rds"))
# 
# loocomparison = loo::loo_compare(list(with_x = model_with_x$loo(), with_x1 = model_with_x1$loo()))
# 
# 
# saveRDS(loocomparison,here::here("Preregistration","Results","loocomparison.rds"))
```

As a sanity check on the assumption of $X_i = 1$, we fitted both hierarchical models to our pilot data and computed approximate leave one out cross-validation, to test if the assumption could be justified. This comparison showed that the two models were virtually indistinguishable with the difference in expected log probability density being `r round(loocomparison[2,1],2)` $\pm$ `r round(loocomparison[2,2],2)`, where generally it is thought that a difference below -4 and or more than two standard errors can be considered as evidence for one model over the other (REF), thus at least for our pilot data for predictive performance there seems to be no difference. Digging further into the distinction between these two models we below plot all pairwise parameters together with the identity line.

```{r,fig.height=6, fig.width=8, message=F,error=F,warning=F}
# parameters = c("beta","sigma_e","sigma_k","sigma_m","confprec","lapse","meta_bias","confprec","meta_bias_beta","sigmam_beta")
# 
# parameter_comparison_plot = rbind(
#     model_with_x$summary(parameters) %>% 
#       mutate(
#         subject  = as.numeric(str_extract(variable, "(?<=\\[)\\d+(?=\\])")),
#         variable = str_remove(variable, "\\[\\d+\\]"),
#         model    = "X"
#       ),
#     
#     model_with_x1$summary(parameters) %>% 
#         mutate(
#         subject  = as.numeric(str_extract(variable, "(?<=\\[)\\d+(?=\\])")),
#         variable = str_remove(variable, "\\[\\d+\\]"),
#         model    = "Xone"
#       )
#     )%>% select(variable,model,mean,q5,q95) %>% 
#       pivot_wider(names_from = "model",values_from = c("mean","q5","q95")) %>% unnest() %>% 
#   ggplot(aes(x = mean_X, y = mean_Xone))+
#   geom_pointrange(aes(ymin = q5_Xone, ymax = q95_Xone))+
#   geom_pointrange(aes(xmin = q5_X, xmax = q95_X))+
#   facet_wrap(~variable, scales = "free")+
#   labs(y = "Model with X = 1", x = "Model with X")+
#   geom_abline()+
#   theme_classic()
# 
# 
# ggsave(here::here("Preregistration","Figures","parameter_comparison_plot.PNG"), height = 6, width = 8, dpi = 300)

magick::image_read(here::here("Preregistration","Figures","parameter_comparison_plot.PNG"))

```

Generally the two models align well in their subject level parameter estimates, further justifying this simplification which makes theoretical sense.

### Posterior predictive checks

Next we show predictive checks of the Hierarchical heuristic model for our pilot participants.

```{r,fig.height=12, fig.width=18, message=F,error=F,warning=F}
magick::image_read(here::here("Preregistration","Figures","Pilot_predictive_checks.PNG"))
```


### Hierarchical parameter recovery

To further assess the robustness of our model implementation and the planned sequential‑sampling procedure, we simulated new subjects from the posterior distribution of the hierarchical model. We did this by extracting a draw from the posterior and then simulating 25 subjects from this draw. We then fitted the model to increasing steps (i.e. 5,7,9, 12,15,20,25) of subjects, to demonstrate that the sequential sampling approach would be feasible. This also allowed us to do parameter recovery on the hierarchical model without additional computational resources. Below we showcase that the model can to an adequate degree recover its parameters on a subject and group level for parameters of interest.

```{r,fig.height=7, fig.width=14, message=F,error=F,warning=F}
magick::image_read(here::here("Preregistration","Figures","parameterrecovery_hier.PNG"))
```


### Sequential sampling

#### Priors

As our sequential sampling approach uses Bayes Factors for the termination of data collection, the prior distribution of this parameter is highly influential. To maximize transparency we visualize the curves that we allow given the prior below, interested readers are also referred to  [Sequential Sampling](../Sequential sampling/) directory, here we provide scripts for fitting visualizing and diagnosing our model.

```{r, fig.height=8, fig.width=12, warning=F,error=F,message=F}
source(here::here("Simulations","Heurestic","scripts","Simulate_mcmc.R"))

n_sims <- 50

df_list <- lapply(1:n_sims, function(i){
  df <- sim_trials(300, seed = i)
  df$sim_id <- i
  df
})

df_all <- do.call(rbind, df_list)

df_all %>% pivot_longer(cols = c("theta","c_mu")) %>% 
  mutate(ACC = as.factor(ifelse(name == "theta","Response",ACC))) %>% 
  group_by(ACC,name,X,sim_id) %>% 
  summarize(mean = mean(value),
            se = sd(value) / sqrt(n())) %>% 
  ggplot(aes(x = X, y = mean, ymin = mean-2*se, ymax = mean+2*se, col = ACC, group = interaction(ACC,sim_id)))+
  # geom_pointrange()+
  facet_wrap(~name,ncol = 1)+
  scale_color_manual(values = c("darkred","darkgreen","black"))+
  geom_smooth(se = F, alpha = 0.25)+
  theme_minimal()
```

The plot above shows the prior predictive checks of the group level parameters (i.e. these curves represent group level curves). As can be seen the curves generally span everything. Now in order to test whether our sequential sampling approach works, we simulate data from draws of the posterior distribution of our pilot data, we then investigate how the Bayes Factors of both hypotheses outlined above evolve over the number of subjects collected. Instead of selecting the posterior draws of our parameters of interest i.e. $\mu_{\Delta{\sigma_{m_s}}}$ and $\mu_{\Delta{\mu_{\beta_c}}}$ we specifically fixed these to either -0.05, 0 or 0.05 to test that we can detect effects of this magnitude in both direction, but also terminate sampling if the effect is zero. Below we display the size of these effects (simulated data from the mean of the group means).

```{r, fig.height=8, fig.width=12, warning=F,error=F,message=F}
effects <- c(-0.05, 0, 0.05)

# Create all combinations (if you had two effects)
grid <- expand.grid(effect1 = effects, effect2 = effects)

df_list <- vector("list", nrow(grid))

for(i in 1:nrow(grid)){
  df_list[[i]] <- sim_trials_fix(300, 1997, grid$effect1[i], grid$effect2[i])
  df_list[[i]]$sigma_m_beta <- grid$effect1[i]
  df_list[[i]]$meta_bias_beta <- grid$effect2[i]
}

df_effects <- do.call(rbind, df_list)


df_effects %>% 
  # filter(ACC == 1) %>% 
  mutate(Accuracy = ACC) %>% 
  mutate(PDW = as.factor(cut(interval,3))) %>% 
  group_by(X,PDW,Accuracy,meta_bias_beta,sigma_m_beta) %>% 
  summarize(mean = mean(c_mu),
            se = sd(c_mu) / sqrt(n())) %>% 
  ggplot()+
  geom_hline(yintercept = 0.5, linetype = 2)+
  scale_color_manual(values = c("orange","orange3","darkred","green","green3","darkgreen"))+
  geom_pointrange(aes(x = X, y = mean, ymin = mean-2*se, ymax = mean+2*se, col = interaction(PDW,Accuracy)))+
  geom_smooth(aes(x = X, y = mean, col = interaction(PDW,Accuracy)),se = F)+
  facet_grid(sigma_m_beta~meta_bias_beta, labeller = label_both)+
  theme_minimal()


```


As mentioned above we simulated subjects from draws of the posterior that was informed by our pilot data. We then fitted the model to increasing steps (i.e. 5,7,9, 12,15,20,25) of subjects, to demonstrate that the sequential sampling approach would be feasible. below we showcase 5 such trajectories of Bayes factors across the number of subjects used, together with another panel showing how the posterior distribution of the parameter of interest changes with increasing subjects.


```{r, fig.height=8, fig.width=12, warning=F,error=F,message=F}
magick::image_read(here::here("Preregistration","Figures","sequential_samping_plot.PNG"))
```


In the above plot we show the development of the effect of the interval between choice and confidence rating on both meta cognitive bias (green) and meta cognitive uncertainty (orange). Circles display a fitted model without divergences, above 400 effective sample size for both the bulk and the tails and rhat values below 1.03. The Left side displays the posterior distribution of each parameter for increasing subjects. The right side displays the development of the Bayes-factors for increasing subjects for the same draw of the posterior distribution fitted to our pilot data.


From the above plot it can be seen that even with this model we might run into convergences issues, below we describe our approach for determining and dealing with convergences issues. 

Firstly, we will collect data from 15 subjects before we fit our hierarchical model the first time and test the size of the bayes-fator. We do this as hierarchical models generally need some initial data to be able to converge (see also plot above for simulation). We will then after each additional 5 subjects check the infered bayes factor as this coincides with the expected number of subjects being collected each day.


