---
title: "Current model"
author: Jesper Fischer Ehmsen^1^,
        Siebe Everaets^xx^,
        Micah G. Allen^1,2,\*^,
        Francesca Fardo^1,3,\*^
output:
  html_document: default
  word_document:
    # reference_docx: Knitting files/docx_template.docx
  pdf_document:
    latex_engine: xelatex
# bibliography: 
  # - "Knitting files/References.bib"
link-citations: yes
linkcolor: blue
# csl: Knitting files/science-advances.csl
always_allow_html: yes
---

::: {custom-style="Title text"}
^1^Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University

^2^Cambridge Psychiatry, University of Cambridge, United Kingdom

^3^Danish Pain Research Center, Department of Clinical Medicine, Aarhus University

^\*^Equal contributions

Corresponding author: Jesper Fischer Ehmsen (jesperfischer@cfin.au.dk)
:::

```{r load packages, warning = F, message = F, echo = F, include=FALSE}
knitr::opts_chunk$set(dpi=300)

# seed
set.seed(123)

# packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load("here", "knitr","yaml")

knitr::opts_chunk$set(echo = TRUE)

required_packages = c("cowplot", "magick", "tidyverse", "flextable","truncnorm")
lapply(required_packages, library, character.only = TRUE)
```


### Static models of metacognition

Here we outline the general idea of static models of perceptual decision making and confidence judgements.

In SDT, an actor makes a binary choice on each trial i, $a_i \in \{-1,+1\}$ based on some internal evidence $e_i$. The action reflects the internal evidence on the true state the actor is trying to infer $D_i \in \{-1,+1\}$. 

In SDT, evidence is drawn from a normal distribution, as perception is a noisy process, with a mean in the direction of the true state scaled by the strength of the signal $X_i$.

$$
e_i \sim \mathcal{N}(X_iD_i,\sigma_e)
$$

Here inorder to encompass **3** from the Patterns of metacognition section we introduce a parameter that can shift this evidence distribution, essentially introducing a bias in the encoding of evidence.

$$
e_i \sim \mathcal{N}(X_iD_i - \beta,\sigma_e) 
$$

To produce an action the actor compares the internal evidence to criterion $c$:

$$
a_i = \begin{cases}+1 & \text{if } e_i > c \\ -1 & \text{if } e_i \le c\end{cases}
$$

This criterion is usually thought to be static within an experimental setting for each agent.

#### Meta-cognition

After having made a choice, the agent might be asked to give a confidence judgement in their choice. A particularly prominent way to conceptualize this is to think that the agent computes the probability that the action that they produced was correct. Below we write one promient idea of how such confidence ratings might be computed that allows for error-monitoring [@FLEMINGSelfEvaluationDecisionMakingGeneralBayesianFrameworkMetacognitive2017].


In this formulation confidence judgement are based on draws from a multivariate normal, which contains the evidence used for the action $e_i$ and partically independent information only accesiable to the rater $r_i$, these quantities are assumed to be correlated with $\rho$ governing information sharing. I.e. if $\rho$ = 1, the confidence judgement is based on the same information as the action, however if $\rho$ = 0 the confidence judgement is based on completely independent information. Note we here write this model as originally proposed by Fleming and Daw without the evidence bias

$$
\begin{bmatrix}e_i \\ r_i\end{bmatrix}\sim \mathcal{N}\left(\begin{bmatrix}X_iD_i \\ X_iD_i\end{bmatrix},\begin{bmatrix}\sigma_e^2 & \rho \cdot \sigma_e \cdot \sigma_m \\ \rho \cdot \sigma_e \cdot \sigma_m & \sigma_m^2\end{bmatrix}\right)
$$

The rater is then thought to use this information to calculate the subjective probability of being correct, to make a confidence judgement.

One of the main arguments for such model and other models with independent information was to explain error-monitoring i.e. agents' ability to know that they produced an error and give low (even below guessing) confidence ratings, see [Design considerations](Models and Data in Perceptual Decision-Making with Confidence.Rmd), for an in depth consideration of this account and how it relates to dynamic models of meta-cognition and espcially post-decisional evidence accumulation.



## Evidence copy (Choice Variability) model of metacogntion (EA/CV - Meta)

Below we formalize the idea that independent information is not necessary to explain error-monitoring, if instead it is assumed that agents have variability in their choices, given the evidence.

This can be formalized as instead of a deterministic decision policy (which is usually assumed in SDT), agents have a noisy decision process that can produce errors even though the evidence is in favor of a different choice.

We therefore extend the standard SDT framework by introducing stochastic choice policy by assuming that the criterion varies from trial-by-trial: $c_i \sim \mathcal{N}(0,\sigma_c)$. This yielding:

$$
a_i = \begin{cases}+1 & \text{if } e_i > c_i \\ -1 & \text{if } e_i \le c_i\end{cases}
$$

Due to the fact that both $e_i$ and $c_i$ are normally distributed, it is possible to marginalize their difference which gives the probability of the action:

$$
P(a = 1 \mid D_i, X_i) = \Phi\left(\frac{D_i \cdot X_i- \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right) \tag{1}
$$

From choices alone, only $\sigma_e^2 + \sigma_c^2$ is identifiable, as both noise sources equivalently flatten the psychometric function. However, by including confidence ratings these can be disentangled.

For confidence ratings, we define:

$$
\text{confidence} = P(\text{correct} \mid r, a) = P(D = a \mid r)
$$
Here we assume that the rater has access to a noisy evidence: $r_i \sim \mathcal{N}(e_i, \sigma_m)$, where $\sigma_m$ is meta cognitive noise. 

This can be shown to yield:
See the full derivation in [Derivations](Derivations/inverselogit.Rmd).

$$
P(D=1 \mid r_i, X_i) = g\left(\frac{2 \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right)
$$

Assuming a uniform prior on D, i.e. $P(D_i = 1) = P(D_i = -1) = 0.5$

Which gives the following equation for the trial by trial confidence ratings.

$$
P(\text{correct} \mid r_i, a, X_i) = \begin{cases}g\left(\frac{2 \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - g\left(\frac{2 \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases} \tag{2}
$$

Here the function $g(.)$ is the inverse logit function i.e. $g(x)= \frac{1}{1+exp(-x)}$.

From confidence alone, only $\sigma_e^2 + \sigma_m^2$ is identifiable. However, combining choices and confidence allows estimating all three parameters: $\sigma_e$, $\sigma_c$, and $\sigma_m$. 

As can be seen from equation (2), to compute the probability of being correct the agent (rater) needs access to the stimulus intensity ($X_i$), even though assumed in some models [@FLEMINGNeuralMediatorsChangesMindPerceptualDecisions2018a], we believe that the agent cannot have access to the objective stimulus intensity for several reasons. We will get back to these reasons and a final formulation without this term later. Interested readers are refereed to [@RAUSCHFoldedXpatternNotNecessarilyStatisticalSignature2019], were simulations are conducted with differing generative processes for the stimulus intensity.



Before demonstrating simulations from this model to provide an intuition for the meaning of the parameters and their interactions we demonstrate how this kind of model can be argued for, consider the below plate notation. 


```{r,fig.height=8, fig.width=12, message=F,error=F,warning=F}

tinytex::latexmk(here::here("Preregistration","Figure scripts","Platenotations", "platenotation.tex"))

img1 <- image_read_pdf(here::here("Preregistration", "Figure scripts","Platenotations", "platenotation.pdf"),
                       density = 600)


print(img1)
```


An agent is presented with a stimulus that has a direction D (the true state of the world) and an intensity (X). This creates an internal evidence representation $e_i$. The brain then makes a copy of this evidence (an efference copy) (red split path). The internal evidence representation is then sent to the premotor and later motor cortex where an efferent signal is produced to generate an action, crucially this additional processing introduces noise 




## Simulations

We demonstrate that this model reproduces key behavioral patterns (folded-X, double-increase) under different parameter combinations. Simulations show that the ratio $\frac{\sigma_e}{\sigma_c}$ determines these patterns: 

A low ratio (high $\sigma_c$ low $\sigma_e$) produces folded-X (strong error-monitoring), while a high ratio (low $\sigma_c$ high $\sigma_e$) produces double-increase (weak error-monitoring).

Metacognitive noise $\sigma_m$ flattens confidence curves regardless of accuracy.

Note that for these simulations we assume that the bias in evidence encoding $\beta$ is 0 as this just shifts the psychometric function as well as the confidence curves horizontally

```{r}
set.seed(199742)

n_trials <- 500

simulate_interaction <- function(sigma_m, sigma_c,sigma_e) {
  X <- abs(rnorm(n_trials,2,1))
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_c)
  p_gauss <- pnorm(e / sigma_c)
  a <- rbinom(n_trials, 1, p_gauss)

  data.frame(e = e,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}


df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_c = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_c,sigma_e))) %>% 
  unnest(data)

```

Now lets turn to how evidence, the physical stimulus and accuracy interacts across parameters:

```{r, fig.height=6, fig.width=10,warning=F,message=F}

ACC = df_interaction %>%
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = X*D, y = e, col = Accuracy)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point()+
  facet_grid(sigma_e~sigma_c, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))

ACC
```

From the above plot its obvious that $\sigma_e$ scales the the range of evidence observed (see y-axis range for the different rows). When $\sigma_c$ is low (left column) choices are consistent (when the the evidence is positive and the signed stimulus is positive i.e. D = 1 then a correct response is given). When $\sigma_c$ increases this relationship essentially becomes more noisy with correct and incorrect choices being less distinguishable across the evidence = 0 line. Another way to display this is by plotting the actions instead of the accuracy: 

here the space is only split in two quadrants instead of 4.

```{r, fig.height=6, fig.width=10,warning=F,message=F}
Act = df_interaction %>%
  mutate(Action = as.factor(a)) %>% 
  ggplot(aes(x = X*D, y = e, col = Action)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point()+
  facet_grid(sigma_e~sigma_c, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))
Act
```


Next we look at the realized confidence ratings as a function of the three different sigmas:

```{r}
set.seed(199742)

n_trials <- 5000

conf_prec <- 100
c0 <- -3
c1 <- 3

simulate_interaction <- function(sigma_m, sigma_c,sigma_e) {
  X <- sample(c(0.01,1,2,5), n_trials, replace = TRUE)
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_c)
  p_gauss <- pnorm(e / sigma_c)
  a <- rbinom(n_trials, 1, p_gauss)
  C_mu <- ifelse(a == 1, 
                 plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)),
                 1 - plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)))
  

  c_mu <- pmax(0.001, pmin(0.999, C_mu))
  C <- ordbetareg::rordbeta(n_trials, c_mu, conf_prec, c(c0, c1))
  
  data.frame(e = e,
             C = C,
             C_mu = C_mu,
             # C_mu2 = C_mu2,
             ehat = ehat,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}

df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_c = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_c,sigma_e))) %>% 
  unnest(data)
```



First we can look to fix $\sigma_c = 1$ and then we look at how $\sigma_e$ & $\sigma_m$ change confidence ratings as a function of the Stimulus:

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_c == 1) %>% 
  mutate(SS = D*X) %>%
  group_by(SS,ACC, sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

From the above plot we see that when we plot confidence on the Signed Stimulus scale we end up with a somewhat double increase pattern, when the stimulus encoding uncertainty is high (i.e. 3 or 5). Furthermore we see a (")half) folded X when stimulus encoding uncertainty  is low. Generally we see that meta uncertainty makes the confidence ratings more flat lined at 0.5 regardless of their correctness.  

Now we fix $\sigma_e = 1$ and then we look at $\sigma_c$ & $\sigma_m$:

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_e == 1) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_c,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_c~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

This plot generally shows the same as the one with $\sigma_c = 1$ but opposite in pattern. I.e. high $\sigma_c$ here is more of a folded X with lower values (espeically below 1) make a double increase.

lastly we might plot $sigma_m = 1$

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_m == 1) %>%
  mutate(sigma = sigma_e/sigma_c) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_c,sigma,sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>%
  mutate(sigma_m = as.factor(sigma_m)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~sigma_c, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We see that the diagonal where $\sigma_e = \sigma_c$ that there generally is a half-folded X, with incorrect trials being flat with stimulus strength with correct trials increasing in confidence with stimulus strength. Going to the upper triangle we move into the space where $\sigma_c > \sigma_e$ where we see folded-X patterns of varying "strengths". Moving to the lower triangle we see the opposite pattern with double increase patterns emerging again with varying "strengths". 

This means, that from the models perspective, double increase and Folded-X patterns arise from the ratio of uncertainty in the encoding of the stimulus and the choice variability. When choice variability is high compared to stimulus encoding a folded X pattern arises whereas choice variability is low compared to stimulus encoding a double increase pattern arises. $\sigma_m$, meta uncertainty is agnostic to the accuracy of the rater and when it increases it makes the curves of the confidence ratings more flat regardless of correctness.

## Fitting

We fit this model using full Bayesian inference with Stan's NUTS sampler, treating latent trial-by-trial evidence as parameters estimated jointly with the noise parameters. Generally, from investigating our pilot data (see below) we saw that on a single-subject fitting this model is possible to fit with sufficiently good sampler diagnostics, however trying to fit this model in a hierarchical framework i.e. estimating group level parameters which subjects level parameters are drawn from we ran into sampling issues. 

We therefore build a heuristic model that approximates the above model, without the need for the trial-by-trial evidence. The main reason for this is to leverage hierarchical bayesian inference as our main hypothesis is on a particular group level parameter of this model while we additionally will be conducting sequential sampling (see Sequential sampling section).

## The Heuristic model

While the above model is powerful, its computational complexity (especially in hierarchical settings) motivate a simpler heuristic approximation. The key challenge is that confidence requires trial-specific evidence $r_i$, which cannot be marginalized analytically, because the random variable is inside the inverse logit function. 

The inverse logit function and the cummulative normal distribution are in practice close to being indistinguishable up to a constant i.e. $g(.) \sim \Phi(\frac{1}{1.7})$, see plot below:

```{r, fig.height=6, fig.width=10,warning=F,message=F}
# Compare inverse logit with cumulative normal approximation
data.frame(
  x = seq(-6, 6, by = 0.01)
) %>% 
  mutate(
  `Inverse Logit g(x)` = plogis(x),
  `Scaled CDF Φ(x/1.7)` = pnorm(x / 1.7)) %>% 
  pivot_longer(cols = -x, names_to = "Function", values_to = "Value") %>% 
  ggplot(aes(x = x, y = Value, color = Function, linetype = Function)) +
  geom_line(size = 1) +
  labs(x = "x", y = "Probability",
       title = "Comparison of Inverse Logit and Scaled Cumulative Normal") +
  theme_minimal() +
  theme(text = element_text(size = 12), legend.position = "bottom")
```

By using this approximation it is possible to marginalize out the trial-by-trial evidence and thereby getting a computationally easier model to fit (and an equation for the expectation of confidence) with the same parameters. So with this approximation we go from equation (2) to the following approximation.

$$
P(\text{correct} \mid r_i, a, X_i) = \begin{cases}\Phi\left(\frac{C \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - \Phi\left(\frac{C \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases} \tag{3}
$$
with $C = \frac{2}{1.7}$. 

Focusing on the instance where the choice was one.

$$
\Phi\left(\frac{C \cdot r_i \cdot X_i}{\sigma_e^2 + \sigma_m^2}\right)
$$

We want to marginalize over $r_i$ i.e. calculating the expectation of above equation across $r_i$. 

This is can be done using the general formula for the expectation of a random continuous variable

$$
\mathbb{E(X)} = \int_{-\infty}^\infty x \ p(x) dx
$$

substituting the variables with the following definition $b = \frac{CX_i}{\sigma_e^2 + \sigma_m^2}$

$$
\int_{-\infty}^\infty \Phi(b \cdot r_i) \ p(r_i) dr_i
$$

This integral can be shown to have the solution:

https://www.tandfonline.com/doi/abs/10.1080/03610918008812164

$$
\Phi \left(\frac{\mu \cdot b}{\sqrt{1+\sigma^2 \cdot b^2}}\right)
$$


with $r_i \sim \mathcal{N}(\mu,\sigma)$

given that we have  $r_i \sim \mathcal{N}(X_iD_i - \beta, \sqrt{\sigma_m^2+\sigma_e^2})$ we get: 

$$
\Phi \left(\frac{\mathop{\mathbb{E}}(r_i) \cdot b}{\sqrt{1+(\sigma_e^2+\sigma_m^2) \cdot b^2}}\right)
$$

Here we write the expectation of $r_i$ instead of the mean of the $r_i$ distribution ($X_iD_i - \beta$), as this will be conditional on the choice. 

Writing out the definition of b and simplify we get:


$$
\Phi\left(\frac{\mathop{\mathbb{E}}(r_i) \cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)
$$

This equation thus gives us the probability of being correct, when the choice is 1, meaning that we need to get the expectation of meta cognitive evidence $r_i$ conditional on the fact that the choice was 1. 

Equation (3) thus now becomes

$$
P(\text{correct} \mid a, X_i) 

\sim 

\begin{cases}\Phi\left(\frac{\mathop{\mathbb{E}}(r_i \mid a = 1) \cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)

& 

\text{if } a = 1 \\ 

1 - \Phi\left(\frac{\mathop{\mathbb{E}}(r_i \mid a = -1) \cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)

& 

\text{if } a = -1\end{cases} \tag{4}
$$

Lastly, we need to calculate is the two expectations of $r_i$ given the choice i.e. $\mathop{\mathbb{E}}(r_i \mid a_i = 1)$ and $\mathop{\mathbb{E}}(r_i \mid a_i = -1)$

#### Calculating the conditional expectations

To calculate these expectations, we need to account for the fact that the choice $a_i$ depends on the relationship between the evidence $e_i$ and the criterion $c_i$. and given that $r_i$ is a noisy readout of $e_i$ there is a dependency structure.

We have the following relationships:

- Evidence: $e_i \sim \mathcal{N}(X_iD_i - \beta, \sigma_e)$
- Criterion: $c_i \sim \mathcal{N}(0, \sigma_c)$
- Metacognitive evidence: $r_i = e_i + \epsilon_m$ where $\epsilon_m \sim \mathcal{N}(0, \sigma_m)$
- Choice: $a_i = 1$ if $e_i > c_i$, otherwise $a_i = -1$


we can treat the difference distribution $e_i - c_i$ as a joint bivariate normal distribution with $r_i$ and calculate the expectation of $r_i$ given that the $e_i - c_i$ random variable is above 0 (i.e. the choice was 1).

The difference $e_i - c_i$ follows a Gaussian distribution, as both random variables are Gaussians.

$$
e_i - c_i \sim \mathcal{N}(X_iD_i - \beta, \sqrt{\sigma_e^2 + \sigma_c^2})
$$
The joint bivariate normal distribution becomes:

$$
\begin{bmatrix}
e_i - c_i \\
r_i
\end{bmatrix}
\sim \mathcal{N}\left(
\begin{bmatrix}
X_iD_i - \beta \\
X_iD_i - \beta
\end{bmatrix},
\begin{bmatrix}
\sigma_e^2 + \sigma_c^2 & \sigma_e^2 \\
\sigma_e^2 & \sigma_e^2 + \sigma_m^2
\end{bmatrix}
\right)
$$

intuitively this makes sense as both the $e_i - c_i$ and $r_i$ distribution share $e_i$, with the same mean but are otherwise independent.


The key mathematical insight is that the covariance between $e_i-c_i$ and $r_i$ is $\sigma_e^2$, not zero. 

writing the full covariance and expanding it using this we get:  (https://www.probabilitycourse.com/chapter5/5_3_1_covariance_correlation.php#:~:text=The%20covariance%20between%20two%20random%20variables%20X,and%20X%20and%20Y%20are%20positively%20correlated.)


$$
\text{Cov}(e_i-c_i, r_i) = \text{Cov}(e_i-c_i, e_i + \epsilon_m)
$$
$$
\text{Cov}(e_i-c_i, r_i) = \text{Cov}(e_i,e_i) + \text{Cov}(e_i, \epsilon_m) - \text{Cov}(c_i,e_i) - \text{Cov}(c_i, \epsilon_m)=  \text{Cov}(e_i,e_i) = \text{Var}(e) =\sigma_e^2
$$
As $e_i$ and $\epsilon_m$ and also $c_i$ are independent, these covariance are 0 and we get:

$$
\text{Cov}(e_i-c_i, r_i) = \text{Cov}(e_i,e_i) = \text{Var}(e) =\sigma_e^2
$$
This shared variance structure means that $e_i-c_i$ and $r_i$ are correlated with correlation coefficient: See definition of the correlation coefficient below.

$$
\rho = \frac{\sigma_e^2}{\sqrt{(\sigma_e^2 + \sigma_c^2)} \cdot \sqrt{(\sigma_e^2 + \sigma_m^2)}} \tag{5}
$$


## Truncated normal distributions

To proceed, we need a general result from probability theory. Suppose we have two jointly normal random variables:

$$
\begin{pmatrix} Z \\ Y \end{pmatrix} \sim \mathcal{N}\left(\begin{pmatrix} \mu_Z \\ \mu_Y \end{pmatrix}, \begin{pmatrix} \sigma_Z^2 & \rho\sigma_Y\sigma_Z \\ \rho\sigma_Y\sigma_Z & \sigma_Y^2 \end{pmatrix}\right)
$$
It can be shown that the expectation of Y given that Z>0 is:
See the full derivation in [Derivation](Derivations/Truncated Normal.Rmd).

$$
\mathbb{E}(Y \mid Z > 0) = \mu_Y + \rho\cdot \sigma_Y\frac{\phi(z_z)}{\Phi(z_z)}
$$

With $z_Z = \frac{\mu_Z}{\sigma_Z}$


Now replacing Y with $r_i$ and Z with $e_i - c_i$ we get:

$$
\mathbb{E}(r_i \mid (e_i-c_i) > 0) =  (X_iD_i - \beta) + \rho \cdot \sqrt{\sigma_e^2 + \sigma_m^2} \cdot \frac{\phi(z_z)}{\Phi(z_z)}
$$

With $z_z = \frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}$

One might here use the correlation coefficient we found from equation (5) can be inserted see [Derivations](Derivations/no_correlation.Rmd).

It can then easily be shown using the same steps that when the action is -1 i.e. $(e_i-c_i) < 0$ this becomes:

$$
\mathbb{E}(r_i \mid (e_i-c_i) > 0) =  (X_iD_i - \beta) - \rho \cdot \sqrt{\sigma_e^2 + \sigma_m^2} \cdot \frac{\phi(-z_z)}{\Phi(-z_z)}
$$

we might define $\lambda(x) = \frac{\phi(x)}{\Phi(x)}$ which is the inverse Mills ratio giving us:

$$
\mathop{\mathbb{E}}(r_i \mid a = 1) = (X_iD_i - \beta) + \rho \cdot \sqrt{\sigma_e^2 + \sigma_m^2} \cdot \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$

For $a = -1$ (when $e_i - c_i \leq 0$):

$$
\mathop{\mathbb{E}}(r_i \mid a = -1) = (X_iD_i - \beta) - \rho \cdot \sqrt{\sigma_e^2 + \sigma_m^2} \cdot \lambda\left(-\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$
Note that $\lambda(-x) = \frac{\phi(x)}{1-\Phi(x)}$ by properties of the normal distribution.

### Final confidence model

Substituting these conditional expectations back into our marginalized confidence equation (4):

$$
P(\text{correct} \mid a = 1, X_i, D_i) = \Phi\left(\frac{\left[(X_iD_i - \beta) + \rho \cdot \sqrt{\sigma_e^2 + \sigma_m^2} 

\cdot \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)\right]


\cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right) 
$$

and simplifying

$$
P(\text{correct} \mid a = 1, X_i, D_i) = 

\Phi\left(
\left[

\frac{(CX_i^2D_i - C\beta X_i)}{\sigma_e^2 + \sigma_m^2} +

\rho \cdot \frac{ \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)CX_i}{\sqrt{\sigma_e^2 + \sigma_m^2} } 

\right]

\cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right) \tag{6a}
$$


which for the other choice becomes:


$$
P(\text{correct} \mid a = -1, X_i, D_i) = 

\Phi\left(
\left[
\frac{(CX_i^2D_i - C\beta X_i)}{\sigma_e^2 + \sigma_m^2} -

\rho \cdot \frac{ \lambda\left(-\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)CX_i}{\sqrt{\sigma_e^2 + \sigma_m^2} } 

\right]

\cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right) \tag{6b}
$$


This gives us a fully marginalized model for confidence ratings that depends only on the parameters $\Theta_1 = (\beta, \sigma_e, \sigma_c, \sigma_m)$ and the observed data (choices $a_i$, stimuli $X_i$, true states $D_i$), without requiring trial-by-trial estimation of latent evidence variables $e_i$ and $r_i$.

Before interpreting on the parts of the terms we need to revert back to one of the key assumptions of this model as was outline just after introducing equation (2). When the subject calculates the subjective probability of being correct it is assumed that the subject has access to the stimulus intensity. Below we have highlighted which of the stimulus intensities this is related to:

$$
P(\text{correct} \mid a = 1, X_i, D_i) = \Phi\left(\frac{\left[(X_iD_i - \beta) + \rho \cdot \sqrt{\sigma_e^2 + \sigma_m^2} 

\cdot \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)\right]

\cdot C {\color{red}{X_i}}}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2{\color{red}{X^2_i}}}{\sigma_e^2+\sigma_m^2}}}\right) 
$$

Interestingly these two stimulus intensities, especially the first one (which is not from the marginalization) serves to make sure that at intensity 0 i.e. $X_i = 0$ the predicted confidence will always be 0.5. 

Not only is this a strong assumption that we will challenge regardless, due to empirical findings of confidence bias (i.e. some subjects reporting higher confidence than others on average), but the main problem with this formulation is that with this anchor at (0,0.5) the confidence curves cannot produce one of the main empirical patterns of confidence ratings, which we highlighted in the [Models and Data Markdown](Models and Data in Perceptual Decision-Making with Confidence.Rmd). Here we formalized that confidence ratings reflect a noisy, subjective probability of being correct, incorporating response biases rather than purely objective performance. With the above formulation this is not possible as there is an objective anchor at $X_i = 0$. 

Due to these consideration, supported by some empirical validation on our pilot data below, we believe that this particular stimulus intensity, coming from an unjustified assumption of the agent having access to the stimulus intensity, needs to be accessed. 

One particular interesting idea of how to deal with this is to assume that the agent has a noisy readout of the stimulus strength, $\hat{X} \sim p(\hat{x}|\theta,c)$. The problem with such formulation is that the defining the probability distribution of the internal stimulus intensity is not trivial as its positively bound and therefore would not easily be marginalized as above. 

Another approach would be to replace this stimulus intensity value with a constant across subject parameter that would serve as a placeholder for idosyncratic differences in how subjects might use the stimulus intensity above and beyond what they should have access to. 

Instead of opting for these options we chose to assume that the agents do not have access to this stimulus intensity and therefore fix it to 1. This approach in its simplicity circumvents some of the troublesome computational aspects as well as obeying to what the agents is assumed to have access to, but just as importantly this makes it possible for the model to accommodate the emperical observation that the confidence curve follows the psychometric function of the binary choices.

This means that the final equations become

$$
P(\text{correct} \mid a = 1, X_i, D_i) = 

\Phi\left(
\left[

\frac{(CX_iD_i - C\beta)}{\sigma_e^2 + \sigma_m^2} +

\rho \cdot \frac{ \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)C}{\sqrt{\sigma_e^2 + \sigma_m^2} } 

\right]

\cdot\frac{1} {\sqrt{1 + \frac{C^2}{\sigma_e^2+\sigma_m^2}}}\right) \tag{7a}
$$


which for the other choice becomes:


$$
P(\text{correct} \mid a = -1, X_i, D_i) = 

\Phi\left(
\left[
\frac{(CX_iD_i - C\beta)}{\sigma_e^2 + \sigma_m^2} -

\rho \cdot \frac{ \lambda\left(-\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)C}{\sqrt{\sigma_e^2 + \sigma_m^2} } 

\right]

\cdot\frac{1} {\sqrt{1 + \frac{C^2}{\sigma_e^2+\sigma_m^2}}}\right) \tag{7b}
$$

with:
$$
\lambda(x) = \frac{\phi(x)}{\Phi(x)}
\\
C = \frac{2}{1.7}
$$

Of note in the above formula is that the terms inside the cummulative normal distribution function have clear interpretations. 

Firstly the common factor is a correction factor for marginalization. thus we define: 

$$
K_i = \frac{1} {\sqrt{1 + \frac{C^2}{\sigma_e^2+\sigma_m^2}}}
$$ 
whereas the two other terms can be thought to have a cognitive interpretation.

looking back at equation 3 we had (now with $X_i = 1$)

$$
P(\text{correct} \mid r_i, a, X_i) = \begin{cases}\Phi\left(\frac{C \cdot r_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - \Phi\left(\frac{C \cdot r_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases} \tag{3}
$$

we observe that the first term from equation 7 is just the expectation of the argument to the cumulative normal distribution function. This argument from the derivation is the log-likelihood ratio, which might also be seen as the average strength of evidence in favor of the true state. We might call this is I;

$$
I = \mathop{\mathbb{E}}\left(\frac{C \cdot r_i}{\sigma_e^2 + \sigma_m^2}\right) = \frac{(CX_iD_i - C\beta)}{\sigma_e^2 + \sigma_m^2}
$$

The latter term in equation 7 adjusts this expectation, because of the action selected before the confidence rating thus we name this the S. 

$$
S = -\rho \cdot \frac{ \lambda\left(-\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)C}{\sqrt{\sigma_e^2 + \sigma_m^2} }
$$

This term serves the purpose of changing the mean expectation to incorporate the information of the action to guide confidence. The addition of actions into the confidence computation have been proposed in the second order rater model (REF), from Fleming and Daw where they also re-analyzed data to show this pattern empirically. 



To summarize; Confidence computation can be thought of as a weighting of the expected normalized evidence with a correction factor for having commited to a choice. 

Now below we introduce a couple of additional parameters in order to encompass idiosyncrasies in the task and participants, as well as probabilistic relate the above formulation to our confidence ratings.

We introduce 5 additional parameters. $\Theta_2 = (\kappa_c, \lambda, \mu_{\beta})$. $\lambda$ is the lapse rate on the binary choices this is a stimulus independent factor that shifts the asympote of the psychometric function away from certainty i.e. the asymptotic becomes $\lambda$ and $1-\lambda$. Further we define $\mu_{\beta}$ as meta cognitive bias, which serves as an independent shift of confidence ratings. The parameter $\kappa_c$ is the precision of the ordered beta likelihood respectively. Lastly we introduce subject level parameters not hierarchically drawn which are the two cutpoints for the ordered beta likelihood $c_0,c_1$. These parameters help generalize the beta-distribution to outcomes that contain 0 and 1 ratings. The last three parameters serve to probabilistically relate the expected value of the confidence ratings to the confidence rating data. See (XXX) for the ordered beta likelihood. 

The complete set of parameters for the model are: 


$$
\Theta_s = \begin{bmatrix}
           beta_s \\
           \sigma_{e_s} \\
           \sigma_{c_s} \\
           \sigma_{m_s} \\
           \mu_{\beta_s} \\
           \kappa_{c_s} \\
           \lambda_s 
         \end{bmatrix} 
         \sim 
         \mathcal{N}(\mu_{\Theta},\Sigma_{\Theta}) 
\\         
\beta_s  \in [-1,1], \sigma_{e_s} >0, \sigma_{c_s} >0, \sigma_{m_s} > 0, 
\\
\mu_{\beta_s} \in \mathop{\mathbb{R}}, \kappa_{c_s} > 0, \lambda_s \in [0,0.5]
\\
(c_{1_s} , c_{0_s}) \sim \delta([1,10,1])
$$

Note in the above formulation that $\mathcal{N}$ is a multivariate normal distribution with mean and covariance. Further all contraints on the parameters as written were acomplised by transformation. All positively constrainted parameters were exponentiated, parameters with an upper and lower bound were transformed with the inverse logit function and multiplied appropriately. The constraint of the cutputs were applied by the prior of a dirchlet distribution. 

and the equations for model becomes:

$$
\\
P(a = 1) = \Phi\left(\frac{D_i \cdot X_i- \beta_s}{\sqrt{\sigma_{e_s}^2 + \sigma_{c_s}^2}}\right)
\\
\mu_{c_i} = \begin{cases}\Phi\left((I_i-S_i)\cdot C_i\right) & \text{if } a_i = 1 
\\ 
1 - \Phi\left((I_i+S_i)\cdot C_i\right) & \text{if } a_i = -1\end{cases}
\\
a_i \sim \mathcal{bernoulli}(p(a_i = 1))
\\
Conf_i \sim \mathcal{ord\beta}(\mu_{c_i},\kappa_{c_s},c_{0_s},c_{1_s})
$$

with

$$
I_i = \frac{(CX_iD_i - C\beta)}{\sigma_{e_s}^2 + \sigma_{m_s}^2},
S_i = -\rho_s \cdot \frac{ \lambda\left(-\frac{X_iD_i - \beta_s}{\sqrt{\sigma_{e_s}^2 + \sigma_{c_s}^2}}\right)C}{\sqrt{\sigma_{e_s}^2 + \sigma_{m_s}^2} },
K_i = \frac{1} {\sqrt{1 + \frac{C^2}{\sigma_{e_s}^2+\sigma_{m_s}^2}}}

$$




Below we showcase the above heuristic model fit to data simulated from equation (2) (with $X_i = 1$) to show that the two are approximately close. 

```{r, error=F,message=F,warning=F, fig.height=10,fig.width=10}
source("~/RDM-metacognition/Preregistration/scripts/Simulation.R")

library(tidyverse)

# Define high/low values
sigma_e_vals <- c(-2, -1.5)   # example low/high
sigma_k_vals <- c(-3, -2)
sigma_m_vals <- c(-1, 0)


param_grid <- expand.grid(
  sigma_e = sigma_e_vals,
  sigma_k = sigma_k_vals,
  sigma_m = sigma_m_vals,
  bias    = 0  # start with bias = 0 for all
)

# Add an extra row for the one case where bias != 0 (if desired)
param_grid <- param_grid%>% 
  mutate(id = row_number())

param_grid <- param_grid %>%
  mutate(facet_label = paste0("S_e=", sigma_e, 
                              ", S_k=", sigma_k, 
                              ", S_m=", sigma_m, 
                              ", B=", bias))

breaks <- seq(-1, 1, by = 0.05)
midpoints <- breaks[-length(breaks)] + 0.1

df_mech = param_grid %>% 
    rowwise() %>%
    mutate(sims = list(simulate_data_p_bias(100000, sigma_e = sigma_e, sigma_k = sigma_k, sigma_m = sigma_m, bias = bias, seed = 1))) %>% 
    ungroup() %>% 
    unnest() %>% 
    mutate(XD = X * D,
    bin = cut(XD, breaks = breaks, include.lowest = TRUE, labels = FALSE),
    XD = (breaks[bin] + breaks[bin + 1]) / 2) %>% 
  mutate(Action = as.factor(a_gauss),
         Accuracy = as.factor(ACC))  

df_heu = param_grid %>% 
    rowwise() %>%
    mutate(sims = list(simulate_data_margin_bias(5000, sigma_e = sigma_e, sigma_k = sigma_k, sigma_m = sigma_m, bias = bias, seed = 1))) %>% 
    ungroup() %>% 
    unnest() %>% 
    mutate(XD = X*D) %>% 
  mutate(bin = NA)%>% 
  mutate(Action = as.factor(a_gauss),
         Accuracy = as.factor(ACC))
    
  

rbind(df_mech,df_heu) %>% ggplot()+
  geom_pointrange(data = df_mech %>% 
                    group_by(XD, Accuracy,facet_label) %>% 
                    summarize(mean = mean(c_mu), 
                              se = sd(c_mu)/sqrt(n())),
                  aes(x = XD,y = mean,ymin = mean-2*se, ymax = mean+2*se, fill = Accuracy), shape = 21, col = "black")+
  geom_pointrange(data = df_mech %>% 
                    group_by(XD, facet_label) %>% 
                    summarize(k = sum(as.numeric(as.character(Action))),
                              n = n(),
                              prop = k / n,
                              a_post = k + 1,
                              b_post = (n - k) + 1,
                              lower = qbeta(0.025, a_post, b_post),
                              upper = qbeta(0.975, a_post, b_post)),
                  aes(x = XD,y = prop,ymin = lower, ymax = upper), col = "black", shape = 21,fill = "black")+
  geom_line(data = df_heu, aes(x = XD,y = c_mu, col = Accuracy), linewidth = 2, alpha = 0.75)+
  geom_line(data = df_heu, aes(x = XD,y = p_gauss), linewidth = 2, alpha = 0.75, col = "black")+
  theme_classic()+
  facet_wrap(~facet_label, scales = "free") +
  labs(x = "Stimulus Intensity")+
  geom_vline(xintercept = 0, linetype = 2)+
  geom_hline(yintercept = 0, linetype = 2)+
  geom_hline(yintercept = 1, linetype = 2)+
  theme(legend.position = "top")
  
  
```


```{r, error=F,message=F,warning=F, fig.height=10,fig.width=10}

# Define high/low values
sigma_e_vals <- c(-1.5)   # example low/high
sigma_k_vals <- c(-2)
sigma_m_vals <- c(-1.2)


param_grid <- expand.grid(
  sigma_e = sigma_e_vals,
  sigma_k = sigma_k_vals,
  sigma_m = sigma_m_vals,
  bias    = seq(-0.7,0.7,length.out = 9)
)

# Add an extra row for the one case where bias != 0 (if desired)
param_grid <- param_grid%>% 
  mutate(id = row_number())

param_grid <- param_grid %>%
  mutate(facet_label = paste0("S_e=", sigma_e, 
                              ", S_k=", sigma_k, 
                              ", S_m=", sigma_m, 
                              ", B=", bias))

breaks <- seq(-1, 1, by = 0.05)
midpoints <- breaks[-length(breaks)] + 0.1

df_mech = param_grid %>% 
    rowwise() %>%
    mutate(sims = list(simulate_data_p_bias(100000, sigma_e = sigma_e, sigma_k = sigma_k, sigma_m = sigma_m, bias = bias, seed = 1))) %>% 
    ungroup() %>% 
    unnest() %>% 
    mutate(XD = X * D,
    bin = cut(XD, breaks = breaks, include.lowest = TRUE, labels = FALSE),
    XD = (breaks[bin] + breaks[bin + 1]) / 2) %>% 
  mutate(Action = as.factor(a_gauss),
         Accuracy = as.factor(ACC))  

df_heu = param_grid %>% 
    rowwise() %>%
    mutate(sims = list(simulate_data_margin_bias(5000, sigma_e = sigma_e, sigma_k = sigma_k, sigma_m = sigma_m, bias = bias, seed = 1))) %>% 
    ungroup() %>% 
    unnest() %>% 
    mutate(XD = X*D) %>% 
  mutate(bin = NA)%>% 
  mutate(Action = as.factor(a_gauss),
         Accuracy = as.factor(ACC))
    
  

rbind(df_mech) %>% ggplot()+
  geom_pointrange(data = df_mech %>% 
                    group_by(XD, Action,facet_label) %>% 
                    summarize(mean = mean(c_mu), 
                              se = sd(c_mu)/sqrt(n())),
                  aes(x = XD,y = mean,ymin = mean-2*se, ymax = mean+2*se, fill = Action), shape = 21, col = "black")+
  geom_pointrange(data = df_mech %>% 
                    group_by(XD, facet_label) %>% 
                    summarize(k = sum(as.numeric(as.character(Action))),
                              n = n(),
                              prop = k / n,
                              a_post = k + 1,
                              b_post = (n - k) + 1,
                              lower = qbeta(0.025, a_post, b_post),
                              upper = qbeta(0.975, a_post, b_post)),
                  aes(x = XD,y = prop,ymin = lower, ymax = upper), shape = 21, col = "black",fill = "black")+
  geom_line(data = df_heu, aes(x = XD,y = c_mu, col = Action), linewidth = 2, alpha = 0.75)+
  geom_line(data = df_heu, aes(x = XD,y = p_gauss), linewidth = 2, alpha = 0.75, col = "black")+
  theme_classic()+
  facet_wrap(~facet_label, scales = "free") +
  geom_vline(xintercept = 0, linetype = 2)+
  geom_hline(yintercept = 0, linetype = 2)+
  geom_hline(yintercept = 1, linetype = 2)+
  labs(x = "Stimulus Intensity")+
  theme(legend.position = "top")
  

  
```

Note that each facet contains the parameters that were used for simulating the data for both models. Datapoints with error bars are derived from averaging 100000 points across the -1 ; 1 stimulus range in increments of 0.05 (error bars contain approximate 95% converage) from equation (2). Lines are the expected values derived from equation (7) from 5000 points. As can be seen the approximation for the expectation of the trial-by-trial model is excellent.



Next, we demonstrate parameter recovery simulating from equation (2) (with $X_i = 1$) and fitting with the heuristic model derived above. 


```{r}
 
```


Lastly, we fit this model to our pilot data hierarchically and compare to the single subject estimated from equation (2). 

## Model comparison

```{r, echo = T}
model_with_x <- readRDS(here::here("Experimental analysis","Heurestic models",
                                   "saved models","saved models hier","fit_finalheu_lapse_cypertron.rds"))


model_with_x1 <- readRDS(here::here("Experimental analysis","Heurestic models",
                                    "saved models","saved models hier","fit_finalheu_lapse_x1.rds"))

loocomparison = loo::loo_compare(list(with_x = model_with_x$loo(), with_x1 = model_with_x1$loo()))
```

As a sanity check on the assumption of $X_i = 1$, we fitted both hierarchical models to our pilot data and computed approximate leave one out cross-validation, to test if the assumption could be justified. This comparison showed that the two models were essentially indistinguishable with the difference in expected log probability density being `r round(loocomparison[2,1],2)` $\pm$ `r round(loocomparison[2,2],2)`, where generally it is thought that a difference below -4 and or more than two standard errors can be considered as evidence for one model over the other (REF), thus atleast for our pilot data for predictive performance there seems to be no difference. Digging further into the distinction between these two models we below plot all pairwise parameters together with the identity line.

```{r,fig.height=6, fig.width=8, message=F,error=F,warning=F}
parameters = c("beta","sigma_e","sigma_k","sigma_m","confprec","lapse","meta_bias","confprec","meta_bias_beta","sigmam_beta")


rbind(
    model_with_x$summary(parameters) %>% 
      mutate(
        subject  = as.numeric(str_extract(variable, "(?<=\\[)\\d+(?=\\])")),
        variable = str_remove(variable, "\\[\\d+\\]"),
        model    = "X"
      ),
    
    model_with_x1$summary(parameters) %>% 
        mutate(
        subject  = as.numeric(str_extract(variable, "(?<=\\[)\\d+(?=\\])")),
        variable = str_remove(variable, "\\[\\d+\\]"),
        model    = "Xone"
      )
    )%>% select(variable,model,mean,q5,q95) %>% 
      pivot_wider(names_from = "model",values_from = c("mean","q5","q95")) %>% unnest() %>% 
  ggplot(aes(x = mean_X, y = mean_Xone))+
  geom_pointrange(aes(ymin = q5_Xone, ymax = q95_Xone))+
  geom_pointrange(aes(xmin = q5_X, xmax = q95_X))+
  facet_wrap(~variable, scales = "free")+
  labs(y = "Model with X = 1", x = "Model with X")+
  geom_abline()+
  theme_classic()

```

Generally the two models align well in their subject level parameter estimates, further justifing this simplification that makes theoretical sense.




```{r,fig.height=12, fig.width=18, message=F,error=F,warning=F}
magick::image_read(here::here("Preregistration","Figures","Pilot_predictive_checks.PNG"))
```


## pilot data fit from Hierarchical Heurestic model

```{r,fig.height=12, fig.width=18, message=F,error=F,warning=F}
magick::image_read(here::here("Preregistration","Figures","Pilot_predictive_checks.PNG"))
```



## Hierarchical parameter recovery

Using the data from the simulations of the sequential sampling we also have access to parameter recovery, where the simulated datasets were extracted from the draws of the above posterior distribution, below we showcase that the model can to an adequte degree recover its parameters on a subject and group level:

```{r,fig.height=7, fig.width=14, message=F,error=F,warning=F}
magick::image_read(here::here("Preregistration","Figures","parameterrecovery_hier.PNG"))
```




