---
title: "Current model"
author: Jesper Fischer Ehmsen^1^,
        Siebe Everaets^xx^,
        Micah G. Allen^1,2,\*^,
        Francesca Fardo^1,3,\*^
output:
  html_document: default
  word_document:
    # reference_docx: Knitting files/docx_template.docx
  pdf_document:
    latex_engine: xelatex
# bibliography: 
  # - "Knitting files/References.bib"
link-citations: yes
linkcolor: blue
# csl: Knitting files/science-advances.csl
always_allow_html: yes
---

::: {custom-style="Title text"}
^1^Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University

^2^Cambridge Psychiatry, University of Cambridge, United Kingdom

^3^Danish Pain Research Center, Department of Clinical Medicine, Aarhus University

^\*^Equal contributions

Corresponding author: Jesper Fischer Ehmsen (jesperfischer@cfin.au.dk)
:::

```{r load packages, warning = F, message = F, echo = F, include=FALSE}
knitr::opts_chunk$set(dpi=300)

# seed
set.seed(123)

# packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load("here", "knitr","yaml")

knitr::opts_chunk$set(echo = TRUE)

#renv::restore(project=here::here())

required_packages = c("cowplot", "magick", "tidyverse", "flextable")
lapply(required_packages, library, character.only = TRUE)
```




### Static models of metacognition

Here we outline the general idea between the static models of perceptual deicison making and confidence judgements.

In SDT, an actor makes a binary choice $a \in \{-1,+1\}$ based on evidence $e$ about the true state $D \in \{-1,+1\}$. Evidence is drawn from a normal distribution:

$$
e_i \sim \mathcal{N}(X_iD_i,\sigma)
$$

Here inorder to incompass **3** from the Patterns of metacognition section we introduce a parameter that can shift this evidence distribution, essentially introducing a bias in the encoding of evidence.

$$
e_i \sim \mathcal{N}(X_iD_i - \beta,\sigma)
$$

where $X_i$ is stimulus strength. The actor compares evidence to criterion $c$:

$$
a_i = \begin{cases}+1 & \text{if } e_i > c \\ -1 & \text{if } e_i \le c\end{cases}
$$

For metacognition, the agent accesses this evidence for information about the correctness of the action.

Models differ in whether the rater has independent information above the evidence used to make the action. Independent information is usually assumed iorder for the agent to do error-monitoring.

Fleming and Daw formalized a second order rater where evidence for the action and evidence for the confidence judgement is draw from a multivariate normal with correlation $\rho$ governing information sharing. I.e. if $\rho$ = 1, the confidence judgement is based on the same information as the action, however if $\rho$ = 0 the confidence judgement is based on completely independent information.

Note we here write this model as orginally proposed by Fleming and Daw without the evidence bias

$$
\begin{bmatrix}e_i \\ r_i\end{bmatrix}\sim \mathcal{N}\left(\begin{bmatrix}X_iD_i \\ X_iD_i\end{bmatrix},\begin{bmatrix}\sigma_e^2 & \rho \cdot \sigma_e \cdot \sigma_m \\ \rho \cdot \sigma_e \cdot \sigma_m & \sigma_m^2\end{bmatrix}\right)
$$

The rater uses this information to compute $P(D = a | r_i)$, the subjective probability of being correct.

One of the main arguments for the development of this and other models with independent information for the confidence judgement was to explain error-monitoring.

Below we formalize the idea that such independent information is not necessary, if instead it is assumeed that agents have varability in their choices, given the evidence. I.e. instead of a deterministic decision policy agents have a noisy decision process that can produce errors even though the evidence is in favor of a different choice.

## Choice Variability

We extend the standard SDT framework by introducing stochastic choice policy. Evidence is drawn as:

$$
e_i \sim \mathcal{N}(X_iD_i- \beta,\sigma_e)
$$

But now the criterion varies trial-by-trial: $c_i \sim \mathcal{N}(0,\sigma_c)$. This yields:

$$
a_i = \begin{cases}+1 & \text{if } e_i > c_i \\ -1 & \text{if } e_i \le c_i\end{cases}
$$

Due to the fact that both $e_i$ and $c_i$ are normally distributed one can marginalize their difference which gives the probability of the action:

$$
P(a = 1 \mid D_i, X_i) = \Phi\left(\frac{D_i \cdot X_i- \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$
Or equivalently:

$$
P(a = 1 \mid e_i) = \Phi\left(\frac{e_i}{\sigma_c}\right)
$$

From choices alone, only $\sigma_e^2 + \sigma_c^2$ is identifiable, both noise sources equivalently flatten the psychometric curve. However, by including confidence ratings these can be disentangled.

For confidence, we define:

$$
\text{confidence} = P(\text{correct} \mid r, a) = P(D = a \mid r)
$$

The rater accesses noisy evidence: $r_i \sim \mathcal{N}(e_i, \sigma_m^2)$, where $\sigma_m^2$ is metacognitive noise. 

This can be shown to yield:

$$
P(D=1 \mid r_i, \mu_i) = g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right)
$$

giving:

$$
P(\text{correct} \mid r_i, a, \mu_i) = \begin{cases}g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases}
$$

From confidence alone, only $\sigma_e^2 + \sigma_m^2$ is identifiable. However, combining choices and confidence allows estimating all three parameters: $\sigma_e$, $\sigma_c$, and $\sigma_m$. 

One problem with the above formulation is that in-order to estimate the parameters one has to estimate the trial-by-trial evidence $r_i$ which can be possible with Bayesian data augmentation. As the normative probability of being correct i.e. $P(correct | r_i,a)$ is a logit-normally distributed variable it cannot be marginalized as the choice probability can. Below we show simulations from this model in order to provide an intuition for the meaning of the parameters and their interactions.


## Simulations

We demonstrate that this model reproduces key behavioral patterns (folded-X, double-increase) under different parameter combinations. Simulations show that the ratio $\frac{\sigma_e}{\sigma_c}$ determines these patterns: 

A low ratio (high $\sigma_c$ low $\sigma_e$) produces folded-X (strong error-monitoring), while a high ratio (low $\sigma_c$ high $\sigma_e$) produces double-increase (weak error-monitoring).

Metacognitive noise $\sigma_m$ flattens confidence curves regardless of accuracy.

Note that for these simulations we assume that the bias in evidence encoding $\beta$ is 0 as this just shifts the psychometric function as well as the confidence curves horizontally

```{r}
set.seed(199742)

n_trials <- 500

simulate_interaction <- function(sigma_m, sigma_k,sigma_e) {
  X <- abs(rnorm(n_trials,2,1))
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_k)
  p_gauss <- pnorm(e / sigma_k)
  a <- rbinom(n_trials, 1, p_gauss)

  data.frame(e = e,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}


df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_k = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_k,sigma_e))) %>% 
  unnest(data)

```

Now lets turn to how evidence, the physical stimulus and accuracy interacts across parameters:

```{r, fig.height=6, fig.width=10,warning=F,message=F}

ACC = df_interaction %>%
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = X*D, y = e, col = Accuracy)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point()+
  facet_grid(sigma_e~sigma_k, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))

ACC
```

From the above plot its obvious that $\sigma_e$ scales the the range of evidence observed (see y-axis range for the different rows). When $\sigma_k$ is low (left column) choices are consistent (when the the evidence is positive and the signed stimulus is positive i.e. D = 1 then a correct response is given). When $\sigma_k$ increases this relationship essentially becomes more noisy with correct and incorrect choices being less distinguishable across the evidence = 0 line. Another way to display this is by plotting the actions instead of the accuracy: 

here the space is only split in two quadrants instead of 4.

```{r, fig.height=6, fig.width=10,warning=F,message=F}
Act = df_interaction %>%
  mutate(Action = as.factor(a)) %>% 
  ggplot(aes(x = X*D, y = e, col = Action)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point()+
  facet_grid(sigma_e~sigma_k, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))
Act
```


Next we look at the realized confidence ratings as a function of the three different sigmas:

```{r}
set.seed(199742)

n_trials <- 5000

conf_prec <- 100
c0 <- -3
c1 <- 3

simulate_interaction <- function(sigma_m, sigma_k,sigma_e) {
  X <- sample(c(0.01,1,2,5), n_trials, replace = TRUE)
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_k)
  p_gauss <- pnorm(e / sigma_k)
  a <- rbinom(n_trials, 1, p_gauss)
  C_mu <- ifelse(a == 1, 
                 plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)),
                 1 - plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)))
  

  c_mu <- pmax(0.001, pmin(0.999, C_mu))
  C <- ordbetareg::rordbeta(n_trials, c_mu, conf_prec, c(c0, c1))
  
  data.frame(e = e,
             C = C,
             C_mu = C_mu,
             # C_mu2 = C_mu2,
             ehat = ehat,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}

df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_k = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_k,sigma_e))) %>% 
  unnest(data)
```



First we can look to fix $\sigma_k = 1$ and then we look at how $\sigma_e$ & $\sigma_m$ change confidence ratings as a function of the Stimulus:

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_k == 1) %>% 
  mutate(SS = D*X) %>%
  group_by(SS,ACC, sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

From the above plot we see that when we plot confidence on the Signed Stimulus scale we end up with a somewhat double increase pattern, when the stimulus encoding uncertainty is high (i.e. 3 or 5). Furthermore we see a (")half) folded X when stimulus encoding uncertainty  is low. Generally we see that meta uncertainty makes the confidence ratings more flat lined at 0.5 regardless of their correctness.  

Now we fix $\sigma_e = 1$ and then we look at $\sigma_k$ & $\sigma_m$:

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_e == 1) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_k,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_k~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

This plot generally shows the same as the one with $\sigma_k = 1$ but opposite in pattern. I.e. high $\sigma_k$ here is more of a folded X with lower values (espeically below 1) make a double increase.

lastly we might plot $sigma_m = 1$

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_m == 1) %>%
  mutate(sigma = sigma_e/sigma_k) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_k,sigma,sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>%
  mutate(sigma_m = as.factor(sigma_m)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~sigma_k, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We see that the diagonal where $\sigma_e = \sigma_k$ that there generally is a half-folded X, with incorrect trials being flat with stimulus strength with correct trials increasing in confidence with stimulus strength. Going to the upper triangle we move into the space where $\sigma_k > \sigma_e$ where we see folded-X patterns of varying "strengths". Moving to the lower triangle we see the opposite pattern with double increase patterns emerging again with varying "strengths". 

This means, that from the models perspective, double increase and Folded-X patterns arise from the ratio of uncertainty in the encoding of the stimulus and the choice variability. When choice variability is high compared to stimulus encoding a folded X pattern arises whereas choice variability is low compared to stimulus encoding a double increase pattern arises. $\sigma_m$, meta uncertainty is agnostic to the accuracy of the rater and when it increases it makes the curves of the confidence ratings more flat regardless of correctness.

## Fitting

We fit this model using full Bayesian inference with Stan's NUTS sampler, treating latent trial-by-trial evidence as parameters estimated jointly with the noise parameters. Generally, from investigating our pilot data we saw that on a single-subject basis this model is possible to fit with sufficiently good sampler diagnoistics, however trying to fit this model in a hierarchical framework i.e. estimating group level parameters which subjects level parameters are drawn from we ran into sampling issues. 

We therefore build a heurestic model that approximates the above model, without the need for the trial-by-trial evidence. This is done such that hierarchical bayesian inference can be done reliably for the Sequential sampling approach used for sample size determined in the current study (see Sequential sampling section).

## A heuristic model

While the mechanistic model is powerful, its computational complexity (especially in hierarchical settings) motivate a simpler heuristic approximation. The key challenge is that confidence requires trial-specific evidence $r_i$, which cannot be marginalized analytically, because the random variable is inside the inverse logit function. Now The inverse logit function and the cummulative normal distribution are basically in practice indistinguishable up to a constant i.e. $g(.) \sim \Phi(\frac{1}{1.7})$ using this approximation it is possible to marginalize out the trial-by-trial evidence and get a computationally easier model to fit (and an equation for the expectation of confidence) with the same parameters. Thus utilizing the above approximation we go from

$$
P(\text{correct} \mid r_i, a, \mu_i) = \begin{cases}g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases}
$$
to

$$
P(\text{correct} \mid r_i, a, \mu_i) = \begin{cases}\Phi\left(\frac{C \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - \Phi\left(\frac{C \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases}
$$

with $C = \frac{2}{1.7}$. Now looking at the condition where the choice was one.

$$
\Phi\left(\frac{C \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right)
$$

We want to marginalize over $r_i$ i.e. calculating the expectation of above equation across $r_i$. This is can be done using the general formula for the expectation of a random continous variable

$$
\mathbb{E(X)} = \int_{-\infty}^\infty x \ p(x) dx
$$

substituting the variables with the following definition $b = \frac{CX_i}{\sigma_e^2 + \sigma_m^2}$

$$
\int_{-\infty}^\infty \Phi(b \cdot r_i) \ p(r_i) dr_i
$$

This integral can be shown to have the solution:

$$
\Phi \left(\frac{\mu \cdot b}{\sqrt{1+\sigma^2 \cdot b^2}}\right)
$$


with $r_i \sim \mathcal{N}(\mu,\sigma)$

given that with  $r_i \sim \mathcal{N}(X_iD_i - \beta, \sqrt{\sigma_m^2+\sigma_e^2})$ we get: 

$$
\Phi \left(\frac{\mathop{\mathbb{E}}(r_i) \cdot b}{\sqrt{1+(\sigma_e^2+\sigma_m^2) \cdot b^2}}\right)
$$

Here we instead of the mean of the $r_i$ distribution ($X_iD_i - \beta$) we write the expectation of $r_i$ as this will be conditional on the choice later. However, before calculating this conditional expectation we simplify the above:

$$
\Phi \left( \frac{\mathop{\mathbb{E}}(r_i) \cdot \frac{CX_i}{\sigma_e^2 + \sigma_m^2}}{\sqrt{1+(\sigma_e^2+\sigma_m^2) \cdot (\frac{CX_i}{\sigma_e^2 + \sigma_m^2})^2}}\right)
$$

$$
\Phi \left( \frac{\mathop{\mathbb{E}}(r_i) \cdot \frac{CX_i}{\sigma_e^2 + \sigma_m^2}}{\sqrt{1+(\sigma_e^2+\sigma_m^2) \cdot (\frac{C^2X_i^2}{(\sigma_e^2+\sigma_m^2)^2})}}\right)
$$

$$
\Phi \left( \frac{\mathop{\mathbb{E}}(r_i) \cdot \frac{CX_i}{\sigma_e^2 + \sigma_m^2}}{\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)
$$



$$
\Phi\left(\frac{\mathop{\mathbb{E}}(r_i) \cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)
$$

This equation thus gives us the probability of being correct, when the choice is 1. Thus the expectation of meta cognitive evidence $r_i$ also has to be conditional on the fact that the choice was 1. Looking back we had:

$$
P(\text{correct} \mid r_i, a, X_i) \sim \begin{cases}\Phi\left( \frac{Cr_iX_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - \Phi\left( \frac{Cr_iX_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases}
$$
which now becomes:

$$
P(\text{correct} \mid a, X_i) 

\sim 

\begin{cases}\Phi\left(\frac{\mathop{\mathbb{E}}(r_i \mid a = 1) \cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)

& 

\text{if } a = 1 \\ 

1 - \Phi\left(\frac{\mathop{\mathbb{E}}(r_i \mid a = -1) \cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)

& 

\text{if } a = -1\end{cases}
$$

Thus, now all that there is left to calculate is the two expectations of $r_i$ given the choice i.e.
$\mathop{\mathbb{E}}(r_i \mid a = 1)$ & $\mathop{\mathbb{E}}(r_i \mid a = -1)$

### Calculating conditional expectations

To calculate these expectations, we need to account for the fact that the choice $a$ depends on the relationship between the evidence $e_i$ and the criterion $c_i$. We have:

- Evidence: $e_i \sim \mathcal{N}(X_iD_i - \beta, \sigma_e^2)$
- Criterion: $c_i \sim \mathcal{N}(0, \sigma_c^2)$
- Metacognitive evidence: $r_i = e_i + \epsilon_m$ where $\epsilon_m \sim \mathcal{N}(0, \sigma_m^2)$
- Choice: $a = 1$ if $e_i > c_i$, otherwise $a = -1$


we now treat the difference distribution $e_i - c_i$ as a joint bivariate normal distribution with $r_i$  and calculate the expectation of $r_i$ given that the $e_i - c_i$ random variable is above 0 (i.e. the choice was 1).

As used before we have that the difference $e_i - c_i$ follows another Gaussian distribution, as both random variables are Gaussians.

$$
e_i - c_i \sim \mathcal{N}(X_iD_i - \beta, \sigma_e^2 + \sigma_c^2)
$$

The joint bivariate normal distribution thus becomes:
$$
\begin{bmatrix}
e_i - c_i \\
r_i
\end{bmatrix}
\sim \mathcal{N}\left(
\begin{bmatrix}
X_iD_i - \beta \\
X_iD_i - \beta
\end{bmatrix},
\begin{bmatrix}
\sigma_e^2 + \sigma_c^2 & \sigma_e^2 \\
\sigma_e^2 & \sigma_e^2 + \sigma_m^2
\end{bmatrix}
\right)
$$

The key insight here is that the **covariance** between $e_i-c_i$ and $r_i$ is $\sigma_e^2$, not zero. This is because $r_i$ is constructed as $e_i$ plus independent noise, so $\text{Cov}(e_i-c_i, r_i) = \text{Cov}(e_i-c_i, r_i + \epsilon_m) = \text{Var}(e) = \sigma_e^2$. This shared variance structure means that $e_i-c_i$ and $r_i$ are correlated with correlation coefficient:

## Truncated normal distributions

To proceed, we need a general result from probability theory. Suppose we have two jointly normal random variables:

$$
\begin{pmatrix} Z \\ Y \end{pmatrix} \sim \mathcal{N}\left(\begin{pmatrix} \mu_Z \\ \mu_Y \end{pmatrix}, \begin{pmatrix} \sigma_Z^2 & \rho\sigma_Y\sigma_Z \\ \rho\sigma_Y\sigma_Z & \sigma_Y^2 \end{pmatrix}\right)
$$

where $\rho$ is the **correlation coefficient**, defined as:

$$
\rho = \frac{\text{Cov}(Z, Y)}{\sigma_Z \sigma_Y}
$$


We want to find $\mathbb{E}(Y \mid Z > 0)$, the expected value of $Y$ conditional on $Z$ being positive (i.e. the choice was 1).

From standard bivariate normal theory, the conditional distribution of $Y$ given $Z = z$ is:

$$
Y \mid Z = z \sim \mathcal{N}\left(\mu_Y + \frac{\rho\sigma_Y}{\sigma_Z}(z - \mu_Z), \sigma_Y^2(1-\rho^2)\right)
$$

So the expectation is:

$$
\mathbb{E}(Y \mid Z = z) = \mu_Y + \frac{\rho\sigma_Y}{\sigma_Z}(z - \mu_Z)
$$

Now we want the expectation of Y when Z > 0 (i.e. the expectation of $r_i$ when the choice was 1) i.e. $\mathbb{E}(Y \mid Z > 0)$ we would have to do the integral which gives:

**Derivation using the Law of Total Expectation:**

Since we want the expected value of $Y$ conditional on $Z > 0$, we need to average over all possible values of $Z$ that satisfy $Z > 0$, weighted by their probability densities. The law of total expectation for conditioning on a subset states:

$$
\mathbb{E}(Y \mid Z > 0) = \frac{\int_{0}^{\infty} \mathbb{E}(Y \mid Z = z) \cdot f_Z(z) \, dz}{\int_{0}^{\infty} f_Z(z) \, dz}
$$

The denominator is simply $P(Z > 0) = \Phi\left(\frac{\mu_Z}{\sigma_Z}\right)$, so:

$$
\mathbb{E}(Y \mid Z > 0) = \frac{1}{P(Z > 0)} \int_{0}^{\infty} \mathbb{E}(Y \mid Z = z) \cdot f_Z(z) \, dz
$$


From earlier, we know that $\mathbb{E}(Y \mid Z = z) = \mu_Y + \frac{\rho\sigma_Y}{\sigma_Z}(z - \mu_Z)$, and the marginal density is $f_Z(z) = \frac{1}{\sigma_Z}\phi\left(\frac{z-\mu_Z}{\sigma_Z}\right)$:

$$
\mathbb{E}(Y \mid Z > 0) = \frac{1}{\Phi\left(\frac{\mu_Z}{\sigma_Z}\right)} \int_{0}^{\infty} \left[\mu_Y + \frac{\rho\sigma_Y}{\sigma_Z}(z - \mu_Z)\right] \cdot \frac{1}{\sigma_Z}\phi\left(\frac{z-\mu_Z}{\sigma_Z}\right) \, dz
$$

**Step 2:** Separate the integral into two parts:

$$
\mathbb{E}(Y \mid Z > 0) = \mu_Y \cdot \frac{\int_{0}^{\infty} \frac{1}{\sigma_Z}\phi\left(\frac{z-\mu_Z}{\sigma_Z}\right) dz}{\Phi\left(\frac{\mu_Z}{\sigma_Z}\right)} + \frac{\rho\sigma_Y}{\sigma_Z} \cdot \frac{\int_{0}^{\infty} (z - \mu_Z) \cdot \frac{1}{\sigma_Z}\phi\left(\frac{z-\mu_Z}{\sigma_Z}\right) dz}{\Phi\left(\frac{\mu_Z}{\sigma_Z}\right)}
$$

The first integral equals $P(Z > 0) = \Phi\left(\frac{\mu_Z}{\sigma_Z}\right)$, so the first term simplifies to $\mu_Y$.

**Step 3:** Standardize the remaining integral using substitution $u = \frac{z-\mu_Z}{\sigma_Z}$:

When $z = 0$, we have $u = \frac{-\mu_Z}{\sigma_Z}$, and when $z \to \infty$, $u \to \infty$. Also, $z - \mu_Z = \sigma_Z u$ and $dz = \sigma_Z du$:

$$
\mathbb{E}(Y \mid Z > 0) = \mu_Y + \rho\sigma_Y \cdot \frac{\int_{\frac{-\mu_Z}{\sigma_Z}}^{\infty} u \cdot \phi(u) \, du}{\Phi\left(\frac{\mu_Z}{\sigma_Z}\right)}
$$

**Step 4:** Recognize the inverse Mills ratio:

The integral $\frac{\int_{-a}^{\infty} u \cdot \phi(u) \, du}{\Phi(a)} = \lambda(a)$ is the inverse Mills ratio, where $\lambda(a) = \frac{\phi(a)}{\Phi(a)}$.

Therefore, substituting $a = \frac{\mu_Z}{\sigma_Z}$:

$$
\mathbb{E}(Y \mid Z > 0) = \mu_Y + \rho\sigma_Y \cdot \lambda\left(\frac{\mu_Z}{\sigma_Z}\right)
$$

Equivalently, using $\rho\sigma_Y = \frac{\text{Cov}(Y,Z)}{\sigma_Z}$:

$$
\mathbb{E}(Y \mid Z > 0) = \mu_Y + \frac{\text{Cov}(Y,Z)}{\sigma_Z} \cdot \lambda\left(\frac{\mu_Z}{\sigma_Z}\right)
$$

For the case where we condition on $Z \leq 0$, a similar derivation gives:

$$
\mathbb{E}(Y \mid Z \leq 0) = \mu_Y - \rho\sigma_Y \cdot \lambda\left(-\frac{\mu_Z}{\sigma_Z}\right)
$$

#### Applying truncated normal theory to our problem

Now we apply this general result to our specific case. We need to identify:

- $Y = r_i$ with $\mu_Y = X_iD_i - \beta$ and $\sigma_Y^2 = \sigma_e^2 + \sigma_m^2$
- $Z = e_i - c_i$ with $\mu_Z = X_iD_i - \beta$ and $\sigma_Z^2 = \sigma_e^2 + \sigma_c^2$
- **Covariance**: $\text{Cov}(r_i, e_i - c_i) = \sigma_e^2$.
- Correlation coefficient: $\rho = \frac{\text{Cov}(r_i, e_i - c_i)}{\sigma_Y \sigma_Z} = \frac{\sigma_e^2}{\sqrt{(\sigma_e^2 + \sigma_m^2)(\sigma_e^2 + \sigma_c^2)}}$

Note that we use the covariance $\sigma_e^2$ that we calculated earlier to determine $\rho$.

The product $\rho\sigma_Y$ becomes

$$
\rho\sigma_Y = \frac{\sigma_e^2}{\sqrt{(\sigma_e^2 + \sigma_m^2)(\sigma_e^2 + \sigma_c^2)}} \cdot \sqrt{\sigma_e^2 + \sigma_m^2} = \frac{\sigma_e^2}{\sqrt{\sigma_e^2 + \sigma_c^2}}
$$

Applying the formula with $a = 1$ corresponding to $Z > 0$:

For $a = 1$ (when $e_i - c_i > 0$):
$$
\mathop{\mathbb{E}}(r_i \mid a = 1) = (X_iD_i - \beta) + \frac{\sigma_e^2}{\sqrt{\sigma_e^2 + \sigma_c^2}} \cdot \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$

For $a = -1$ (when $e_i - c_i \leq 0$):

$$
\mathop{\mathbb{E}}(r_i \mid a = -1) = (X_iD_i - \beta) - \frac{\sigma_e^2}{\sqrt{\sigma_e^2 + \sigma_c^2}} \cdot \lambda\left(-\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$

where $\lambda(x) = \frac{\phi(x)}{\Phi(x)}$ is the **inverse Mills ratio**, with $\phi(\cdot)$ being the standard normal PDF and $\Phi(\cdot)$ being the standard normal CDF.

Note that $\lambda(-x) = \frac{\phi(x)}{1-\Phi(x)}$ by properties of the normal distribution.

### Final confidence model

Substituting these conditional expectations back into our marginalized confidence equation from earlier:

$$
P(\text{correct} \mid a = 1, X_i, D_i) = \Phi\left(\frac{\left[(X_iD_i - \beta) + \frac{\sigma_e^2}{\sqrt{\sigma_e^2 + \sigma_c^2}} 

\cdot \lambda\left(\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)\right]


\cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)
$$

$$
P(\text{correct} \mid a = -1, X_i, D_i) = 1 - \Phi\left(\frac{\left[(X_iD_i - \beta) - \frac{\sigma_e^2}{\sqrt{\sigma_e^2 + \sigma_c^2}} 

\cdot \lambda\left(-\frac{X_iD_i - \beta}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)\right]


\cdot CX_i}{ \sigma_e^2 + \sigma_m^2} \cdot\frac{1} {\sqrt{1 + \frac{C^2X^2_i}{\sigma_e^2+\sigma_m^2}}}\right)
$$

with:

$$
\lambda(x) = \frac{\phi(x)}{\Phi(x)}
$$


where:
- $C = \frac{2}{1.7}$ (the scaling constant from the logit-normal approximation)
- $\lambda(x) = \frac{\phi(x)}{\Phi(x)}$ (inverse Mills ratio)

This gives us a fully marginalized model for confidence ratings that depends only on the parameters $\Theta = (\beta, \sigma_e, \sigma_c, \sigma_m)$ and the observed data (choices $a_i$, stimuli $X_i$, true states $D_i$), without requiring trial-by-trial estimation of latent evidence variables $e_i$ and $r_i$.
