<!-- Part 3: Design, Methods, and Hypotheses -->

# Part 3: Design, Methods, and Hypotheses

## Experimental study

The current study examines motion discrimination and meta-cognition in a Random Dot Motion (RDM) task. Specifically, we are interested in the question of whether post-decisional timing influences meta-cognitive uncertainty. We conceptualize this as whether confidence ratings follow the probability of being correct more closely when subjects are forced to wait longer before giving their confidence ratings, compared to when they can respond directly. 

Dynamic models of meta-cognition assume that evidence is accumulated until a response boundary is reached and a binary decision is initialized. After the binary choice is given, additional evidence is assumed to be accumulated (i.e. post-decisionally) in order to give a confidence rating to the binary choice. If this post-decisional evidence accumulation is informative about the correctness of the choice, one would expect that increasing waiting time should increase the quality of the confidence ratings (i.e. confidence ratings would be closer to the objective probability that the choice was correct). However, if evidence used for the confidence rating is already computed around the time of the decision (i.e. pre-decisional) one would expect there to be no-difference or even expect the opposite effect. The reverse effect would indicate a situation where there exists some leakage of the information (perhaps due to memory constraints). In this study we manipulate the post decisional time window to examine its effects on the quality of the confidence ratings.

Previous studies investigating post decisional evidence accumulation have generally found that if post-decisional evidence accumulation exists it is easiest to detect in speeded decision making tasks for easy trials, where increased post-decisional time can decrease confidence ratings for incorrect choices. In this first study we set out to test the opposite case, a non-speeded perceptual decision making task with a range of trials from very easy to very hard (individually calibrated). 

<!-- Ths s the mre severe test f the thery  -->

As the evidence is not clear and we do not necessarily expect to find a significant difference in how post-decisional timing changes meta-cognitive uncertainty, we set out to test this in a Bayesian framework using sequential sampling. This not only let us provide evidence for the NULL-hypothesis (post-decisional time does not change the quality of the confidence ratings), but also allows us to collect as many participants as strictly needed to provide sufficient evidence.

### Hypotheses

1. **Hypothesis 1:** Increasing the post-decisional time window will lead to no improvement or deterioration in the relationship between confidence and accuracy.

    - **Test specification:**  
      This hypothesis is tested by allowing the $\sigma_m$ parameter to vary by the trial-by-trial post-decisional time window. As this is the main parameter of interest, we will sequentially test whether this parameter has reached a a Bayesfator of either 30 or 1/30, which is our termination criteria. 

2. **Hypothesis 2:** Increasing the post-decisional time window will lead to no difference in confidence bias.

    - **Test specification:**  
      This hypothesis is tested by allowing the $\mu_\beta$ parameter to vary by the post-decisional time window. The primary test evaluates whether the group-level parameter differs from 0 (i.e. does the 95 Highest density interval include 0), here we also aim to report the bayesfactor to determine the strength of the evidence compared to the NULL (i.e. there is no difference). This test will be conducted when the data collection has been terminated.


### Priors and Sequential Sampling.

To limit the number of subjects needed to get sufficient evidence for our inference we will conduct a sequential sampling paradigm. Here we choose to terminate recruitment when the above explained Bayesfactor reaches either 30 or 1/30, or we have collected 50 subjects. We will start examining the Bayesfactors in increments of 5 subjects when we have reached 15 total subjects. This is mainly due to stability of the model estimation, as well as expecting to test around 5 subjects per day.

#### Priors

As our sequential sampling approach uses bayesfactors for the termination of data collection, the prior distribution of this parameter is highly influential. To maximize transparency we visualize the curves that we allow given the prior below, interested readers are also refereed to  [Sequential Sampling](../Sequential sampling/) directory, here we provide scripts for fitting visualizing and diagnosing our model.

```{r, fig.height=8, fig.width=12, warning=F,error=F,message=F}
source(here::here("Simulations","Heurestic","scripts","Simulate_mcmc.R"))

n_sims <- 50

df_list <- lapply(1:n_sims, function(i){
  df <- sim_trials(300, seed = i)
  df$sim_id <- i
  df
})

df_all <- do.call(rbind, df_list)

df_all %>% pivot_longer(cols = c("theta","c_mu")) %>% 
  mutate(ACC = as.factor(ifelse(name == "theta","Response",ACC))) %>% 
  group_by(ACC,name,X,sim_id) %>% 
  summarize(mean = mean(value),
            se = sd(value) / sqrt(n())) %>% 
  ggplot(aes(x = X, y = mean, ymin = mean-2*se, ymax = mean+2*se, col = ACC, group = interaction(ACC,sim_id)))+
  # geom_pointrange()+
  facet_wrap(~name,ncol = 1)+
  scale_color_manual(values = c("darkred","darkgreen","black"))+
  geom_smooth(se = F, alpha = 0.25)+
  theme_minimal()
```

The plot above shows the prior predictive checks of the group level parameters (i.e. these curves represent group level curves). As can be seen the curves generally span everything. Now in order to test whether our sequential sampling approach works we simulate data from draws of the posterior distribution of our pilot data, we then investigate how the bayesfactors of both hypotheses outlined above evolve over time. Instead of drawing the parameters of interest we fix these to either -0.05, 0 or 0.05 to test that we can detect effects of this magnitude in both direction, but also terminate sampling if there is no effect. Below we display the size of these effects (simulated data from the mean of the group means).

```{r, fig.height=8, fig.width=12, warning=F,error=F,message=F}
effects <- c(-0.05, 0, 0.05)

# Create all combinations (if you had two effects)
grid <- expand.grid(effect1 = effects, effect2 = effects)

df_list <- vector("list", nrow(grid))

for(i in 1:nrow(grid)){
  df_list[[i]] <- sim_trials_fix(300, 1997, grid$effect1[i], grid$effect2[i])
  df_list[[i]]$sigma_m_beta <- grid$effect1[i]
  df_list[[i]]$meta_bias_beta <- grid$effect2[i]
}

df_effects <- do.call(rbind, df_list)


df_effects %>% 
  # filter(ACC == 1) %>% 
  mutate(Accuracy = ACC) %>% 
  mutate(ITI = as.factor(cut(interval,3))) %>% 
  group_by(X,ITI,Accuracy,meta_bias_beta,sigma_m_beta) %>% 
  summarize(mean = mean(c_mu),
            se = sd(c_mu) / sqrt(n())) %>% 
  ggplot()+
  geom_hline(yintercept = 0.5, linetype = 2)+
  scale_color_manual(values = c("orange","orange3","darkred","green","green3","darkgreen"))+
  geom_pointrange(aes(x = X, y = mean, ymin = mean-2*se, ymax = mean+2*se, col = interaction(ITI,Accuracy)))+
  geom_smooth(aes(x = X, y = mean, col = interaction(ITI,Accuracy)),se = F)+
  facet_grid(sigma_m_beta~meta_bias_beta, labeller = label_both)+
  theme_minimal()


```

#### Sequential sampling

As mentioned above we simulated subjects from draws of the posterior that was informed by our pilot data. We then fitted the model to increasing steps (i.e. 5,7,9, 12,15,20,25) of subjects, to demonstrate that the sequential sampling approach would be feasible. below we showcase 5 such trajectories of Bayes factors across the number of subjects used, together with another panel showing how the posterior distribution of the parameter of interest changes with increasing subjects.


```{r, fig.height=8, fig.width=12, warning=F,error=F,message=F}
magick::image_read(here::here("Preregistration","Figures","sequential_samping_plot.PNG"))
```
In the above plot we show the development of the effect of the interval between choice and confidence rating on both meta cognitive bias (green) and meta cognitive uncertainty (orange). Circles display a fitted model without divergences, above 400 effective sample size for both the bulk and the tails and rhat values below 1.03. The Left side displays the posterior distribution of each parameter for increasing subjects. The right side displays the development of the Bayes-factors for increasing subjects for the same draw of the posterior distribution fitted to our pilot data.


From the above plot it can be seen that even with this model we might run into convergences issues, below we describe our approach for determining and dealing with convergences issues. 

Firstly, we will collect data from 15 subjects before we fit our hierarchical model the first time and test the size of the bayes-fator. We do this as hierarchical models generally need some initial data to be able to converge (see also plot above for stimulation). We will then after each additional 5 subjects check the infered bayes factor as this coincides with the expected number of subjects being collected each day.

Below we define our initial sampler parameters, then what we deem sufficiently good sampler diagnostics, and lastly contingency plans if we do not get convergence.


##### Initial Sampler Configuration

The model is initially estimated using the following sampler settings:

- Number of chains: 4  
- Number of iterations per chain: 1000  
- Warmup iterations: 1000 
- Target acceptance rate (`adapt_delta`): 0.95  
- Maximum tree depth (`max_treedepth`): 12

These values represent our baseline configuration for model estimation.

##### Convergence Criteria

Sampler convergence and reliability will be assessed using the following diagnostics:

- $\hat{R}$ values below 1.03 for all group level parameters  
- Effective sample size (ESS) above 400 for all group level parameters  
- Zero divergent transitions after warmup  
- No saturation of the maximum tree depth  
- well-mixing trace-plots

If these criteria are met, the sampling procedure will be considered satisfactory.

##### Contingency Procedures

If the above convergence criteria are not met, we will apply the following steps depending on the convergence failure:

1. Increase the number of iterations.
2. Increase the target acceptance rate (`adapt_delta`).
3. Increase the maximum tree depth (`max_treedepth`).
5. Reconsider model parameterization (i.e. centered vs non-centered parameterization).
6. Estimate the lapse-rate parameter non-hierarchically

These steps are intended to improve sampler stability and ensure reliable posterior estimation.

Interested readers are referred to the directory used to perform the [Sequential Sampling](../Sequential sampling/). Here the full Stan model and the scripts for checking diagnostics and model fit are present. 


## Participants 

Eligible participants will be 18–40 years old, have normal or corrected‑to‑normal vision, and be proficient in English. Individuals with neurological or psychiatric diagnoses or those taking medication that may affect cognitive performance are also not eligible to participate. Subjects will be recruited through COBE Lab its participant pool to insure fast and effective recruitment and will be rewarded with 110DKK.


## Design

This study consists of a single lab visit where participants perform a computer experiment on a COMPUTERSPEC(XXX) run using Psychopy [@PEIRCEPsychoPy2ExperimentsBehaviorMadeEasy2019]. The experiment uses a random dot motion (RDM) decision‑making task. On each trial, participants view a random dot kinematogram in which a majority of dots move either upward or downward. While the dots are displayed, participants indicate whether the overall motion is ‘up’ or ‘down’ by pressing a corresponding key. They have 2 seconds to respond, trials without a response within this window are classified as no‑response trials. After indicating their response, participants view a post‑decisional fixation cross. The duration of this interval is the main experimental manipulation and is randomly drawn from a uniform distribution ranging from the maximum response time of 2 seconds to 5 seconds ($PostD_i \sim \text{U}(2, 5)$). After the interval, participants provide a metacognitive evaluation of their decision. They are asked, ‘How confident were you in your decision?’ and respond using a scale ranging from ‘certainly wrong’ to ‘certainly right.’ Participants navigate the scale using the arrow keys and confirm their response with a key press. If no response is made within 3 seconds, the trial is classified as a no‑response trial. Before the start of each new trial, an inter‑trial interval is presented consisting of a fixation cross. The duration of this interval is drawn from a log‑normal distribution with a mean of 1 seconds and a standard deviation of 0.1 ($ITI_i \sim \text{Lognormal}(1,0.1)$).


```{r, echo=FALSE} 
n_trials = 36
n_blocks = 8
n_total = n_trials * n_blocks
```

In total, participants will complete `r n_total` trials, spread over `r n_blocks` blocks of `r n_trials` trials. Before the main experiment begins, participants receive instructions on how to perform the decision task. They then complete six easy, non‑speeded practice trials without confidence ratings, followed by 120 more difficult practice trials. Afterward, participants receive additional instructions about the confidence scale and complete 10 easy, non‑speeded practice trials that include confidence ratings, after which the main experiment begins. The key variables of interest are response choice, accuracy, response times, confidence ratings, and the response times associated with those confidence ratings.

To ensure that task difficulty is calibrated at the individual level, the 120 harder practice trials are used as part of a staircase procedure. In a random dot kinematogram, task difficulty is determined by several stimulus properties, which can be broadly grouped into two categories: stimulus parameters and stimulus noisiness. To ensure that difficulty is optimally calibrated for each participant, we use the first 40 of the harder practice trials to staircase the dotlife parameter in the Dotstimclass in [psychopy](https://www.psychopy.org/api/visual/dotstim.html). This is a stimulus parameter that determines how many frames a dot lives before disappearing. Once an appropriate dot‑life value is identified, we start staircasing coherence. Coherence controls the stimulus signal-to-noise ratio and is the most commonly manipulated difficulty parameter in the literature. It ranges from 0 to 1 and determines the proportion of signal to noise dots, with higher values indicating a greater percentage of signal dots.

The remaining 80 trials are used to estimate the slope and threshold of the participant’s psychometric function on coherence. Both staircases are implemented in PsychoPy using the QUEST+ Bayesian adaptive procedure, which selects each stimulus to minimize posterior entropy over the psychometric parameters (Watson, 2017; implementation: https://www.psychopy.org/api/data.html#psychopy.data.QuestPlusHandler). Using the estimated psychometric function, we determine the coherence levels for the main experiment based on a set of predefined target accuracies  ($ACC = [0.51, 0.6, 0.7, 0.8, 0.99]$). The corresponding coherence values are then sampled in a quasi‑normal fashion: the extreme accuracy levels (0.51 and 0.99) appear once per direction, the intermediate levels appear twice, and the 0.70 level appears three times. In each block of the main experiment, trials are presented in a pseudo‑randomized sequence of upward and downward motion such that no more than three consecutive trials share the same direction. This reduces the risk of implicit directional bias. All of this ensures that participants use the full confidence scale and that difficulty is balanced across upward and downward motion directions.

## Exclusion criteria

```{r, echo=FALSE} 
ACC_cut = 60
no_resp_cut = 30
```

Because the task is a two‑alternative forced‑choice decision, chance performance is 50%. For each participant, we will compute overall accuracy across all main‑task trials and will exclude all participants with an accuracy lower than `r ACC_cut`%. For our trial count, accuracy below this threshold is not reliably above chance under a binomial model of guessing. Furthermore, given the stimulus calibration procedure, an optimal observer is expected to achieve approximately 71% accuracy.

Taken together, an accuracy below `r ACC_cut`% is inconsistent with meaningful participation in the task. We will also exclude participants who have a no‑response rate greater than `r no_resp_cut`% (combining decision and confidence responses), as a high proportion of missed responses also indicates insufficient engagement with the task.


## Physiological measures

As an additional stream of data we also collect physiological measures from each of our subjects. Specifically, we will be collecting eyetracking (and pupil size) using an Eyelink 1000, furthermore we will collect electrocardiogram (ECG) and respiration measures using Biopac Systems. To ensure accurate eye-tracking calibration, the experimental script performs recalibration after each experimental block, which also provides participants with regular breaks.

These physiological measures will allow us to test the theoretical assumptions of our model of confidence using physiological data. In particular, we will investigate whether autonomic responses, especially changes in pupil size, reflect the agent’s appraisal of their own action. We hypothesize that an initial physiological response will occur following stimulus presentation. Crucially, we further predict a second autonomic response shortly after the initial decision response. This response is expected to unfold within a time window of approximately 500–2500 ms after the initial response, corresponding to the period during which the action is being appraised (Note the general sluggishness of the autonomic signals making this range wide). Importantly, this physiological response is expected to occur before the explicit confidence rating, suggesting that the appraisal process, and computation of confidence, might begin prior to the overt response. Additionally we expect the initial response to the stimulus to be conditional on the stimulus intensity, such that higher coherence leads to larger amplitudes of the autonomic response. We further believe that this second response to the appraisal of the action will be conditional on the prediction error of the comparator model. Practially, we expect that the physiological signature of action appraisal, scales with the difference between the action and the expected probability of that action. Meaning that an action that would be produced in 90% of instances would lead to a small prediction error, whereas the complement action would generate a high prediction error as this would only be produced in 10% of instances of the given stimulus.

### Exploratory analyses

1. **Ex-Hypothesis 1:** Autonomic responses, particularly pupil size, will show an initial response to the stimulus, but also a second response after the decision, reflecting action appraisal.


2. **Ex-Hypothesis 2:** The Autonomic responses for the initial response to the stimulus will scale with the stimulus intensity, whereas the second response will scale with the prediction error of the action and the expected value.


As the analysis plan of these two hypotheses were not fully developed before data collect we highlight them as exploratory.
