<!-- Part 3: Design, Methods, and Hypotheses -->

# Part 3: Design, Methods, and Hypotheses

## Experimental study

The current study examines motion discrimination and meta-cognition in a Random Dot Motion (RDM) task. Specifically, we are interested in the question of whether the time between first order choice and confidence judgements (post-decisional time) influences meta-cognitive uncertainty. We conceptualize this as whether confidence ratings follow the probability of being correct more closely when subjects are forced to wait longer before giving their confidence ratings, compared to when they can respond directly. 

Dynamic models of meta-cognition assume that evidence is accumulated until a response boundary is reached and a binary decision is initialized. After the binary choice is given, additional evidence is assumed to be accumulated (i.e. post-decisionally) in order to give a confidence rating to the binary choice. If this post-decisional evidence accumulation is informative about the correctness of the choice, one would expect that increasing waiting time should increase the quality of the confidence ratings. Meaning that confidence ratings would be closer to the objective probability that the choice was correct. However, if evidence used for the confidence rating is already fully computed around the time of the decision (i.e. pre-decisional) one would expect there to be no-difference or even expect the opposite effect. If confidence ratings become less consistent with the probability of the choice being correct  this would indicate a situation where some leakage of the information is taking place (perhaps due to memory constraints). In this study we manipulate the post-decisional time window to examine its effects on the quality of the confidence ratings. Previous studies investigating post-decisional evidence accumulation have generally found that if post-decisional evidence accumulation exists it is easiest to detect in decision making tasks where speed is emphasized particularly for easy trials, here increased post-decisional time can decrease confidence ratings for incorrect choices. 

In this first study we set out to test a non-speeded perceptual decision making task with a range of trials from very easy to very hard (individually calibrated). Here we test the theory that post-decisional evidence accumulation might unfold over a longer time period even in non-speeded perceptual decision making tasks. As previous literature has shown inconclusive evidence of whether post-decisional evidence accumulation can matter in non-speeded decision making tasks, we set out to test this in a Bayesian framework using sequential sampling. This not only let us provide evidence for the null-hypothesis, post-decisional time does not change the quality of the confidence ratings, but also allows us to only collect the necessary number of participants.

### Computational model

To test whether the post-decisional response window changes how informative confidence ratings are about choice correctness, we constructed a hierarchical Bayesian model grounded in principles from signal detection theory and motor control theory. In brief, the model assumes that agents exhibit variability in the encoding of the physical stimulus, as in traditional signal detection theory, but also variability in the selection of a choice given the internal evidence. Consistent with motor control theories, the model further assumes that an efference copy is generated before the efferent signal is sent to the acting limb. This efference copy is assumed to be corrupted by additional noise, consistent with other generative metacognition models, before being combined with the executed action to compute confidence in the chosen response. The model naturally accounts for error monitoring through the inclusion of choice variability. For the full derivation and simulations of the model, see [the accompanying model markdown](Current model.Rmd). Here it suffices to note that the metacognitive uncertainty parameter $\sigma_m$ governs the relationship between confidence ratings and the correctness of the choice: as $\sigma_m$ increases, confidence becomes less informative about whether the choice was correct.


### Hypotheses

1. **Hypothesis 1:** Increasing the post-decisional time window will lead to no improvement in the relationship between confidence and accuracy.

    - **Test specification:**  
      This hypothesis is tested by allowing the meta-certainty ($\sigma_m$) parameter to vary linearly on the logarithmic-scale with post-decisional time window. As this is the main parameter of interest, we will sequentially test whether this parameter has reached a Bayes factor of either 30 or 1/30, which is our termination criteria. 

2. **Hypothesis 2:** Increasing the post-decisional time window will lead to no difference in confidence bias.

    - **Test specification:**  
      This hypothesis is tested by allowing the $mu_{\beta_c}$ parameter to vary by post-decisional time window. The primary test evaluates whether the group-level parameter differs from 0 (i.e. does the 95 Highest density interval include 0), here we also aim to report the Bayes factor to determine the strength of the evidence compared to the NULL (i.e. there is no difference). This test will be conducted when the data collection has been terminated.

### Priors and Sequential Sampling.

To limit the number of subjects needed to get sufficient evidence for our inference we will conduct a sequential sampling paradigm. Here we choose to terminate recruitment when the above explained Bayes factor reaches either 30 or 1/30, or we have collected 50 subjects. We will start examining the Bayes factors in increments of 5 subjects when we have reached 15 total subjects that are non-excluded (see exclusion citeria). This is mainly due to stability of the model estimation, as well as expecting to test around 5 subjects per day. The Bayes factor used in the current study is the savage dickey Bayes factor evaluated at 0 [@WAGENMAKERSBayesianHypothesisTestingPsychologistsTutorialSavage2010a].


Below we define our initial sampler parameters, then what we deem sufficiently good sampler diagnostics, and lastly contingency plans if we do not get convergence.


#### Initial Sampler Configuration

The model is initially estimated using the following sampler settings:

- Number of chains: 4  
- Number of iterations per chain: 1000  
- Warmup iterations: 1000 
- Target acceptance rate (`adapt_delta`): 0.95  
- Maximum tree depth (`max_treedepth`): 12

These values represent our baseline configuration for model estimation.

##### Convergence Criteria

Sampler convergence and reliability will be assessed using the following diagnostics:

- $\hat{R}$ values below 1.03 for all group level parameters  
- Effective sample size (ESS) above 400 for all group level parameters  
- Zero divergent transitions after warmup  
- No saturation of the maximum tree depth  
- well-mixing trace-plots

If these criteria are met, the sampling procedure will be considered satisfactory.

##### Contingency Procedures

If the above convergence criteria are not met, we will apply the following steps depending on the convergence failure:

1. Increase the number of iterations.
2. Increase the target acceptance rate (`adapt_delta`).
3. Increase the maximum tree depth (`max_treedepth`).
4. Reconsider model parameterization (i.e. centered vs non-centered parameterization).
5. Estimate the lapse-rate parameter non-hierarchically
6. Estimate the choice variability parameter non-hierarchically


These steps are intended to improve sampler stability and ensure reliable posterior estimation.

Interested readers are referred to the directory used to perform the Sequential Sampling (../Sequential Sampling). Here the full Stan model and the scripts for checking diagnostics and model fit are present. 


## Participants 

Eligible participants will be 18–40 years old, have normal or corrected‑to‑normal vision, and be proficient in English. Individuals with neurological or psychiatric diagnoses or those taking medication that may affect cognitive performance are also not eligible to participate. Subjects will be recruited through COBE Lab's participant pool to insure fast and effective recruitment and will be rewarded with 110DKK.


## Design

This study consists of a single lab visit of an hour, where participants perform a computer experiment on a COMPUTERSPEC(XXX) run using Psychopy [@PEIRCEPsychoPy2ExperimentsBehaviorMadeEasy2019]. The experiment uses a random dot motion (RDM) decision‑making task. On each trial, participants view a random dot kinematogram in which a majority of dots move either upward or downward. While the dots are displayed, participants indicate whether the overall motion is ‘up’ or ‘down’ by pressing a corresponding key. The dots stayed on the screen for at most 2 seconds or when the participant responded, trials without a response within this window are classified as no‑response trials. After indicating their response, participants view a post‑decisional fixation cross. The duration of this interval is the main experimental manipulation and is randomly drawn from a uniform distribution ranging from 2 to 5 seconds ($PostD_i \sim \text{U}(2, 5)$). After the interval, participants provide a metacognitive evaluation of their decision. They are asked, ‘How confident were you in your decision?’ and respond using a visual analog scale ranging from ‘certainly wrong’ to ‘certainly right.’ Participants navigate the scale using the arrow keys and confirm their response with a key press. If no response is made within 3 seconds, the trial is classified as a no‑response trial. Before the start of each new trial, an inter‑trial interval is presented consisting of a fixation cross. The duration of this interval is drawn from a log‑normal distribution with a mean of 1 seconds and a standard deviation of 0.1 ($ITI_i \sim \text{Lognormal}(1,0.1)$).


```{r, echo=FALSE} 
n_trials = 36
n_blocks = 8
n_total = n_trials * n_blocks
```

In total, participants will complete `r n_total` trials, spread over `r n_blocks` blocks of `r n_trials` trials, which from our pilot data has shown good parameter estimation for subject-level-parameters. Before the main experiment begins, participants receive instructions on how to perform the decision task. They then complete six easy, non‑speeded practice trials without confidence ratings, followed by 120 more difficult practice trials. Afterward, participants receive additional instructions about the confidence scale and complete 10 easy, non‑speeded practice trials that include confidence ratings, after which the main experiment begins. The key variables of interest are response choice, accuracy, response times, confidence ratings, and the response times associated with those confidence ratings.

To ensure that task difficulty is calibrated at the individual level, the 120 harder practice trials are used as part of a staircase procedure. In a random dot kinematogram, task difficulty is determined by several stimulus properties, which can be broadly grouped into two categories: stimulus parameters and stimulus noisiness. To ensure that difficulty is optimally calibrated for each participant, we use the first 40 of the harder practice trials to staircase the dotlife parameter in the Dotstimclass in [psychopy](https://www.psychopy.org/api/visual/dotstim.html). This is a stimulus parameter that determines how many frames a dot lives before disappearing. Once an appropriate dot‑life value is identified, we start staircasing coherence. Coherence controls the stimulus signal-to-noise ratio and is the most commonly manipulated difficulty parameter in the literature. It ranges from 0 to 1 and determines the proportion of signal to noise dots, with higher values indicating a greater percentage of signal dots.

The remaining 80 trials are used to estimate the slope and threshold of the participant’s psychometric function on coherence. Both staircases are implemented in PsychoPy using the QUEST+ Bayesian adaptive procedure, which selects each stimulus to minimize posterior entropy over the psychometric parameters [@WATSONQUESTGeneralMultidimensionalBayesianAdaptivePsychometric2017]. Using the estimated psychometric function, we determine the coherence levels for the main experiment based on a set of predefined target accuracies  ($ACC = [0.51, 0.6, 0.7, 0.8, 0.99]$). The corresponding coherence values are then sampled in a quasi‑normal fashion: the extreme accuracy levels (0.51 and 0.99) appear once per direction, the intermediate levels appear twice, and the 0.70 level appears three times. In each block of the main experiment, trials are presented in a pseudo‑randomized sequence of upward and downward motion such that no more than three consecutive trials share the same direction. This reduces the risk of implicit directional bias. All of this ensures that participants use the full confidence scale and that difficulty is balanced across upward and downward motion directions.

## Exclusion criteria

```{r, echo=FALSE} 
ACC_cut = 60
no_resp_cut = 30
min_RT = 150
```

Because the task is a two‑alternative forced‑choice decision, chance performance is 50%. For each participant, we will compute overall accuracy across all main‑task trials and will exclude all participants with an accuracy lower than `r ACC_cut`%. For our trial count, accuracy below this threshold is not reliably above chance under a binomial model of guessing. Furthermore, given the stimulus calibration procedure, an optimal observer is expected to achieve approximately 71% accuracy.

Taken together, an accuracy below `r ACC_cut`% is inconsistent with meaningful participation in the task. Additionally, trials with a response time below `r min_RT` ms will be coded as no response trials and excluded from further analysis, as responses faster than this are unlikely to reflect genuine perceptual processing. We will exclude participants who have a no‑response rate greater than `r no_resp_cut`% (combining decision, confidence responses and response times below `r min_RT` ms), as a high proportion of missed responses also indicates insufficient engagement with the task.


## Physiological measures

As an additional stream of data we also collect physiological measures from each of our subjects. Specifically, we will be collecting eyetracking (and pupil size) using an Eyelink 1000, furthermore we will collect electrocardiogram (ECG) and respiration measures using Biopac Systems [@DataAcquisitionLoggersAmplifiersTransducersElectrodes]. To ensure accurate eye-tracking calibration, the experimental script performs re-calibration after each experimental block, which also provides participants with regular breaks.

These physiological measures will allow us to test the theoretical assumptions of our model of confidence ratings using physiological data. In particular, we will investigate whether autonomic responses, reflect the agent’s appraisal of their own action. We expect that an initial physiological response will occur following stimulus presentation i.e. the orienting response [@GABAYOrientingAttentionPupilSizeNorepinephrineSystem2011]. Crucially, we predict a second response of the physiological measures, shortly after the initial decision. This physiological response is expected to unfold within a time window of approximately 500–2500 ms after the initial response, corresponding to the period during which the action is being appraised (Note the general sluggishness of the autonomic signals making this range wide). Importantly, this physiological response is expected to occur before the explicit presentation of the confidence rating, which would indicate that the appraisal process and computation of confidence begin prior to the overt response of confidence. Additionally, we expect the initial response to the stimulus to be conditional on the stimulus intensity, such that higher coherence leads to larger amplitudes of the physiological responses. We further believe that the second physiological response, the appraisal of the action, will be conditional on the internal prediction error between the expected evidence distribution and the observed action, quantified by the internal prediction error $S_i$. This quantity is largest for errors at high stimulus intensity and decreases toward threshold, while for correct responses it is largest near threshold and decays to zero at high stimulus intensity. These predictions are directly derivable from the model and generate two dissociable physiological signatures: a stimulus-locked response scaling with |XD| which should not be conditional on the action nor the choice accuracy, and an action-locked response scaling with $S_i$ which diverges by accuracy as stimulus intensity increases. 


### Physiological analyses

1. **Physiological-Hypothesis 1:** Autonomic responses, particularly pupil size, will show an initial response to the stimulus, but also a second response after the decision, reflecting action appraisal.


2. **Physiological-Hypothesis 2:** The Autonomic responses for the initial response to the stimulus will scale with the stimulus intensity, whereas the second response will scale with the internal prediction error $S_i$. Which predicts the greatest response for high stimulus intensity incorrect choices, an intermediate response near threshold regardless of accuracy, and the smallest response for high stimulus intensity correct choices.


As the analysis plan of these two hypotheses are not fully developed, and do not depend on the sequential sampling, we highlight them as exploratory.
