# Part 4: Sequential Sampling Data Collection Progress

```{r load packages, warning = F, message = F, echo = F, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dpi = 300,
  fig.align = 'center'
)

set.seed(123)


if (!require("pacman")) install.packages("pacman")
pacman::p_load("here", "knitr","yaml","cowplot", "magick",
               "tidyverse", "flextable","ordbetareg","patchwork","truncnorm")



files = list.files(here::here("Results","SS"),
                   full.names = TRUE)

# 
# files = list.files(here::here("Preregistration","Results","SS"),
#                    full.names = TRUE)




big_list = list()

for(file in files){
  big_list = append(big_list, list(readRDS(file)))
}
```


## Overview

This section provides live updates on the sequential sampling experiment data collection. As participants complete the experiment.

## Data Collection Status

At the moment there has been excluded `r big_list[[length(big_list)]][[1]]`, due to poor accuracy, and `r big_list[[length(big_list)]][[2]]`, due to no-response rates above. This in total means we have excluded `r big_list[[length(big_list)]][[3]]`.


## Model convergence and Sequential sampling

```{r, fig.height=6, fig.width=10,warning=F,message=F}
big_list[[length(big_list)]][[4]]
```



```{r, fig.height=4, fig.width=7,warning=F,message=F}
df = do.call(rbind, lapply(big_list, function(x) x[[5]]))


df %>% 
  ggplot()+
  geom_line(aes(x = subjects, y =log(BF)),col = "orange")+
  geom_point(aes(x = subjects, y = log(BF)),col = "orange")+
  labs(y = "Log(BF)", x = "N_subjects")+
  geom_hline(yintercept = c(log(1/30),log(30)), linetype = 2)+
  scale_y_continuous(breaks = c(round(log(1/30),2),0,round(log(30),2)))+
  scale_x_continuous(limits = c(0,30), breaks = seq(0,30,by = 10), labels = seq(0,30,by = 10))+
  theme_classic()+
  theme(legend.position = "top")

```

