# Part 4: Sequential Sampling Data Collection Progress

```{r load packages, warning = F, message = F, echo = F, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dpi = 300,
  fig.align = 'center'
)

set.seed(123)


if (!require("pacman")) install.packages("pacman")
pacman::p_load("here", "knitr","yaml","cowplot", "magick",
               "tidyverse", "flextable","ordbetareg","patchwork","truncnorm")


files = list.files(here::here("Preregistration","Results","SS"),
                   full.names = TRUE)

big_list = list()

for(file in files){
  big_list = append(big_list, list(readRDS(file)))
}
```


## Overview

This section provides live updates on the sequential sampling experiment data collection. As participants complete the experiment.

## Data Collection Status

At the moment there has been 


## Participant Progress

```{r, message = FALSE, warning = FALSE, echo = FALSE}

# Placeholder: Plot of participant completion status
# Can include: enrollment timeline, trial completion per participant, data quality metrics

```

## Model Fits

```{r, message = FALSE, warning = FALSE, echo = FALSE}

# Placeholder: Model fit visualizations from Run_SS
# Examples:
# - Posterior distributions of key parameters
# - Prior vs posterior comparisons
# - Convergence diagnostics (Rhat, n_eff)
# - Parameter recovery (if simulations run)

```

## Preliminary Analyses

```{r, message = FALSE, warning = FALSE, echo = FALSE}

# Placeholder: Preliminary result plots
# Examples:
# - Confidence vs accuracy relationships
# - Post-decisional window effects (if sufficient data)
# - Between-subject parameter heterogeneity

```

## Data Access

The raw sequential sampling data and full fitting results are stored in the [`Sequential sampling/`](../../Sequential sampling/) directory and remain private. Fitted models and diagnostic outputs are available there for reproducibility.
