---
title: "Metacognitive uncertainty in Correct and Error trials across inter-stimulus-intervals"
author: Jesper Fischer Ehmsen^1^,
        Siebe Everaets^xx^,
        Micah G. Allen^1,2,\*^,
        Francesca Fardo^1,3,\*^
output:
  html_document: default
  word_document:
    # reference_docx: Knitting files/docx_template.docx
  pdf_document:
    latex_engine: xelatex
# bibliography: 
  # - "Knitting files/References.bib"
link-citations: yes
linkcolor: blue
# csl: Knitting files/science-advances.csl
always_allow_html: yes
---

::: {custom-style="Title text"}
^1^Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University

^2^Cambridge Psychiatry, University of Cambridge, United Kingdom

^3^Danish Pain Research Center, Department of Clinical Medicine, Aarhus University

^\*^Equal contributions

Corresponding author: Jesper Fischer Ehmsen (jesperfischer@cfin.au.dk)

:::

```{r load packages, warning = F, message = F, echo = F, include=FALSE}
knitr::opts_chunk$set(dpi=300)

# seed
set.seed(123)

# packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load("here", "knitr","yaml")

knitr::opts_chunk$set(echo = TRUE)

#renv::restore(project=here::here())

required_packages = c("cowplot", "magick", "tidyverse", "flextable")
lapply(required_packages, library, character.only = TRUE)
```


This pre-registration consists of 3 parts; 
1. **considerations of models and empirical data of perceptual decision making tasks with confidence ratings**
2. **Outlining two computational models**
3. **Design, methods, hypotheses and sample size considerations of the current study.**

These three parts are covered in 3 separate documents, with this document including them all.

```{r child1, child="Models and Data in Perceptual Decision-Making with Confidence.Rmd"}
```

```{r child2, child="Current model.Rmd"}
```

```{r child3, child="Design methods hypotheses.Rmd"}
```





# References

::: {#refs}
:::