---
title: "Metacognitive uncertainty in Correct and Error trials across inter-stimulus-intervals"
author: Jesper Fischer Ehmsen^1^,
        Siebe Everaets^xx^,
        Micah G. Allen^1,2,\*^,
        Francesca Fardo^1,3,\*^
output:
  html_document: default
  word_document:
    # reference_docx: Knitting files/docx_template.docx
  pdf_document:
    latex_engine: xelatex
# bibliography: 
  # - "Knitting files/References.bib"
link-citations: yes
linkcolor: blue
# csl: Knitting files/science-advances.csl
always_allow_html: yes
---

::: {custom-style="Title text"}
^1^Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University

^2^Cambridge Psychiatry, University of Cambridge, United Kingdom

^3^Danish Pain Research Center, Department of Clinical Medicine, Aarhus University

^\*^Equal contributions

Corresponding author: Jesper Fischer Ehmsen (jesperfischer@cfin.au.dk)

:::

```{r load packages, warning = F, message = F, echo = F, include=FALSE}
knitr::opts_chunk$set(dpi=300)

# seed
set.seed(123)

# packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load("here", "knitr","yaml")

knitr::opts_chunk$set(echo = TRUE)

#renv::restore(project=here::here())

required_packages = c("cowplot", "magick", "tidyverse", "flextable")
lapply(required_packages, library, character.only = TRUE)
```


# Abstract



# Introduction

Metacognition, the ability to reflect on one's own cognitive processes, is commonly studied through confidence ratings in decision-making tasks. Three key findings have emerged from this research:

First, confidence ratings can be thought to reflect a noisy, subjective probability of being correct, incorporating response biases rather than purely objective performance.

Second, confidence judgments most often underperform relative to first-order accuracy, (i.e. a difficultly level produces 90% Accuracy might produce 80% confidence).

Third, metacognition and error-monitoring are functionally related systems. Being confident about an error is a form of error-monitoring. 


## Models of metacognition

Computational models of metacognition typically fall into two categories: **static models** based on Signal Detection Theory (SDT) and **dynamic models** based on evidence accumulation. 

Static models describe choice and confidence, while dynamic models additionally account for response times. 
Both can be understood as (approximately) optimal solutions to partially observable Markov decision processes under different assumptions. 
One key distinction between these two approaches is how evidence is encoded: 
static models assume fixed evidence per trial  (drawn from a distribution), while dynamic models assume evidence accumulates over time until some decision bound is reached.

The two types of models are not mutually exclusive and are many times thought of being extentions of oneanother. I.e. the dynamic models being extentions of the static models.

### Patterns of metacognition

Metacognitive models should capture several empirical patterns:

1. **Performance increases with stimulus strength**: Stronger stimuli yield higher accuracy
2. **Confidence increases with stimulus strength**: Stronger stimuli yield higher Confidence ratings
3. **Confidence varies with accuracy**: The relationship between confidence and stimulus strength depends on response correctness, with correct trials showing higher confidence than incorrect.
4. **"Folded-X" pattern**: Confidence increases with stimulus strength for correct responses but decreases (or plateaus) for errors—considered a signature of metacognition
5. **"Double-increase" pattern**: Confidence increases with stimulus strength for both correct and incorrect responses

Recent work suggests these two latter patterns of confidence arise from different experimental manipulations: 
* folded-X emerges when manipulating decision-relevant stimulus properties (e.g., coherence in RDM)
* double-increase occurs when manipulating auxiliary difficulty markers (e.g., stimulus visibility). 

From an error-monitoring perspective, folded-X reflects functioning error-monitoring (confidence decreases for errors), while double-increase suggests limited error-monitoring.

## Confidence scales, Stimulus granularity, and measurement considerations

Several methodological issues deserve attention:

**Confidence measurement**: Confidence is collected either discretely or continuously. While continuous scales offer more information and statistical power, 
most computational models require discrete data, leading many to discretize continuous confidence.

**Stimulus strength**: Similarly often binary or discretized, though continuous measurement provides more statistical power.

**Scale construction**: A critical yet underexamined issue is how confidence scale anchors constrain expressivity.
A **"half-scale"** (e.g., "guessing" to "confident") prevents expressing certainty about errors, agents cannot indicate they are sure they made a mistake. This produces a **"half-folded-X"** (flat incorrect branch) rather than a true folded-X. 
In contrast, a **"full scale"** (e.g., "confidently incorrect" to "confidently correct") enables full error-monitoring expression and can produce complete folded-X patterns. 
Simultaneous binary-confidence paradigms, where choice and confidence are indicated together, necessarily produce double-increase patterns, as otherwise would mean that agents deliberately choose the wrong answer.

**Choice variability and error-monitoring**: A key question is why agents make errors while knowing they erred. 
We propose this at least partly arises from variability in the choice process itself. Most models of perceptual decision making assume that after evidence has been acquired, a deterministic choice is given, i.e. the choice is fixed given the evidence. 
This means evidence noise and choice variability are confounded when observing only binary responses. 
However, by incorporating confidence ratings (and potentially response times), we can disentangle these noise sources.
We demonstrate that our model, which includes choice variability, can produce the full range of patterns (folded-X to double-increase) and represents a specific instantiation of Fleming and Daw's second-order framework.

### Static models of metacognition

Here we outline the general idea between the static models of perceptual deicison making and confidence judgements.

In SDT, an actor makes a binary choice $a \in \{-1,+1\}$ based on evidence $e$ about the true state $D \in \{-1,+1\}$. Evidence is drawn from a normal distribution:

$$
e_i \sim \mathcal{N}(X_iD_i,\sigma)
$$

where $X_i$ is stimulus strength. The actor compares evidence to criterion $c$:

$$
a_i = \begin{cases}+1 & \text{if } e_i > c \\ -1 & \text{if } e_i \le c\end{cases}
$$

For metacognition, the agent accesses this evidence for information about the correctness of the action.

Models differ in whether the rater has independent information above the evidence used to make the action.
Fleming and Daw formalized a second order rater where evidence for the action and evidence for the confidence judgement is draw from a multivariate normal with correlation $\rho$ governing information sharing. I.e. if $\rho$ = 1, the confidence judgement is based on the same information as the action, however if $\rho$ = 0 the confidence judgement is based on completely independent information.

$$
\begin{bmatrix}e_i \\ r_i\end{bmatrix}\sim \mathcal{N}\left(\begin{bmatrix}X_iD_i \\ X_iD_i\end{bmatrix},\begin{bmatrix}\sigma_e^2 & \rho \cdot \sigma_e \cdot \sigma_m \\ \rho \cdot \sigma_e \cdot \sigma_m & \sigma_m^2\end{bmatrix}\right)
$$

The rater uses this information to compute $P(D = a | r_i)$, the subjective probability of being correct.

One of the main arguments for the development of this and other models with independent information for the confidence judgement was to explain error-monitoring. The question is; how can an agent be confident that they made the wrong decision if they do not have access to independent information from the action. Phrased different, why do agents sometime make the wrong choice eventhough they know that it was the wrong choice?

Below we formalize the idea that such independent information is not necessary if instead it is assumeed that agents have varability in their choice given the evidence. I.e. instead of a deterministic decision policy agents have a noisy decision process that can produce errors eventhough the evidence is in favor of a different choice.

## Choice Variability

We extend the standard SDT framework by introducing stochastic choice policy. Evidence is drawn as:

$$
e_i \sim \mathcal{N}(X_iD_i,\sigma_e)
$$

But now the criterion varies trial-by-trial: $c_i \sim \mathcal{N}(0,\sigma_c)$. This yields:

$$
P(a = 1 \mid D) = \Phi\left(\frac{D_i \cdot X_i}{\sqrt{\sigma_e^2 + \sigma_c^2}}\right)
$$
From choices alone, only $\sigma_e^2 + \sigma_c^2$ is identifiable—both noise sources equivalently flatten the psychometric curve. However, by including confidence ratings these can be disentangled 

For confidence, we define:

$$\text{confidence} = P(\text{correct} \mid r, a) = P(D = a \mid r)$$

The rater accesses noisy evidence: $r_i \sim \mathcal{N}(e_i, \sigma_m^2)$, where $\sigma_m^2$ is metacognitive noise. This can be shown to yield:

$$P(D=1 \mid r_i) = g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right)$$

giving:

$$
P(\text{correct} \mid r_i, a) = \begin{cases}g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = 1 \\ 1 - g\left(\frac{2 \cdot r_i \cdot \mu_i}{\sigma_e^2 + \sigma_m^2}\right) & \text{if } a = -1\end{cases}
$$

From confidence alone, only $\sigma_e^2 + \sigma_m^2$ is identifiable. However, combining choices and confidence allows estimating all three parameters: $\sigma_e$, $\sigma_c$, and $\sigma_m$. This requires Bayesian data augmentation to estimate latent trial-by-trial evidence, has this normative probability of being correct i.e. $P(correct | r_i,a)$ is a logit-normally distributed variable and cannot be marginalized, like the choice probability can.


## Simulations

We demonstrate that this model reproduces key behavioral patterns (folded-X, double-increase) under different parameter combinations. Simulations show that the ratio $\frac{\sigma_e}{\sigma_c}$ determines these patterns: 

A low ratio (high $\sigma_c$ low $\sigma_e$) produces folded-X (strong error-monitoring), while a high ratio (low $\sigma_c$ high $\sigma_e$) produces double-increase (weak error-monitoring).

Metacognitive noise $\sigma_m$ flattens confidence curves regardless of accuracy.

```{r}
set.seed(199742)

n_trials <- 500

simulate_interaction <- function(sigma_m, sigma_k,sigma_e) {
  X <- abs(rnorm(n_trials,2,1))
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_k)
  p_gauss <- pnorm(e / sigma_k)
  a <- rbinom(n_trials, 1, p_gauss)

  data.frame(e = e,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}


df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_k = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_k,sigma_e))) %>% 
  unnest(data)

```

Now lets turn to how evidence, the physical stimulus and accuracy interacts across parameters:

```{r, fig.height=6, fig.width=10,warning=F,message=F}

ACC = df_interaction %>%
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = X*D, y = e, col = Accuracy)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point()+
  facet_grid(sigma_e~sigma_k, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))

ACC
```

From the above plot its obvious that $\sigma_e$ scales the the range of evidence observed (see y-axis range for the different rows). When $\sigma_k$ is low (left column) choices are consistent (when the the evidence is positive and the signed stimulus is positive i.e. D = 1 then a correct response is given). When $\sigma_k$ increases this relationship essentially becomes more noisy with correct and incorrect choices being less distinguishable across the evidence = 0 line. Another way to display this is by plotting the actions instead of the accuracy: 

here the space is only split in two quadrants instead of 4.

```{r, fig.height=6, fig.width=10,warning=F,message=F}
Act = df_interaction %>%
  mutate(Action = as.factor(a)) %>% 
  ggplot(aes(x = X*D, y = e, col = Action)) +
  labs(x = "Signed Stimulus", y = "Evidence") +
  geom_point()+
  facet_grid(sigma_e~sigma_k, scales = "free", labeller = label_both)+
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5)+
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)+
  theme_minimal()+
  theme(text = element_text(size = 14))
Act
```


Next we look at the realized confidence ratings as a function of the three different sigmas:

```{r}
set.seed(199742)

n_trials <- 5000

conf_prec <- 100
c0 <- -3
c1 <- 3

simulate_interaction <- function(sigma_m, sigma_k,sigma_e) {
  X <- sample(c(0.01,1,2,5), n_trials, replace = TRUE)
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_k)
  p_gauss <- pnorm(e / sigma_k)
  a <- rbinom(n_trials, 1, p_gauss)
  C_mu <- ifelse(a == 1, 
                 plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)),
                 1 - plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)))
  

  c_mu <- pmax(0.001, pmin(0.999, C_mu))
  C <- ordbetareg::rordbeta(n_trials, c_mu, conf_prec, c(c0, c1))
  
  data.frame(e = e,
             C = C,
             C_mu = C_mu,
             # C_mu2 = C_mu2,
             ehat = ehat,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}

df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_k = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_k,sigma_e))) %>% 
  unnest(data)
```



First we can look to fix $\sigma_k = 1$ and then we look at how $\sigma_e$ & $\sigma_m$ change confidence ratings as a function of the Stimulus:

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_k == 1) %>% 
  mutate(SS = D*X) %>%
  group_by(SS,ACC, sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

From the above plot we see that when we plot confidence on the Signed Stimulus scale we end up with a somewhat double increase pattern, when the stimulus encoding uncertainty is high (i.e. 3 or 5). Furthermore we see a (")half) folded X when stimulus encoding uncertainty  is low. Generally we see that meta uncertainty makes the confidence ratings more flat lined at 0.5 regardless of their correctness.  

Now we fix $\sigma_e = 1$ and then we look at $\sigma_k$ & $\sigma_m$:

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_e == 1) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_k,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_k~ sigma_m, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

This plot generally shows the same as the one with $\sigma_k = 1$ but opposite in pattern. I.e. high $\sigma_k$ here is more of a folded X with lower values (espeically below 1) make a double increase.

lastly we might plot $sigma_m = 1$

```{r, fig.height=6, fig.width=10, message=F, warning=F}
df_interaction %>%
  filter(sigma_m == 1) %>%
  mutate(sigma = sigma_e/sigma_k) %>% 
  mutate(SS = X*D) %>% 
  group_by(SS,ACC, sigma_k,sigma,sigma_e,sigma_m) %>% 
  summarize(mean = mean(C_mu),
            q5 = mean - 2 * sd(C_mu) / sqrt(n()),
            q95 = mean + 2 * sd(C_mu) / sqrt(n())) %>% 
  mutate(Accuracy = as.factor(ACC)) %>%
  mutate(sigma_m = as.factor(sigma_m)) %>% 
  ggplot(aes(x = SS, y = mean, ymin = q5, ymax = q95, col = Accuracy)) +
    geom_pointrange(position = position_dodge(width =0.4))+
  facet_grid( sigma_e~sigma_k, labeller = label_both) +
    geom_hline(yintercept = c(0,0.5,1), linetype = 2)+
  labs(x = "Signed Stimulus", y = "Confidence") +
  theme_minimal()+
  ylim(c(0,1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We see that the diagonal where $\sigma_e = \sigma_k$ that there generally is a half-folded X, with incorrect trials being flat with stimulus strength with correct trials increasing in confidence with stimulus strength. Going to the upper triangle we move into the space where $\sigma_k > \sigma_e$ where we see folded-X patterns of varying "strengths". Moving to the lower triangle we see the opposite pattern with double increase patterns emerging again with varying "strengths". 

This means that from the models perspective double increase and Folded-X patterns of meta cognition arise from the ratio of uncertainty in the encoding of the stimulus and the choice variability. When choice variability is high compared to stimulus encoding a folded X pattern arises whereas choice variability is low compared to stimulus encoding a double increase pattern arises. $\sigma_m$, meta uncertainty is agnostic to the accuracy of the rater and as it increases it makes the curves of the confidence ratings more flat when increased.

## Fitting

We fit this model using full Bayesian inference with Stan's NUTS sampler, treating latent trial-by-trial evidence as parameters estimated jointly with the noise parameters.

## A heuristic model

While the mechanistic model is powerful, its computational demands (especially in hierarchical settings) motivate a simpler heuristic approximation. The key challenge is that confidence requires trial-specific evidence $r_i$, which cannot be marginalized analytically.

**Heuristic approach**: Instead of computing confidence from trial-specific evidence, we compute it from the marginalized psychometric function:

$$P(a = D) = \Phi\left(\frac{|D_i \cdot X_i|}{\sqrt{\sigma^2 + \sigma_m^2}}\right)$$

where $\sigma = \sqrt{\sigma_e^2 + \sigma_c^2}$.

**Critical distinction**: The mechanistic model computes $\mathbb{E}_{r_i}[g(r_i)]$ (average confidence over evidence), while the heuristic computes $g(\mathbb{E}_{r_i}[r_i])$ (confidence from average evidence). By Jensen's inequality, these differ.

**Asymmetry**: Simulations show the heuristic requires different slopes for correct vs. error trials due to different evidence distributions conditional on accuracy. For correct trials: $C_i = g(\beta_{\text{correct}} \cdot |X_i|)$; for errors: $C_i = g(\beta_{\text{error}} \cdot |X_i|)$.

**When to use each**: 
- **Mechanistic**: When fundamental noise parameters are of interest, or when physiological measures constrain evidence
- **Heuristic**: When computational constraints are severe (hierarchical models), or when the mechanistic model fails to converge

## simulations Heuristic model


```{r}
set.seed(199742)

n_trials <- 5000

simulate_interaction <- function(sigma_m, sigma_k,sigma_e) {
  X <- sample(c(0.01,1,2,5), n_trials, replace = TRUE)
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, e, sigma_m)
  k_gauss <- rnorm(n_trials, 0, sigma_k)
  p_gauss <- pnorm(e / sigma_k)
  a <- rbinom(n_trials, 1, p_gauss)
  C_mu <- ifelse(a == 1, 
                 plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)),
                 1 - plogis(2 * ehat * X / (sigma_e^2 + sigma_m^2)))
  
  
  C_mu2 = pnorm(abs(D*X) / sqrt((sigma_e^2 + sigma_m^2)))
  

  data.frame(e = e,
             C_mu = C_mu,
             C_mu2 = C_mu2,
             ehat = ehat,
             a = a,
             X = X,
             D=D,
             ACC = ifelse(a == 0 & D == -1,1,ifelse(a == 1 & D == 1,1,0)))
}

df_interaction <- expand.grid(
  sigma_m = c(0.1, 0.5 ,1,3, 5),
  sigma_k = c(0.1, 0.5 ,1,3, 5),
  sigma_e = c(0.1, 0.5 ,1,3, 5)
  
) %>%
  rowwise() %>%
  mutate(data = list(simulate_interaction(sigma_m, sigma_k,sigma_e))) %>% 
  unnest(data)



```


```{r, fig.height=6, fig.width=10,warning=F,message=F}
# Compute approximation errors
error_summary <- df_interaction %>%
  group_by(sigma_e, sigma_m, sigma_k, ACC, X) %>%
  summarize(
    mean_mechanistic = mean(C_mu),
    mean_heuristic = mean(C_mu2),
    abs_error = abs(mean(C_mu) - mean(C_mu2)),
    n_trials = n(),
    .groups = "drop"
  )

error_summary %>%
  filter(ACC == 1) %>% 
  group_by(sigma_e, sigma_m, sigma_k, ACC) %>%
  summarize(
    mean_abs_error = mean(abs_error),
    q5_abs_error = quantile(abs_error,0.05),
    q95_abs_error = quantile(abs_error,0.95),
    .groups = "drop"
  ) %>% ggplot(aes(x = sigma_e, y = mean_abs_error, col = as.factor(ACC)))+
  geom_pointrange(aes(ymin = q5_abs_error,ymax = q95_abs_error))+
  facet_grid(sigma_k~sigma_m,scales = "free", labeller = label_both)


error_summary %>%
  filter(ACC == 0) %>% 
  group_by(sigma_e, sigma_m, sigma_k, ACC) %>%
  summarize(
    mean_abs_error = mean(abs_error),
    q5_abs_error = quantile(abs_error,0.05),
    q95_abs_error = quantile(abs_error,0.95),
    .groups = "drop"
  ) %>% ggplot(aes(x = sigma_e, y = mean_abs_error, col = as.factor(ACC)))+
  geom_pointrange(aes(ymin = q5_abs_error,ymax = q95_abs_error))+
  facet_grid(sigma_k~sigma_m,scales = "free", labeller = label_both)
```


```{r, fig.height=6, fig.width=10,warning=F,message=F}
df_interaction %>%
  filter(ACC == 1) %>% 
  mutate(dif = C_mu-C_mu2)  %>% 
  ggplot(aes(x = as.factor(sigma_e), y = dif, col = as.factor(ACC)))+
  geom_boxplot()+
  facet_grid(sigma_k~sigma_m,scales = "free", labeller = label_both)




df_interaction %>%
  filter(ACC == 0) %>% 
  mutate(dif = C_mu-C_mu2)  %>% 
  ggplot(aes(x = as.factor(sigma_e), y = dif, col = as.factor(ACC)))+
  geom_boxplot()+
  facet_grid(sigma_k~sigma_m,scales = "free", labeller = label_both)


```







### Mathematical Distinction

The mechanistic model marginalizes after computing confidence (via MCMC), while the heuristic marginalizes before. By Jensen's inequality, these differ. The heuristic requires separate slopes for correct/error trials due to different conditional evidence distributions.

**When to use**: Mechanistic for fundamental parameters; heuristic for hierarchical models or when mechanistic fails to converge.











