---
title: "Confidence with Accuracy Prior: Full Derivation"
author: ""
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: hide
    df_print: paged
---

# Introduction

This document derives the confidence formula when the agent has a **prior belief about their accuracy**. In the basic framework, we assumed equal priors P(D=1) = P(D=-1) = 0.5, which implies the agent believes they are equally likely to be correct or incorrect. Here, we relax this assumption and allow the agent to have a prior belief π about being correct.

# Review: Confidence Without Accuracy Prior

## Basic Setup

- True state: $D \in \{-1, +1\}$
- Stimulus strength: $X > 0$ (unsigned)
- Evidence: $e \sim \mathcal{N}(D \cdot X, \sigma_e^2)$
- Metacognitive read: $\hat{e} \sim \mathcal{N}(e, \sigma_m^2)$
- Action taken: $a \in \{-1, +1\}$

## Confidence Definition

Confidence is the subjective probability of being correct:

$$
\text{confidence} = P(\text{correct} \mid \hat{e}, a) = P(D = a \mid \hat{e})
$$

## Standard Derivation (Equal Priors)

With equal priors P(D=1) = P(D=-1) = 0.5, the posterior for action $a=1$ is:

$$
P(D=1 \mid \hat{e}) = \frac{P(\hat{e} \mid D=1) P(D=1)}{P(\hat{e})}
$$

Since $\hat{e} \mid D \sim \mathcal{N}(D \cdot X, \sigma_e^2 + \sigma_m^2)$, the log-odds are:

$$
\log\frac{P(D=1 \mid \hat{e})}{P(D=-1 \mid \hat{e})} = \log\frac{P(\hat{e} \mid D=1)}{P(\hat{e} \mid D=-1)} + \log\frac{P(D=1)}{P(D=-1)}
$$

The likelihood ratio:
$$
\log\frac{P(\hat{e} \mid D=1)}{P(\hat{e} \mid D=-1)} = \log\frac{\exp\left(-\frac{(\hat{e}-X)^2}{2(\sigma_e^2+\sigma_m^2)}\right)}{\exp\left(-\frac{(\hat{e}+X)^2}{2(\sigma_e^2+\sigma_m^2)}\right)}
$$

$$
= \frac{(\hat{e}+X)^2 - (\hat{e}-X)^2}{2(\sigma_e^2+\sigma_m^2)} = \frac{4\hat{e}X}{2(\sigma_e^2+\sigma_m^2)} = \frac{2\hat{e}X}{\sigma_e^2+\sigma_m^2}
$$

With equal priors, the prior ratio is 0:
$$
\log\frac{P(D=1)}{P(D=-1)} = \log\frac{0.5}{0.5} = 0
$$

Therefore:
$$
P(D=1 \mid \hat{e}) = g\left(\frac{2\hat{e}X}{\sigma_e^2+\sigma_m^2}\right)
$$

where $g(\cdot) = \text{inv\_logit}(\cdot) = \frac{1}{1+\exp(-\cdot)}$

# Introducing the Accuracy Prior

## Motivation

In reality, agents may have **biased beliefs** about their accuracy. For example:
- Overconfident agents believe they are correct more often than they actually are
- Underconfident agents believe they are correct less often than they actually are
- Well-calibrated agents have accurate beliefs about their accuracy

We model this with a parameter $\pi \in [0, 1]$ representing the agent's **prior belief about being correct**.

## Reframing the Problem

Instead of asking "What is P(D=1 | ê)?", we now ask:

**"What is P(correct | ê, a)?"**

where "correct" means the state matches the action (D = a).

## The Prior

The agent has a prior belief:
$$
P(\text{correct}) = \pi
$$
$$
P(\text{incorrect}) = 1 - \pi
$$

This is a **meta-level prior** about accuracy, distinct from priors about stimulus direction.

# Full Derivation with Accuracy Prior

## Step 1: Apply Bayes' Rule

For action $a$, confidence is:

$$
P(\text{correct} \mid \hat{e}, a) = \frac{P(\hat{e} \mid \text{correct}, a) \cdot P(\text{correct})}{P(\hat{e} \mid a)}
$$

By the law of total probability:
$$
P(\hat{e} \mid a) = P(\hat{e} \mid \text{correct}, a) \cdot P(\text{correct}) + P(\hat{e} \mid \text{incorrect}, a) \cdot P(\text{incorrect})
$$

## Step 2: Map to State Space

When action $a=1$:
- "correct" means $D=1$ → $\hat{e} \sim \mathcal{N}(X, \sigma_e^2 + \sigma_m^2)$
- "incorrect" means $D=-1$ → $\hat{e} \sim \mathcal{N}(-X, \sigma_e^2 + \sigma_m^2)$

When action $a=-1$:
- "correct" means $D=-1$ → $\hat{e} \sim \mathcal{N}(-X, \sigma_e^2 + \sigma_m^2)$
- "incorrect" means $D=1$ → $\hat{e} \sim \mathcal{N}(X, \sigma_e^2 + \sigma_m^2)$

## Step 3: Compute Log-Odds

The odds form of Bayes' rule:

$$
\frac{P(\text{correct} \mid \hat{e}, a)}{P(\text{incorrect} \mid \hat{e}, a)} = \frac{P(\hat{e} \mid \text{correct}, a)}{P(\hat{e} \mid \text{incorrect}, a)} \cdot \frac{P(\text{correct})}{P(\text{incorrect})}
$$

Taking logarithms:

$$
\log\frac{P(\text{correct} \mid \hat{e}, a)}{P(\text{incorrect} \mid \hat{e}, a)} = \log\frac{P(\hat{e} \mid \text{correct}, a)}{P(\hat{e} \mid \text{incorrect}, a)} + \log\frac{\pi}{1-\pi}
$$

### Connecting "Correct/Incorrect" to State D

Now we need to express the likelihood terms using the actual state D. For action $a=1$:
- "correct" means the state matches the action: $D = a = 1$
- "incorrect" means the state opposes the action: $D = -a = -1$

Therefore, when $a=1$:
$$
P(\hat{e} \mid \text{correct}, a=1) = P(\hat{e} \mid D=1)
$$
$$
P(\hat{e} \mid \text{incorrect}, a=1) = P(\hat{e} \mid D=-1)
$$

**Important**: Once we condition on the true state $D$, the action $a$ provides no additional information about $\hat{e}$ (the action was already taken based on evidence, but $\hat{e}$ depends only on the true state). Hence we can drop the conditioning on $a$ from the right-hand side.

Substituting back into our log-odds:
$$
\log\frac{P(\text{correct} \mid \hat{e}, a=1)}{P(\text{incorrect} \mid \hat{e}, a=1)} = \log\frac{P(\hat{e} \mid D=1)}{P(\hat{e} \mid D=-1)} + \log\frac{\pi}{1-\pi}
$$

## Step 4: Evaluate the Likelihood Ratio

Now we evaluate the likelihood ratio using the Gaussian distributions. Since $\hat{e} \mid D \sim \mathcal{N}(D \cdot X, \sigma_e^2 + \sigma_m^2)$:

For $D=1$: $\hat{e} \mid D=1 \sim \mathcal{N}(X, \sigma_e^2 + \sigma_m^2)$

For $D=-1$: $\hat{e} \mid D=-1 \sim \mathcal{N}(-X, \sigma_e^2 + \sigma_m^2)$

The likelihood ratio becomes:

$$
\log\frac{P(\hat{e} \mid D=1)}{P(\hat{e} \mid D=-1)} = \log\frac{\exp\left(-\frac{(\hat{e}-X)^2}{2(\sigma_e^2+\sigma_m^2)}\right)}{\exp\left(-\frac{(\hat{e}+X)^2}{2(\sigma_e^2+\sigma_m^2)}\right)}
$$

Expanding the squares:
$$
= \frac{(\hat{e}+X)^2 - (\hat{e}-X)^2}{2(\sigma_e^2+\sigma_m^2)}
$$

$$
= \frac{(\hat{e}^2 + 2\hat{e}X + X^2) - (\hat{e}^2 - 2\hat{e}X + X^2)}{2(\sigma_e^2+\sigma_m^2)}
$$

$$
= \frac{4\hat{e}X}{2(\sigma_e^2+\sigma_m^2)} = \frac{2\hat{e}X}{\sigma_e^2+\sigma_m^2}
$$

This is the **evidence term** - how much the observed $\hat{e}$ supports being correct vs incorrect.

## Step 5: Evaluate the Prior Ratio

The prior odds ratio is:
$$
\log\frac{P(\text{correct})}{P(\text{incorrect})} = \log\frac{\pi}{1-\pi} = \text{logit}(\pi)
$$

This is the **prior term** - the agent's baseline belief before considering evidence.

## Step 6: Combine Terms

The complete log-odds:

$$
\log\frac{P(\text{correct} \mid \hat{e}, a)}{P(\text{incorrect} \mid \hat{e}, a)} = \frac{2\hat{e}X}{\sigma_e^2+\sigma_m^2} + \text{logit}(\pi)
$$

Converting back to probability:

$$
P(\text{correct} \mid \hat{e}, a) = g\left(\frac{2\hat{e}X}{\sigma_e^2+\sigma_m^2} + \text{logit}(\pi)\right)
$$

where $g(x) = \frac{1}{1+e^{-x}}$ is the inverse logit (logistic) function.

# Final Formula

## For Action a=1

$$
\text{confidence} = g\left(\frac{2\hat{e}X}{\sigma_e^2+\sigma_m^2} + \text{logit}(\pi)\right)
$$

## For Action a=-1

Since "correct" now means D=-1, the evidence points in the opposite direction:

$$
\text{confidence} = g\left(\frac{-2\hat{e}X}{\sigma_e^2+\sigma_m^2} + \text{logit}(\pi)\right)
$$

## Unified Formula

For any action $a \in \{-1, +1\}$:

$$
\text{confidence} = g\left(\frac{2a \cdot \hat{e}X}{\sigma_e^2+\sigma_m^2} + \text{logit}(\pi)\right)
$$

# Interpretation

## The Evidence Component: $\frac{2a \cdot \hat{e}X}{\sigma_e^2+\sigma_m^2}$

- **Numerator**: Alignment between evidence and action
  - Large positive: Evidence strongly supports the action → high confidence
  - Near zero: Evidence is weak or ambiguous → low confidence
  - Large negative: Evidence contradicts the action → low confidence

- **Denominator**: Total uncertainty (evidence noise + metacognitive noise)
  - Larger $\sigma_e^2 + \sigma_m^2$: More uncertainty → confidence less sensitive to evidence

## The Prior Component: $\text{logit}(\pi)$

This shifts **all** confidence judgments:

| π | logit(π) | Effect |
|---|----------|--------|
| 0.5 | 0 | Neutral (standard model) |
| 0.7 | 0.85 | Overconfident (+0.85 shift in log-odds) |
| 0.8 | 1.39 | Very overconfident |
| 0.3 | -0.85 | Underconfident |
| 0.2 | -1.39 | Very underconfident |

**Key insight**: The prior acts **additively in log-odds space**, meaning:
- It shifts the entire confidence distribution up or down
- The sensitivity to evidence (slope) remains unchanged
- Only the baseline level of confidence changes

## Comparison to Meta-Bias

This accuracy prior π is mathematically equivalent to what we previously called "meta_bias":

$$
\text{meta_bias} = \text{logit}(\pi)
$$

But framing it as an **accuracy prior** makes the interpretation clearer:
- π = 0.7 means "I believe I'm correct 70% of the time"
- meta_bias = 0.85 is just the log-odds version of this belief

# Simulation Example

```{r, message=FALSE, warning=FALSE}
library(tidyverse)

# Parameters
sigma_e <- 1.0
sigma_m <- 0.6
X <- 0.8  # stimulus strength
D <- 1    # true state
n_trials <- 10000

# Three different accuracy priors
pi_values <- c(0.3, 0.5, 0.7)  # underconfident, calibrated, overconfident

# Simulate
results <- map_dfr(pi_values, function(pi) {
  # Sample evidence and metacognitive read
  e <- rnorm(n_trials, mean = D * X, sd = sigma_e)
  ehat <- rnorm(n_trials, mean = e, sd = sigma_m)
  
  # Compute confidence with accuracy prior
  evidence_term <- 2 * ehat * X / (sigma_e^2 + sigma_m^2)
  prior_term <- qlogis(pi)  # logit(pi)
  
  confidence <- plogis(evidence_term + prior_term)
  
  tibble(
    pi = pi,
    ehat = ehat,
    confidence = confidence,
    correct = (e > 0)  # action matches state when e > 0
  )
})

# Plot confidence distributions
ggplot(results, aes(x = confidence, fill = as.factor(pi))) +
  geom_histogram(alpha = 0.6, bins = 30, position = "identity") +
  facet_wrap(~pi, labeller = label_both) +
  labs(
    title = "Confidence Distributions with Different Accuracy Priors",
    x = "Confidence",
    y = "Count",
    fill = "Accuracy Prior (π)"
  ) +
  theme_minimal()
```

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}
# Plot confidence vs evidence for different priors
results %>%
  mutate(ehat_bin = cut(ehat, breaks = 20)) %>%
  group_by(pi, ehat_bin) %>%
  summarize(
    mean_ehat = mean(ehat),
    mean_conf = mean(confidence),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = mean_ehat, y = mean_conf, color = as.factor(pi))) +
  geom_line(size = 1.2) +
  geom_point() +
  labs(
    title = "Confidence as Function of Evidence for Different Accuracy Priors",
    x = "Metacognitive Evidence (ê)",
    y = "Mean Confidence",
    color = "Accuracy Prior (π)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Key Observations

1. **Parallel curves**: Different priors shift the confidence function up/down but preserve the shape
2. **Same slope**: Sensitivity to evidence (derivative) is unchanged
3. **Global shift**: π affects confidence equally across all evidence levels

# Implications for Modeling

## Parameter Interpretation

When fitting the model to data:

- **σₑ** (evidence noise): Determines slope of psychometric curve
- **σₖ** (criterion noise): Determines choice variability 
- **σₘ** (metacognitive noise): Determines confidence-evidence coupling strength
- **π** (accuracy prior): Determines overall confidence level (overconfidence/underconfidence)

## Identifiability

The accuracy prior π is **identifiable** from:
- Mean confidence level across trials
- Difference between confidence on easy vs hard trials
- Calibration curves (confidence vs actual accuracy)

Without π (fixing it at 0.5), any systematic over/underconfidence would be misattributed to σₘ or other parameters.

## When to Include It

Include π as a free parameter when:
- Subjects show systematic over/underconfidence
- Interest in metacognitive bias separate from metacognitive sensitivity
- Comparing calibration across conditions or subjects

Fix π = 0.5 when:
- Subjects are well-calibrated on average
- Reducing model complexity is important
- Prior knowledge suggests balanced beliefs

# Connection to the Full Model

In your Stan model, implement this as:

```stan
parameters {
  // ... existing parameters ...
  real<lower=0, upper=1> accuracy_prior;  // π
}

transformed parameters {
  real prior_logit = logit(accuracy_prior);
  
  for(i in 1:N) {
    // ... evidence calculations ...
    
    // Confidence with accuracy prior
    real evidence_term = 2 * a[i] * ehat[i] * X[i] / 
                         (exp(sigma_e_log)^2 + exp(sigma_m_log)^2);
    
    mu_conf[i] = inv_logit(evidence_term + prior_logit);
  }
}

model {
  // Prior: centered at 0.5 (well-calibrated) with moderate uncertainty
  accuracy_prior ~ beta(10, 10);
  
  // ... rest of model ...
}
```

This cleanly separates:
- **Metacognitive noise** (σₘ): How well confidence tracks evidence
- **Metacognitive bias** (π): Overall confidence level

# Summary

The accuracy prior π enters the confidence formula as an **additive term in log-odds space**:

$$
\text{confidence} = g\left(\underbrace{\frac{2a \cdot \hat{e}X}{\sigma_e^2+\sigma_m^2}}_{\text{evidence}} + \underbrace{\text{logit}(\pi)}_{\text{prior}}\right)
$$

This formulation:
1. ✅ Is mathematically principled (derived from Bayes' rule)
2. ✅ Has clear interpretation (π = belief about accuracy rate)
3. ✅ Separates bias from sensitivity
4. ✅ Is identifiable from data
5. ✅ Reduces to standard model when π = 0.5
