---
title: "Correlated Evidence Sources for Choice and Confidence"
author: ""
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: hide
    df_print: paged
---

# Introduction

This document implements the correlated evidence framework from Fleming & Daw (2017) "Self-evaluation of decision-making: A general Bayesian framework for metacognitive computation". 

Instead of assuming confidence is based on a noisy readout of the same evidence used for choice, we model **two correlated evidence samples**: one for action ($X_{act}$) and one for confidence ($X_{conf}$).

# Theoretical Framework

## Generative Model

The true state of the world is $D \in \{-1, +1\}$ with stimulus strength $\mu > 0$.

Two evidence samples are drawn from a **bivariate normal distribution**:

$$
\begin{bmatrix} X_{act} \\ X_{conf} \end{bmatrix} \sim \mathcal{N}\left(
\begin{bmatrix} D \cdot \mu \\ D \cdot \mu \end{bmatrix}, 
\begin{bmatrix} \sigma^2 & \rho\sigma^2 \\ \rho\sigma^2 & \sigma^2 \end{bmatrix}
\right)
$$

where:
- $\sigma^2$: variance of evidence (same for both)
- $\rho \in [-1, 1]$: correlation between action and confidence evidence

## Special Cases

- **$\rho = 1$**: Perfect correlation → $X_{act} = X_{conf}$ (standard model, no metacognitive noise)
- **$\rho = 0$**: Independent evidence sources (maximal metacognitive noise)
- **$0 < \rho < 1$**: Partial correlation (realistic metacognition)

## Decision and Confidence

**Choice**: Based on $X_{act}$
$$
a = \begin{cases}
1 & \text{if } X_{act} > k \\
-1 & \text{if } X_{act} \le k
\end{cases}
$$
where $k \sim \mathcal{N}(\mu_k, \sigma_k^2)$ is a stochastic threshold.

**Confidence**: Based on $X_{conf}$ and action $a$
$$
\text{confidence} = P(D = a \mid X_{conf})
$$


Thus:
- $\rho \to 1$ → $\sigma_m \to 0$ (perfect metacognition)
- $\rho \to 0$ → $\sigma_m \to \infty$ (poor metacognition)

# Simulation

```{r setup, message=FALSE, warning=FALSE}
pacman::p_load(tidyverse, patchwork, MASS)

set.seed(1997)
```

## Parameters

```{r parameters}
# Simulation parameters
n_trials <- 1000

# Parameter levels to explore (3x3x3 factorial design)
rho_levels <- c(0.3, 0.6, 0.9)
sigma_k_levels <- c(0.5, 1, 2)
sigma_m_levels <- c(0.5, 1, 2)

# Fixed parameters
mu_k <- 0  # Mean threshold (no bias)
```

## Generate Data Across Conditions

```{r generate_data}
# Create parameter grid
param_grid <- expand_grid(
  rho = rho_levels,
  sigma_k = sigma_k_levels,
  sigma_m = sigma_m_levels
)

# Generate data for all conditions
all_sim_data <- map_dfr(1:nrow(param_grid), function(idx) {
  
  # Extract parameters
  rho <- param_grid$rho[idx]
  sigma_k <- param_grid$sigma_k[idx]
  sigma_m <- param_grid$sigma_m[idx]
  
  # Trial-specific stimulus strengths
  mu <- sample(c(0.5, 1, 2), n_trials, replace = TRUE)
  
  # True states
  D <- sample(c(-1, 1), n_trials, replace = TRUE)
  
  # Construct covariance matrix
  Sigma <- matrix(c(sigma_k^2, rho * sigma_m * sigma_k,
                    rho * sigma_m * sigma_k, sigma_m^2), 
                  nrow = 2)
  
  # Generate correlated evidence samples
  evidence_samples <- mvrnorm(n = n_trials, mu = c(0, 0), Sigma = Sigma)
  
  X_act <- D * mu + evidence_samples[, 1]
  X_conf <- D * mu + evidence_samples[, 2]
  
  # Generate thresholds
  k <- rnorm(n_trials, mean = mu_k, sd = sigma_k)
  
  # Choices
  a <- ifelse(X_act > k, 1, -1)
  
  # Accuracy
  correct <- (a == D)
  
  # Confidence based on X_conf
  conf_prob_D1 <- plogis(2 * X_conf * mu / sigma_m^2)
  confidence <- ifelse(a == 1, conf_prob_D1, 1 - conf_prob_D1)
  
  # Return tibble with condition information
  tibble(
    condition = idx,
    rho = rho,
    sigma_k = sigma_k,
    sigma_m = sigma_m,
    trial = 1:n_trials,
    mu = mu,
    D = D,
    X_act = X_act,
    X_conf = X_conf,
    k = k,
    a = a,
    correct = correct,
    confidence = confidence
  )
})

# Display summary
all_sim_data %>%
  group_by(rho, sigma_k, sigma_m) %>%
  summarise(
    n_trials = n(),
    accuracy = mean(correct),
    mean_conf = mean(confidence),
    .groups = "drop"
  )
```

## Examine Evidence Correlation

```{r evidence_correlation, fig.width=12, fig.height=10}
all_sim_data %>%
  mutate(
    condition_label = sprintf("ρ=%.1f, σk=%.1f, σm=%.1f", rho, sigma_k, sigma_m)
  ) %>%
  filter(condition <= 9) %>%  # Show first 9 conditions for clarity
  ggplot(aes(x = X_act, y = X_conf, color = factor(D))) +
  geom_point(alpha = 0.3, size = 0.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  facet_wrap(~condition_label, ncol = 3) +
  labs(title = "Correlated Evidence Across Conditions (First 9)",
       x = "Action Evidence (X_act)",
       y = "Confidence Evidence (X_conf)",
       color = "True State (D)") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

# Analysis

## Choice Accuracy

```{r choice_accuracy, fig.width=10, fig.height=6}
# Overall accuracy by condition
accuracy_summary <- all_sim_data %>%
  group_by(rho, sigma_k, sigma_m) %>%
  summarise(
    accuracy = mean(correct),
    .groups = "drop"
  )

# Plot accuracy across conditions
p1 <- accuracy_summary %>%
  ggplot(aes(x = factor(sigma_k), y = accuracy, color = factor(rho), group = rho)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  facet_wrap(~sigma_m, labeller = label_both) +
  labs(title = "Accuracy by Threshold Noise and Correlation",
       subtitle = "Facets show different σm levels",
       x = "σk (Threshold Noise)",
       y = "Accuracy",
       color = "ρ") +
  theme_minimal()

p1

# Psychometric function for one example condition
all_sim_data %>%
  filter(rho == 0.6, sigma_k == 1, sigma_m == 1) %>%
  mutate(stim_signed = D * mu) %>%
  group_by(stim_signed) %>%
  summarise(p_choice_up = mean(a == 1),
            n = n()) %>%
  ggplot(aes(x = stim_signed, y = p_choice_up)) +
  geom_point(size = 3) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(title = "Psychometric Function (ρ=0.6, σk=1, σm=1)",
       x = "Signed Stimulus (D × μ)",
       y = "P(Choice = 1)") +
  theme_minimal()
```



## Confidence by Evidence Strength

```{r conf_by_evidence, fig.width=12, fig.height=8}
# Confidence vs X_conf
all_sim_data %>%
  mutate(condition_label = sprintf("ρ=%.1f, σk=%.1f, σm=%.1f", rho, sigma_k, sigma_m)) %>%
  filter(condition %in% c(1, 5, 9, 13, 17, 21, 25, 27)) %>%  # Select subset
  ggplot(aes(x = X_conf, y = confidence, color = correct)) +
  geom_point(alpha = 0.2, size = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~condition_label, ncol = 4) +
  labs(title = "Confidence vs Confidence Evidence (Selected Conditions)",
       x = "X_conf",
       y = "Confidence",
       color = "Correct") +
  theme_minimal()

# Confidence vs stimulus strength
all_sim_data %>%
  mutate(
    stim_strength = D * mu,
    rho_label = sprintf("ρ = %.1f", rho)
  ) %>%
  group_by(rho, sigma_k, sigma_m, stim_strength, correct) %>%
  filter(rho == 0.3) %>% 
  summarise(mean_conf = mean(confidence), q5 = mean_conf - 2*sd(confidence)/sqrt(n()), q95 = mean_conf +2* sd(confidence)/sqrt(n()), .groups = "drop") %>%
  ggplot(aes(x = stim_strength, y = mean_conf, color = correct)) +
  geom_line(size = 1) +
  geom_pointrange(aes(ymin = q5, ymax = q95)) +
  facet_grid(sigma_m ~ sigma_k, labeller = label_both) +
  labs(title = "Mean Confidence vs Stimulus Strength",
       subtitle = "Averaged across ρ levels",
       x = "Signed Stimulus Strength (D × μ)",
       y = "Mean Confidence",
       color = "Correct") +
  theme_minimal()


all_sim_data %>%
  mutate(
    stim_strength = D * mu,
    rho_label = sprintf("ρ = %.1f", rho)
  ) %>%
  group_by(rho, sigma_k, sigma_m, stim_strength, correct) %>%
  filter(rho == 0.9) %>% 
  summarise(mean_conf = mean(confidence), q5 = mean_conf - 2*sd(confidence)/sqrt(n()), q95 = mean_conf +2* sd(confidence)/sqrt(n()), .groups = "drop") %>%
  ggplot(aes(x = stim_strength, y = mean_conf, color = correct)) +
  geom_line(size = 1) +
  geom_pointrange(aes(ymin = q5, ymax = q95)) +
  facet_grid(sigma_m ~ sigma_k, labeller = label_both) +
  labs(title = "Mean Confidence vs Stimulus Strength",
       subtitle = "Averaged across ρ levels",
       x = "Signed Stimulus Strength (D × μ)",
       y = "Mean Confidence",
       color = "Correct") +
  theme_minimal()
```

## Metacognitive Sensitivity

```{r metacog_sensitivity, fig.width=10, fig.height=6}
library(pROC)

# Calculate AUROC for each condition
auroc_results <- all_sim_data %>%
  group_by(rho, sigma_k, sigma_m) %>%
  summarise(
    auroc = auc(roc(correct, confidence, quiet = TRUE)),
    .groups = "drop"
  )

# Plot AUROC across conditions
auroc_results %>%
  ggplot(aes(x = factor(sigma_k), y = as.numeric(auroc), 
             color = factor(rho), group = rho)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  facet_wrap(~sigma_m, labeller = label_both) +
  labs(title = "Metacognitive Sensitivity (Type 2 AUROC)",
       subtitle = "Facets show different σm levels",
       x = "σk (Threshold Noise)",
       y = "Type 2 AUROC",
       color = "ρ") +
  theme_minimal()
```
